---
title: 'SS2 (for AG Hefendehl): Integration'
author: "Desirée Brösamle"
date: "Dec 2021"
output:
  html_document:
    toc: true
---

```{r ,message=FALSE, warning=FALSE, include=FALSE}
home=getwd()

ep_str_wrap <- function(string, width) {
    x <- gregexpr(' ', string)
    vapply(seq_along(x),
           FUN = function(i) {
               y <- x[[i]]
               n <- nchar(string[i])
               len <- (c(y,n) - c(0, y)) ## length + 1
               idx <- len > width
               j <- which(!idx)
               if (length(j) && max(j) == length(len)) {
                   j <- j[-length(j)]
               }
               if (length(j)) {
                   idx[j] <- len[j] + len[j+1] > width
               }
               idx <- idx[-length(idx)] ## length - 1
               start <- c(1, y[idx] + 1)
               end <- c(y[idx] - 1, n)
               words <- substring(string[i], start, end)
               paste0(words, collapse="\n")
           },
           FUN.VALUE = character(1)
    )
}

```


```{r message=FALSE, warning=FALSE, include=FALSE}
library(Seurat)
library(ggplot2)
library(SingleR)
library(EnhancedVolcano)
library(pheatmap)
library(tidyverse)
library(gprofiler2)


knitr::opts_chunk$set(fig.height=8, fig.width=15)

plates_hefendehl<-readRDS("./data/20230203/seu-SCT-harmony_02-02-23.rds")
samples.integrated <- plates_hefendehl



```


# **Preprocessing**

**QC metrics from AG Hefendehl (Stroke)**

Combined datasets and filter cells that have unique feature counts (gene number) over 7000 or less than 200 and cells with >5%  mitochondrial counts.

```{r message=FALSE, warning=FALSE, fig.width=8, fig.height=4}

# Visualize QC metrics as a violin plot
Idents(plates_hefendehl) <- "Plate"
VlnPlot(plates_hefendehl, features = c("nFeature_RNA", "nCount_RNA", "percent.mt","percent.rp"), 
        group.by = "Plate", ncol = 4)

SampleMeta=plates_hefendehl@meta.data
cormt<- cor(SampleMeta$nCount_RNA, SampleMeta$percent.mt, method="spearman")
corfeat<- cor(SampleMeta$nCount_RNA, SampleMeta$nFeature_RNA, method="spearman")

plot1 <- ggplot(plates_hefendehl@meta.data,aes(x=nCount_RNA, y=percent.mt, col=Plate))+
  geom_point()+theme_classic()+geom_abline(slope = 0, intercept = 5)+labs(title=cormt)
plot2 <- ggplot(plates_hefendehl@meta.data,aes(x=nCount_RNA, y=nFeature_RNA, col=Plate))+geom_point()+theme_classic()+geom_abline(slope = 0, intercept = 7000)+
  ggtitle(corfeat)

plot1 | plot2


```


# **Perform integrated analysis**

```{r echo=FALSE, results='hide',message=FALSE,warning=FALSE}
Idents(samples.integrated) <- "Plate"

DimPlot(samples.integrated, reduction = "harmony", group.by = "Mouse_ID")
DimHeatmap(samples.integrated, dims = 1:2, cells = 200, balanced = TRUE)

FeaturePlot(samples.integrated, reduction = "umap", features = c("Aif1", "Ptprc"), dims = c(1,2))


## filter by Microglia Exclude Cells negative for Markers (Aif1 and CD45)
excl = (samples.integrated@assays$RNA@counts["Aif1", ]==0 |
          samples.integrated@assays$RNA@counts["Ptprc", ]==0)

samples.integrated <- samples.integrated[,!excl]

```

PCA, tSNE and UMAP from the first 30 dimensions and with a resolution of 0.8

```{r message=FALSE, warning=FALSE,}
# Visualization

Idents(samples.integrated) <- "seurat_clusters"
DimPlot(samples.integrated, reduction = "umap", split.by = "Treatment")
```


```{r message=FALSE, warning=FALSE,fig.width=12, fig.height=4}
p1<- DimPlot(samples.integrated, reduction = "umap", group.by = "Plate")
p1a<- DimPlot(samples.integrated, reduction = "umap", group.by = "Mouse_ID")
p1b<- DimPlot(samples.integrated, reduction = "umap", group.by = "Genotype_Treatment")
  

p1 + p1a +  p1b
```


```{r message=FALSE, warning=FALSE,fig.width=4, fig.height=4}
DimPlot(samples.integrated, group.by = "seurat_clusters", pt.size =1, label = T) + NoLegend()
```


```{r message=FALSE, warning=FALSE,fig.width=9, fig.height=3}
DimPlot(samples.integrated, reduction = "umap", split.by = "Plate", pt.size= 1)
DimPlot(samples.integrated, reduction = "umap", split.by = "Genotype_Treatment", pt.size =1)

```


**Cell cycle scoring**

First, a score is assigned to each cell (Tirosh et al. 2016), based on its expression of G2/M and S phase markers. These markers should be anticorrelated in their expression levels and cells expression neither are likely not cycling and in G1 phase.
Note: For downstream cell cylce regression the quantitative scores for G2/M and S phase are used, not the dicrete classification.

```{r cell cycle scoring genes, echo=FALSE, results='hide',message=FALSE,warning=FALSE, fig.width=8, fig.height=4}
Idents(samples.integrated) <- "Phase"
# View cell cycle scores and phase assignment
head(samples.integrated[[]])
 
# visualize the distribution of cell cycle markers across
RidgePlot(samples.integrated, features = c("Pcna", "Top2a", "Mcm6", "Mki67"), ncol = 2)
```


```{r cell cycle scoring, echo=FALSE, results='hide',message=FALSE,warning=FALSE, fig.width=4, fig.height=4}
DimPlot(samples.integrated)+ theme_classic()
```


**Cluster distribution**

```{r cluster distribution, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
 
DimPlot(object = samples.integrated, pt.size = 1,reduction = "umap", group.by="Age",label = F) +
  ggtitle("Cluster distribution according to Age") + theme_classic()| DimPlot(object = samples.integrated, pt.size = 1,reduction = "umap", group.by="Genotype_Treatment",label = F) +
  ggtitle("Cluster distribution according to Genotype + Treatment")+ theme_classic()

DimPlot(object = samples.integrated, pt.size = 1,reduction = "umap", group.by="Plate",label = F) +
  ggtitle("Cluster distribution according to Plate")+ theme_classic() | DimPlot(object = samples.integrated, pt.size = 1,reduction = "umap", group.by = "seurat_clusters", label = T) +
  ggtitle("Clustering")+ theme_classic()



```


```{r cell distribution per cluster, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
Idents(samples.integrated) <- "seurat_clusters"

# use same colors for clusters as in plots
require(scales)

samples.integrated$seurat_clusters = factor(samples.integrated$seurat_clusters, levels=sort(as.character(unique(samples.integrated$seurat_clusters))))
identities <- levels(samples.integrated$seurat_clusters) # Create vector with levels of object@ident
cluster_colors <- hue_pal()(length(identities)) # Create vector of default ggplot2 colors


# number of cells in each cluster 
cluster_nCell <- as.data.frame.matrix(table(samples.integrated$seurat_clusters,
                                            samples.integrated$Genotype_Treatment))
cluster_nCell["Total" ,] = colSums(cluster_nCell)
cluster_nCell


# % of cells in each cluster , grouped by genotype_Age

cluster_per_Cell <- data.frame(table(samples.integrated$seurat_clusters, 
                              samples.integrated$Age))
colnames(cluster_per_Cell) <- c("Cluster", "Age", "Frequency")
cluster_per_Cell


cluster_percent_Cell <- data.frame(round((prop.table(x = table(samples.integrated$seurat_clusters, 
                              samples.integrated$Age), margin = 2)*100),2))
colnames(cluster_percent_Cell) <- c("Cluster", "Age", "Frequency")
cluster_percent_Cell


cluster_per_Treatment <- data.frame(table(samples.integrated$seurat_clusters, 
                              samples.integrated$Genotype_Treatment))
colnames(cluster_per_Treatment) <- c("Cluster", "Genotype_Treatment", "Frequency")
cluster_per_Treatment

cluster_percent_Treatment <- data.frame(round((prop.table(x = table(samples.integrated$seurat_clusters, 
                              samples.integrated$Genotype_Treatment), margin = 2)*100),2))
colnames(cluster_percent_Treatment) <- c("Cluster", "Genotype_Treatment", "Frequency")
cluster_percent_Treatment
```


```{r plot cell distribution per cluster, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
# Grouped barchart of cell proportions
a1<- ggplot(cluster_percent_Cell, aes(fill=Age, 
                                 y=Frequency, 
                                 x=Cluster)) +
  geom_bar(position="dodge", stat="identity")+
  ggtitle("Cell distribution according to Age [%]") +theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=1)) +
  xlab("")


# Grouped barchart of cell proportions
b1 <- ggplot(cluster_percent_Cell, aes(fill=Cluster, y=Frequency, x=Age)) + 
  geom_bar(position="dodge", stat="identity")+
  ggtitle("Cell distribution according to Age [%]") +theme_classic()+
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=1))


c1 <-ggplot(cluster_percent_Cell, aes(fill=Cluster, y=Frequency, x=Age)) + 
    geom_bar(position="stack", stat="identity")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=1))

a1 | b1 | c1

```
```{r plot cell distribution per GenoTReamtne, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
# Grouped barchart of cell proportions
a1<- ggplot(cluster_percent_Treatment, aes(fill=Genotype_Treatment, 
                                 y=Frequency, 
                                 x=Cluster)) +
  geom_bar(position="dodge", stat="identity")+
  ggtitle("Cell distribution according to Genotype_Treatment [%]") +theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=1)) +
  xlab("")


# Grouped barchart of cell proportions
b1 <- ggplot(cluster_percent_Treatment, aes(fill=Cluster, y=Frequency, x=Genotype_Treatment)) + 
  geom_bar(position="dodge", stat="identity")+
  ggtitle("Cell distribution according to Genotype_Treatment [%]") +theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=1))


c1 <-ggplot(cluster_percent_Treatment, aes(fill=Cluster, y=Frequency, x=Genotype_Treatment)) + 
    geom_bar(position="stack", stat="identity")+theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=1))

a1 | b1 | c1

```

# **Cell type identification**

## **Finding differentially expressed features (cluster biomarkers)**

```{r Marker genes, message=FALSE, warning=FALSE, fig.width=8, fig.height=8}

Idents(samples.integrated) <- samples.integrated$seurat_clusters
samples.integrated=PrepSCTFindMarkers(samples.integrated)

markers <- FindAllMarkers(object = samples.integrated, 
                          only.pos = TRUE,
                          logfc.threshold = 0.25)  


DefaultAssay(samples.integrated) = "SCT"

markers %>%
    group_by(cluster) %>%
    top_n(n = 20, wt = avg_log2FC) -> top20

DoHeatmap(samples.integrated, features = top20$gene) + NoLegend() + ggtitle("Top20 cluster marker genes")
```


## **SingleR**
SingleR is an automatic annotation method for (scRNAseq) data (Aran et al. 2019). Given a reference dataset of samples (single-cell or bulk) with known labels, it labels new cells from a test dataset based on similarity to the reference set. Here we use the built-in references "Immgen" (830 microarray samples of sorted hematopoetic and immune cell populations) 	 and "Mouse RNA-Seq" (358 non-specific mouse RNA-seq samples). 

**Immgen reference**


```{r SingleR: Immgen, echo=FALSE, results='hide',message=FALSE,warning=FALSE,   fig.width=8, fig.height=4}

# Based on clusters

#table(pred.immgen$first.labels) #main cell types
#table(pred.immgen$labels) #fine tuned cell types
#table(pred.immgen$pruned.labels) #cell type after pruning

#plotScoreHeatmap(pred.immgen, show.labels = TRUE,
#                 annotation_col=data.frame(cluster=unique(samples.integrated@meta.data$seurat_clusters),
#                                           row.names=rownames(pred.immgen)))

# Based on single cells
  
#table(pred.immgen.sc$first.labels)
#table(pred.immgen.sc$labels)

#plotScoreHeatmap(pred.immgen.sc)

#Visualization
DimPlot(samples.integrated,group.by ="Immgen_cluster_labels" ,label = F) | DimPlot(samples.integrated, reduction = "umap",group.by = "seurat_clusters", label = T) 
DimPlot(samples.integrated,group.by ="Immgen_sc_labels", label = F ) + NoLegend() | DimPlot(samples.integrated, reduction = "umap",group.by = "seurat_clusters", label = T)


```

**MouseRNA-Seq reference**
Section Skipped as pred.mouseRNA object is not available 

```{r SingleR: MouseRNA-Seq, echo=FALSE, results='hide',message=FALSE,warning=FALSE,   fig.width=8, fig.height=4}

# Based on clusters

#table(pred.mouseRNA$first.labels) #main cell types
#table(pred.mouseRNA$labels) #fine tuned cell types
#table(pred.mouseRNA$pruned.labels) #cell type after pruning

# plotScoreHeatmap(pred.mouseRNA, show.labels = TRUE,
#                  annotation_col=data.frame(cluster=unique(samples.integrated$seurat_clusters),
#                                            row.names=rownames(pred.mouseRNA)))
# 
# Based on single cells
  
# table(pred.mouseRNA.sc$first.labels)
# table(pred.mouseRNA.sc$labels)
# 
# plotScoreHeatmap(pred.mouseRNA.sc)
# 
#Visualization

DimPlot(samples.integrated,group.by ="MouseRNASeq_cluster_labels" ,label = T) + NoLegend() | DimPlot(samples.integrated, reduction = "umap",group.by = "seurat_clusters", label=T)+ NoLegend()
DimPlot(samples.integrated,group.by ="MouseRNASeq_sc_labels", label=F) + NoLegend() | DimPlot(samples.integrated, reduction = "umap",group.by = "seurat_clusters", label=T)+ NoLegend()

```


```{r cell distribution MouseRNASeq, message=FALSE, warning=FALSE, fig.with=6, fig.height=6}

# number of cells in each cluster 
cluster_nCell <- data.frame(table(samples.integrated$MouseRNASeq_sc_labels,samples.integrated$Genotype_Treatment))
colnames(cluster_nCell) <- c( "MouseRNASeq_sc_labels", "Genotype_Treatment","Number")

#cluster_nCell["Total" ,] = colSums(cluster_nCell)
# cluster_nCell

# % of cells in each cluster 
cluster_percent_Cell <- data.frame(round((prop.table(x = table(samples.integrated$MouseRNASeq_sc_labels, 
                              samples.integrated$Genotype_Treatment), margin = 2)*100),2))
colnames(cluster_percent_Cell) <- c("MouseRNASeq_sc_labels", "Genotype_Treatment", "Frequency")
# cluster_percent_Cell


# Grouped barchart of absolute cell numbers
ggplot(cluster_nCell, aes(fill=Genotype_Treatment, y=Number, x=MouseRNASeq_sc_labels)) + 
  geom_bar(position="dodge", stat="identity") +
  ggtitle("Cell distribution according to MouseRNASeq reference (absolut values)")+ theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=0.5)) +
  xlab("")

# Grouped barchart of cell proportions
ggplot(cluster_percent_Cell, aes(fill=Genotype_Treatment, y=Frequency, x=MouseRNASeq_sc_labels)) + 
  geom_bar(position="dodge", stat="identity")+
  ggtitle("Cell distribution according to MouseRNASeq reference [%]")+ theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=0.5)) +
  xlab("")

```

```{r cell distribution immgen, message=FALSE, warning=FALSE, fig.with=8, fig.height=6}

# number of cells in each cluster 
cluster_nCell <- data.frame(table(samples.integrated$Immgen_sc_labels,samples.integrated$Genotype_Treatment))
colnames(cluster_nCell) <- c( "Immgen_sc_labels", "Genotype_Treatment","Number")

#cluster_nCell["Total" ,] = colSums(cluster_nCell)
# cluster_nCell

# % of cells in each cluster 
cluster_percent_Cell <- data.frame(round((prop.table(x = table(samples.integrated$Immgen_sc_labels, 
                              samples.integrated$Genotype_Treatment), margin = 2)*100),2))
colnames(cluster_percent_Cell) <- c( "Immgen_sc_labels", "Genotype_Treatment","Frequency")
# cluster_percent_Cell


# Grouped barchart of absolute cell numbers
ggplot(cluster_nCell, aes(fill=Genotype_Treatment, y=Number, x=Immgen_sc_labels)) + 
  geom_bar(position="dodge", stat="identity") +
  ggtitle("Cell distribution according to Immgen reference (absolut values)") + theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=0.5)) +
  xlab("")

# Grouped barchart of cell proportions
ggplot(cluster_percent_Cell, aes(fill=Genotype_Treatment, y=Frequency, x=Immgen_sc_labels)) + 
  geom_bar(position="dodge", stat="identity")+
  ggtitle("Cell distribution according to Immgen reference [%]") + theme_classic()+
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=0.5)) +
  xlab("")

```
Renaming Clusters

```{r}
Idents(samples.integrated) = "Celltype"
DimPlot(samples.integrated, reduction = "umap", label=T) + NoLegend()+ theme_classic()
```


# **Zoom into Microglia clusters**

A second Seurat cluster is generated, where all non-microglial immune cells are removed according to the mouseRNASeq_sc_labels.

```{r subclustering microglia, message=FALSE, warning=FALSE, fig.with=8, fig.height=4}

samples.microglia <- samples.integrated[,grepl("Microgli.*", samples.integrated$MouseRNASeq_sc_labels)]

p3 <- DimPlot(samples.microglia, reduction = "umap", label=T, group.by = "MouseRNASeq_sc_labels")  + theme_classic()+ NoLegend()

p3a <- DimPlot(samples.microglia, reduction = "umap", label=T, group.by = "seurat_clusters") + theme_classic()+ NoLegend()

p3 + p3a

```


## **DAMs (Cell. 2017 Jun 15;169(7):1276-1290.e17. doi: 10.1016/j.cell.2017.05.018)**

**DAM plotting**

```{r Stage 2 DAMs, echo=FALSE, results='hide',message=FALSE,warning=FALSE, fig.width=8, fig.height=4}

load("./data/DAM_genelists.RData")

samples.microglia <- AddModuleScore(object = samples.microglia, 
                                    features = list(DAM_up), name = "DAM_score")
p4 <- FeaturePlot(object = samples.microglia, features = "DAM_score1")+scale_color_viridis_c()
p4a <- DimPlot(object = samples.microglia, reduction = "umap")

p4 | p4a

```

```{r,  fig.width=8, fig.height=4}

DAM2_marker_gene_list <- list(Stage2_DAM_up)
samples.microglia <- AddModuleScore(object = samples.microglia, 
                                    features = DAM2_marker_gene_list, name = "DAM2_score")
p5 <- FeaturePlot(object = samples.microglia, features = "DAM2_score1")+scale_color_viridis_c()
p5a <- DimPlot(object = samples.microglia, reduction = "umap")

p5 | p5a

```

**DAM 2 Markers**

**DAM2 upreagulated markers**
```{r echo=FALSE, message=FALSE,results='hide',warning=FALSE}
VlnPlot(samples.microglia, features = DAM2_up_1) 
VlnPlot(samples.microglia, features = DAM2_up_2) 
VlnPlot(samples.microglia, features = DAM2_up_3)
```

**DAM2 dowreagulated markers**
```{r echo=FALSE, message=FALSE,results='hide',warning=FALSE}
VlnPlot(samples.microglia, features = DAM2_down_1)
VlnPlot(samples.microglia, features = DAM2_down_2) 
VlnPlot(samples.microglia, features = DAM2_down_3) 
```

## **Renaming clusters**

```{r cluster renaming, message=FALSE,warning=FALSE}

Idents(samples.microglia) <- "Celltype"
DimPlot(samples.microglia, reduction = "umap", label = TRUE, pt.size = 1)+ NoLegend()

```


## **Signature enrichment**

All enrichment tests were done with gprofiler 2 

Tested Ontologies and Signatures 
GO:MF = Molecular Function
GO:BP = Biological Processes
GO:CC = Cellular Compartment),
KEGG = pathways from KEGG Reactome, 
TF = regulatory motif matches from TRANSFAC
HPA = tissue specificity from Human Protein Atlas;
CORUM = protein complexes from CORUM 
HP = human disease phenotypes  from Human Phenotype Ontology. 

For more dteails see 
Website: https://biit.cs.ut.ee/gprofiler/gost"


### **DEGs and GSEA across microglia of different genotypes**

log fold-change of the average expression between the two groups: Positive values indicate that the gene is more highly expressed in the target group (e.g. the APPPS1+ group)


#### **WT vs App**

```{r general DEGs WT vs APPPS1, echo=FALSE, message=FALSE,warning=FALSE,   fig.width=8, fig.height=8}

Idents(samples.microglia) <- "Genotype"

# WT vs App
samples.microglia <- PrepSCTFindMarkers(samples.microglia)
markers_WtvsApp <- FindMarkers(samples.microglia, ident.1 = "APPPS1")

library(DESeq2)

DT::datatable(markers_WtvsApp, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of Genotype"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))

EnhancedVolcano(markers_WtvsApp,
                lab = rownames(markers_WtvsApp),
                x = 'avg_log2FC', 
                y = 'p_val_adj', xlim = c(-1.5, 1.5), ylim=c(0,15), pCutoff = 0.05,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(Padj)),
                title = "DEGs", subtitle = "Wt vs APPPS1+",
                caption = "pval-adj cutoff  0.05 \n FC cutoff  1.5")


sig.markers_WtvsApp <- markers_WtvsApp[abs(markers_WtvsApp$avg_log2FC)>log(1.5,2) 
                                       &markers_WtvsApp$p_val_adj<0.05,]


DT::datatable(sig.markers_WtvsApp, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of Genotype"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))




genelist=split(rownames(sig.markers_WtvsApp), f = sig.markers_WtvsApp$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("APPPS1_down", "APPPS1_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]


DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of Genotype"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))

```

#### **WT vs MX04+**
log fold-chage of the average expression between the two groups: Positive values indicate that the gene is more highly expressed in the Methoxy positive group

```{r echo=FALSE, echo=FALSE, message=FALSE,warning=FALSE,   fig.width=8, fig.height=8}
Idents(samples.microglia) <- "Methoxy" 
markers_WtvsMX <- FindMarkers(samples.microglia, ident.1 = "pos")


EnhancedVolcano(markers_WtvsMX,
                lab = rownames(markers_WtvsMX),
                x = 'avg_log2FC', 
                y = 'p_val_adj', xlim = c(-1.5, 1.5), ylim=c(0,15), pCutoff = 0.05,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(Padj)),
                title = "DEGs", subtitle = "MX04+ vs MX04-",
                caption = "pval-adj cutoff  0.05 \n FC cutoff  1.5")


sig.markers_WtvsMX <- markers_WtvsMX[abs(markers_WtvsMX$avg_log2FC)>log(1.5,2) 
                                       &markers_WtvsMX$p_val_adj<0.05,]


DT::datatable(sig.markers_WtvsMX, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of Genotype"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))




genelist=split(rownames(sig.markers_WtvsMX), f = sig.markers_WtvsMX$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("Methoxy_down", "Methoxy_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]


DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of MX04+ across Genotypes"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))

```

#### **MX04+ vs MX04- in APPPS1 cohort only**

```{r echo=FALSE, message=FALSE,warning=FALSE,   fig.width=8, fig.height=8}
samples.microglia.app <- samples.microglia[,samples.microglia$Genotype=="APPPS1"]

samples.microglia.app <- SCTransform(samples.microglia.app)
samples.microglia.app <- RunPCA(samples.microglia.app)
samples.microglia.app <- RunUMAP(samples.microglia.app, dims = 1:40)

markers_MXAPP <- FindMarkers(samples.microglia.app, ident.1 = "pos")


EnhancedVolcano(markers_MXAPP,
                lab = rownames(markers_MXAPP),
                x = 'avg_log2FC', 
                y = 'p_val_adj', xlim = c(-1.5, 1.5), ylim=c(0,15), pCutoff = 0.05,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(Padj)),
                title = "DEGs", subtitle = "MX04+ vs MX04- in APPPS1+",
                caption = "pval-adj cutoff  0.05 \n FC cutoff  1.5")


sig.markers_MXAPP <- markers_MXAPP[abs(markers_MXAPP$avg_log2FC)>log(1.5,2) 
                                       &markers_MXAPP$p_val_adj<0.05,]


DT::datatable(sig.markers_MXAPP, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of MX04+ in APPPS1+"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))




genelist=split(rownames(sig.markers_MXAPP), f = sig.markers_MXAPP$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("MX04pos_APPPS_down", "MX04pos_APPPS_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]


DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of MX04pos in APPPS"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))

```



### **DEGs and GSEA of microglia of different genotype and treatment**

#### **WT_Ctrl vs WT_Stroke**

```{r echo=FALSE, message=FALSE,warning=FALSE,  fig.width=8, fig.height=8}

samples.microglia.WT <- samples.microglia[,samples.microglia$Genotype=="WT"]

samples.microglia.WT <- SCTransform(samples.microglia.WT)
samples.microglia.WT <- RunPCA(samples.microglia.WT)
samples.microglia.WT <- RunUMAP(samples.microglia.WT, dims = 1:40)

Idents(samples.microglia.WT) <- "Treatment"
markers_StrokeWT <- FindMarkers(samples.microglia.WT, ident.1 = "Stroke")


EnhancedVolcano(markers_StrokeWT,
                lab = rownames(markers_StrokeWT),
                x = 'avg_log2FC', 
                y = 'p_val_adj', xlim = c(-1.5, 1.5), ylim=c(0,15), pCutoff = 0.05,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(Padj)),
                title = "DEGs", subtitle = "Stroke- vs Stroke+ in WT",
                caption = "pval-adj cutoff  0.05 \n FC cutoff  1.5")


sig.markers_StrokeWT <- markers_StrokeWT[abs(markers_StrokeWT$avg_log2FC)>log(1.5,2) 
                                       &markers_StrokeWT$p_val_adj<0.05,]


DT::datatable(sig.markers_StrokeWT, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of Stroke in WT"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))


genelist=split(rownames(sig.markers_StrokeWT), f = sig.markers_StrokeWT$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("WT_Stroke_down", "WT_Stroke_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]

DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of Stroke in WT"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))

```


#### **APP_Ctrl vs APP_Stroke**

```{r echo=FALSE, message=FALSE,warning=FALSE,   fig.width=8, fig.height=8}

Idents(samples.microglia.app) <- "Treatment"
markers_StrokeAPPPS <- FindMarkers(samples.microglia.app, ident.1 = "Stroke")


EnhancedVolcano(markers_StrokeAPPPS,
                lab = rownames(markers_StrokeAPPPS),
                x = 'avg_log2FC', 
                y = 'p_val_adj', xlim = c(-1.5, 1.5), ylim=c(0,15), pCutoff = 0.05,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(Padj)),
                title = "DEGs", subtitle = "Stroke+ vs Stroke- in WT",
                caption = "pval-adj cutoff  0.05 \n FC cutoff  1.5")


sig.markers_StrokeAPPPS <- markers_StrokeAPPPS[abs(markers_StrokeAPPPS$avg_log2FC)>log(1.5,2) 
                                       &markers_StrokeAPPPS$p_val_adj<0.05,]


DT::datatable(sig.markers_StrokeAPPPS, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of Stroke in APPPS"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))




genelist=split(rownames(sig.markers_StrokeAPPPS), f = sig.markers_StrokeAPPPS$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("APPPS_Stroke_down", "APPPS_Stroke_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]


DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of STroke in APPPS1+"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))

```


#### **WT_Ctrl vs APP_Ctrl**

```{r echo=FALSE, message=FALSE,warning=FALSE,   fig.height=8, fig.width=8}

samples.microglia.Ctrl <- samples.microglia[,samples.microglia$Treatment=="Ctrl"]

samples.microglia.Ctrl <- SCTransform(samples.microglia.Ctrl)
samples.microglia.Ctrl <- RunPCA(samples.microglia.Ctrl)
samples.microglia.Ctrl <- RunUMAP(samples.microglia.Ctrl, dims = 1:40)

Idents(samples.microglia.Ctrl) <- "Genotype"
markers_GenotypeCtrl <- FindMarkers(samples.microglia.Ctrl, ident.1 = "APPPS1")


EnhancedVolcano(markers_GenotypeCtrl,
                lab = rownames(markers_GenotypeCtrl),
                x = 'avg_log2FC', 
                y = 'p_val_adj', xlim = c(-1.5, 1.5), ylim=c(0,15), pCutoff = 0.05,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(Padj)),
                title = "DEGs", subtitle = "APPPS1- vs APPPS1+ in Ctrl",
                caption = "pval-adj cutoff  0.05 \n FC cutoff  1.5")


sig.markers_GenotypeCtrl <- markers_GenotypeCtrl[abs(markers_GenotypeCtrl$avg_log2FC)>log(1.5,2) 
                                       &markers_GenotypeCtrl$p_val_adj<0.05,]


DT::datatable(sig.markers_GenotypeCtrl, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of APPPS1 in Ctrl"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))




genelist=split(rownames(sig.markers_GenotypeCtrl), f = sig.markers_GenotypeCtrl$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("Ctrl_APPPS1_down", "Ctrl_APPPS1_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]

DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of APPPS1 in Ctrls"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))


```


#### **WT_Stroke vs APP_Stroke**


```{r echo=FALSE, message=FALSE,warning=FALSE,   fig.width=8, fig.height=8} 


samples.microglia.Stroke <- samples.microglia[,samples.microglia$Treatment=="Stroke"]

samples.microglia.Stroke <- SCTransform(samples.microglia.Stroke)
samples.microglia.Stroke <- RunPCA(samples.microglia.Stroke)
samples.microglia.Stroke <- RunUMAP(samples.microglia.Stroke, dims = 1:40)

Idents(samples.microglia.Stroke) <- "Genotype"
markers_GenotypeStroke <- FindMarkers(samples.microglia.Stroke, ident.1 = "APPPS1")


EnhancedVolcano(markers_GenotypeStroke,
                lab = rownames(markers_GenotypeStroke),
                x = 'avg_log2FC', 
                y = 'p_val_adj', xlim = c(-1.5, 1.5), ylim=c(0,15), pCutoff = 0.05,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(Padj)),
                title = "DEGs", subtitle = "APPPS1- vs APPPS1+ in Stroke",
                caption = "pval-adj cutoff  0.05 \n FC cutoff  1.5")


sig.markers_GenotypeStroke <- markers_GenotypeStroke[abs(markers_GenotypeStroke$avg_log2FC)>log(1.5,2) 
                                       &markers_GenotypeStroke$p_val_adj<0.05,]


DT::datatable(sig.markers_GenotypeStroke, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of APPPS1 in Stroke"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))




genelist=split(rownames(sig.markers_GenotypeStroke), f = sig.markers_GenotypeStroke$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("Stroke_APPPS1_down", "Stroke_APPPS1_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]

DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of APPPS1 in Strokes"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))

```


### **DEGs and GSEA of each cluster**

log fold-chage of the average expression between the two groups: Positive values indicate that the gene is more highly expressed in the target group (e.g. Microglia 0) versus all others. 

#### **Microglia_0**


```{r DEGs GSEA_output_ctrl Microglia_0, echo=FALSE, message=FALSE,warning=FALSE,   fig.height=8, fig.width=8}

Idents(samples.microglia)="seurat_clusters"
markers.Microglia_0 = FindMarkers(samples.microglia, ident.1 = "0")

EnhancedVolcano(markers.Microglia_0,
                lab = rownames(markers.Microglia_0),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-3, 3), ylim=c(0,55), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Microglia_0",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")


sig.markers.Microglia_0 <- markers.Microglia_0[abs(markers.Microglia_0$avg_log2FC)>log(1.5,2) 
                                       &markers.Microglia_0$p_val_adj<0.05,]

DT::datatable(sig.markers.Microglia_0, extensions = "Buttons",
              filter="top",
              caption = paste0("Microglia_0 markers"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))




genelist=split(rownames(sig.markers_GenotypeCtrl), f = sig.markers_GenotypeCtrl$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("Microglia_0_down", "Microglia_0_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]

DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Microglia_0_Marker_Enrichment"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))

```


#### **Microglia_1**

```{r DEGs GSEA_output_ctrl Microglia_1, echo=FALSE, message=FALSE,warning=FALSE,   fig.height=8, fig.width=8}

Idents(samples.microglia)="seurat_clusters"
markers.Microglia_1 = FindMarkers(samples.microglia, ident.1 = "1")

EnhancedVolcano(markers.Microglia_1,
                lab = rownames(markers.Microglia_1),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-3, 3), ylim=c(0,55), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Microglia_1",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")


sig.markers.Microglia_1 <- markers.Microglia_1[abs(markers.Microglia_1$avg_log2FC)>log(1.5,2) 
                                       &markers.Microglia_1$p_val_adj<0.05,]

DT::datatable(sig.markers.Microglia_1, extensions = "Buttons",
              filter="top",
              caption = paste0("Microglia_1 markers"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))




genelist=split(rownames(sig.markers_GenotypeCtrl), f = sig.markers_GenotypeCtrl$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("Microglia_1_down", "Microglia_1_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]

DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Microglia_1_Marker_Enrichment"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))
```


#### **Microglia_2**
```{r DEGs GSEA_output_ctrl Microglia_2, echo=FALSE, message=FALSE,warning=FALSE,   fig.height=8, fig.width=8}

Idents(samples.microglia)="seurat_clusters"
markers.Microglia_2 = FindMarkers(samples.microglia, ident.1 = "2")

EnhancedVolcano(markers.Microglia_2,
                lab = rownames(markers.Microglia_2),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-3, 3), ylim=c(0,55), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Microglia_2",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")


sig.markers.Microglia_2 <- markers.Microglia_2[abs(markers.Microglia_2$avg_log2FC)>log(1.5,2) 
                                       &markers.Microglia_2$p_val_adj<0.05,]

DT::datatable(sig.markers.Microglia_2, extensions = "Buttons",
              filter="top",
              caption = paste0("Microglia_2 markers"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))




genelist=split(rownames(sig.markers_GenotypeCtrl), f = sig.markers_GenotypeCtrl$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("Microglia_2_down", "Microglia_2_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]

DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Microglia_2_Marker_Enrichment"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))

```


#### **Microglia_3**
```{r DEGs GSEA_output_ctrl Microglia_3, echo=FALSE, message=FALSE,warning=FALSE,   fig.height=8, fig.width=8}

Idents(samples.microglia)="seurat_clusters"
markers.Microglia_3 = FindMarkers(samples.microglia, ident.1 = "3")

EnhancedVolcano(markers.Microglia_3,
                lab = rownames(markers.Microglia_3),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-3, 3), ylim=c(0,55), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Microglia_3",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")


sig.markers.Microglia_3 <- markers.Microglia_3[abs(markers.Microglia_3$avg_log2FC)>log(1.5,2) 
                                       &markers.Microglia_3$p_val_adj<0.05,]

DT::datatable(sig.markers.Microglia_3, extensions = "Buttons",
              filter="top",
              caption = paste0("Microglia_3 markers"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))




genelist=split(rownames(sig.markers_GenotypeCtrl), f = sig.markers_GenotypeCtrl$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("Microglia_3_down", "Microglia_3_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]

DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Microglia_3_Marker_Enrichment"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))

```


#### **Microglia_4**
```{r DEGs GSEA_output_ctrl Microglia_4, echo=FALSE, message=FALSE,warning=FALSE,   fig.height=8, fig.width=8}

Idents(samples.microglia)="seurat_clusters"
markers.Microglia_4 = FindMarkers(samples.microglia, ident.1 = "4")

EnhancedVolcano(markers.Microglia_4,
                lab = rownames(markers.Microglia_4),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-3, 3), ylim=c(0,55), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Microglia_4",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")


sig.markers.Microglia_4 <- markers.Microglia_4[abs(markers.Microglia_4$avg_log2FC)>log(1.5,2) 
                                       &markers.Microglia_4$p_val_adj<0.05,]

DT::datatable(sig.markers.Microglia_4, extensions = "Buttons",
              filter="top",
              caption = paste0("Microglia_4 markers"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))




genelist=split(rownames(sig.markers_GenotypeCtrl), f = sig.markers_GenotypeCtrl$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("Microglia_4_down", "Microglia_4_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]

DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Microglia_4_Marker_Enrichment"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))

```


#### **Microglia_5**
```{r DEGs GSEA_output_ctrl Microglia_5, echo=FALSE, message=FALSE,warning=FALSE,   fig.height=8, fig.width=8}

Idents(samples.microglia)="seurat_clusters"
markers.Microglia_5 = FindMarkers(samples.microglia, ident.1 = "5")

EnhancedVolcano(markers.Microglia_5,
                lab = rownames(markers.Microglia_5),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-3, 3), ylim=c(0,55), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Microglia_5",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")


sig.markers.Microglia_5 <- markers.Microglia_5[abs(markers.Microglia_5$avg_log2FC)>log(1.5,2) 
                                       &markers.Microglia_5$p_val_adj<0.05,]

DT::datatable(sig.markers.Microglia_5, extensions = "Buttons",
              filter="top",
              caption = paste0("Microglia_5 markers"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))




genelist=split(rownames(sig.markers_GenotypeCtrl), f = sig.markers_GenotypeCtrl$avg_log2FC>0)
genelist = genelist[c("FALSE", "TRUE")]
names(genelist) = c("Microglia_5_down", "Microglia_5_up")

resgost = gost(genelist, organism = "mmusculus", evcodes = T)

gostplot(resgost)

restab <- resgost$result[,c("query", "p_value", "term_id", "source", "term_name", "intersection")]

DT::datatable(restab, extensions = "Buttons",
              filter="top",
              caption = paste0("Microglia_5_Marker_Enrichment"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))

```


## Genotype + Treatment comparison in each cluster

```{r prep data}
Idents(samples.microglia) <- "Genotype_Treatment"

mg1 <- samples.microglia[,samples.microglia$Celltype == "Microglia_1"]
mg1 <- SCTransform(mg1)
mg1.markers.Genotype_Treatment <- FindAllMarkers(mg1)
if(length(mg1.markers.Genotype_Treatment)!=0){
mg1_top20 = mg1.markers.Genotype_Treatment %>% group_by(cluster) %>% top_n(n=20, wt=abs(avg_log2FC))
mg1_top20chk=T
} else {mg1_top20chk=F}

mg2 <- samples.microglia[,samples.microglia$Celltype == "Microglia_2"]
mg2 <- SCTransform(mg2)
mg2.markers.Genotype_Treatment <- FindAllMarkers(mg2)
if(length(mg2.markers.Genotype_Treatment)!=0){
mg2_top20 = mg2.markers.Genotype_Treatment %>% group_by(cluster) %>% top_n(n=20, wt=abs(avg_log2FC))
mg2_top20chk=T
} else {mg2_top20chk=F}

mg3 <- samples.microglia[,samples.microglia$Celltype == "Microglia_3"]
mg3 <- SCTransform(mg3)
mg3.markers.Genotype_Treatment <- FindAllMarkers(mg3)
if(length(mg3.markers.Genotype_Treatment)!=0){
mg3_top20 = mg3.markers.Genotype_Treatment %>% group_by(cluster) %>% top_n(n=20, wt=abs(avg_log2FC))
mg3_top20chk=T
} else {mg3_top20chk=F}

mg4 <- samples.microglia[,samples.microglia$Celltype == "Microglia_4"]
mg4 <- SCTransform(mg4)
mg4.markers.Genotype_Treatment <- FindAllMarkers(mg4)
if(length(mg4.markers.Genotype_Treatment)!=0){
mg4_top20 = mg4.markers.Genotype_Treatment %>% group_by(cluster) %>% top_n(n=20, wt=abs(avg_log2FC))
mg4_top20chk=T
} else {mg4_top20chk=F}

mg5 <- samples.microglia[,samples.microglia$Celltype == "Microglia_5"]
mg5 <- SCTransform(mg5)
mg5.markers.Genotype_Treatment <- FindAllMarkers(mg5)
if(length(mg5.markers.Genotype_Treatment)!=0){
mg5_top20 = mg5.markers.Genotype_Treatment %>% group_by(cluster) %>% top_n(n=20, wt=abs(avg_log2FC))
mg5_top20chk=T
} else {mg5_top20chk=F}

```

### **Microglia_1**

```{r echo=FALSE, message=FALSE,warning=FALSE, fig.width=8,fig.height=8}
if(mg1_top20chk){
  
gg = DoHeatmap(mg1, features = mg1_top20$gene) +
  ggtitle("Top 20 Marker genes for Genotype_Treatment in Microglia_1")
print(gg)

DT::datatable(mg1.markers.Genotype_Treatment, extensions = "Buttons",
              filter="top",
              caption = paste0("Genotype_Treatment in Microglia_1"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))
} else print("no significant markers identified")

```

### **Microglia_2**

```{r echo=FALSE, message=FALSE,warning=FALSE, fig.width=8,fig.height=8}
if(mg2_top20chk){
gg = DoHeatmap(mg2, features = mg2_top20$gene)  +
  ggtitle("Top 20 Marker genes for Genotype_Treatment in Microglia_2")
print(gg)

DT::datatable(mg2.markers.Genotype_Treatment, extensions = "Buttons",
              filter="top",
              caption = paste0("Genotype_Treatment in Microglia_2"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))
} else print("no significant markers identified")
```

### **Microglia_3**

```{r echo=FALSE, message=FALSE,warning=FALSE, fig.width=8,fig.height=8}
if(mg3_top20chk){
gg = DoHeatmap(mg3, features = mg3_top20$gene)  +
  ggtitle("Top 20 Marker genes for Genotype_Treatment in Microglia_3")
print(gg)

DT::datatable(mg3.markers.Genotype_Treatment, extensions = "Buttons",
              filter="top",
              caption = paste0("Genotype_Treatment in Microglia_3"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))
} else print("no significant markers identified")
```

### **Microglia_4**

```{r echo=FALSE, message=FALSE,warning=FALSE, fig.width=8,fig.height=8}
if(mg4_top20chk){
gg =DoHeatmap(mg4, features = mg4_top20$gene)  +
  ggtitle("Top 20 Marker genes for Genotype_Treatment in Microglia_4")
print(gg)

 DT::datatable(mg4.markers.Genotype_Treatment, extensions = "Buttons",
              filter="top",
              caption = paste0("Genotype_Treatment in Microglia_4"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))
} else print("no significant markers identified")
```


### **Microglia_5**

```{r echo=FALSE, message=FALSE,warning=FALSE, fig.width=8,fig.height=8}
if(mg5_top20chk){
gg = DoHeatmap(mg5, features = mg5_top20$gene)  +
  ggtitle("Top 20 Marker genes for Genotype_Treatment in Microglia_5")
print(gg)

DT::datatable(mg5.markers.Genotype_Treatment, extensions = "Buttons",
              filter="top",
              caption = paste0("Genotype_Treatment in Microglia_5"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))
} else print("no significant markers identified")
```



```{r}
save.image("./output/integrated_analysis.RData")
```






