---
title: 'SS2 (for AG Hefendehl): Integration'
author: "Desirée Brösamle"
date: "Dec 2021"
output:
  html_document:
    toc: no
---

```{r ,message=FALSE, warning=FALSE, include=FALSE}
home=getwd()


ep_str_wrap <- function(string, width) {
    x <- gregexpr(' ', string)
    vapply(seq_along(x),
           FUN = function(i) {
               y <- x[[i]]
               n <- nchar(string[i])
               len <- (c(y,n) - c(0, y)) ## length + 1
               idx <- len > width
               j <- which(!idx)
               if (length(j) && max(j) == length(len)) {
                   j <- j[-length(j)]
               }
               if (length(j)) {
                   idx[j] <- len[j] + len[j+1] > width
               }
               idx <- idx[-length(idx)] ## length - 1
               start <- c(1, y[idx] + 1)
               end <- c(y[idx] - 1, n)
               words <- substring(string[i], start, end)
               paste0(words, collapse="\n")
           },
           FUN.VALUE = character(1)
    )
}

```


```{r message=FALSE, warning=FALSE, include=FALSE}
library(Seurat)
library(ggplot2)
library(SingleR)
library(EnhancedVolcano)
library(pheatmap)


knitr::opts_chunk$set(fig.height=8, fig.width=15)

load(paste0(home,"/data/ReloadAllData_Hefendehl_Stroke_Dec'21.RData"))


```


# **Preprocessing**

**QC metrics from AG Hefendehl (Stroke)**

```{r message=FALSE, warning=FALSE}

# Visualize QC metrics as a violin plot
Idents(plates_hefendehl) <- "Plate"
VlnPlot(plates_hefendehl, features = c("nFeature_RNA", "nCount_RNA", "percent.mito","percent.ribo"), 
        group.by = "Plate", pt.size = 1, ncol = 4, )

plot1 | plot2

```




**QC metrics from AG Neher (own additional WT control)**

```{r message=FALSE, warning=FALSE}
# plates_neher <- subset(plates_neher, subset = Genotype == "wt")
# Visualize QC metrics as a violin plot
Idents(plates_neher) <- "Plate"
VlnPlot(plates_neher, features = c("nFeature_RNA", "nCount_RNA", "percent.mito","percent.ribo"), 
        group.by = "Plate",pt.size =1, ncol = 4)

plot1a | plot2a

```



Combine both seurat objects and filter cells that have unique feature counts (gene number) over 7000 or less than 200 and cells with >5%  mitochondrial counts.

**QC metrics after filtering**

```{r message=FALSE, warning=FALSE}

plot3 | plot4

```

# **Perform integrated analysis**

```{r echo=FALSE, results='hide',message=FALSE,warning=FALSE}
Idents(samples.integrated) <- "Plate"

DimPlot(samples.integrated, reduction = "harmony", group.by = "Mouse_ID")
DimHeatmap(samples.integrated, dims = 1:2, cells = 200, balanced = TRUE)

```

PCA, tSNE and UMAP from the first 30 dimensions and with a resolution of 0.8

```{r message=FALSE, warning=FALSE}
# Visualization
Idents(samples.integrated) <- "seurat_clusters"

DimPlot(samples.integrated, reduction = "umap", split.by = "Treatment")

p1 + p1a +  p1b

DimPlot(samples.integrated, group.by = "seurat_clusters", pt.size =1, label = T) + NoLegend()
DimPlot(samples.integrated, reduction = "umap", split.by = "Plate", pt.size = 1)
DimPlot(samples.integrated, reduction = "umap", split.by = "Genotype_Treatment", pt.size =1)

```


*Cell cycle scoring*

First, a score is assigned to each cell (Tirosh et al. 2016), based on its expression of G2/M and S phase markers. These markers should be anticorrelated in their expression levels and cells expression neither are likely not cycling and in G1 phase.
Note: For downstream cell cylce regression the quantitative scores for G2/M and S phase are used, not the dicrete classification.

```{r cell cycle scoring, echo=FALSE, results='hide',message=FALSE,warning=FALSE}
 Idents(samples.integrated) <- "Phase"
# View cell cycle scores and phase assignment
head(samples.integrated[[]])
 
# visualize the distribution of cell cycle markers across
RidgePlot(samples.integrated, features = c("Pcna", "Top2a", "Mcm6", "Mki67"), ncol = 2)
DimPlot(samples.integrated)

```


**Cluster distribution**

```{r cluster distribution, message=FALSE, warning=FALSE}
 
DimPlot(object = samples.integrated, pt.size = 1,reduction = "umap", group.by="Age",label = F) +
  ggtitle("Cluster distribution according to Age") | DimPlot(object = samples.integrated, pt.size = 1,reduction = "umap", group.by="Genotype_Treatment",label = F) +
  ggtitle("Cluster distribution according to Genotype + Treatment")

DimPlot(object = samples.integrated, pt.size = 1,reduction = "umap", group.by="Plate",label = F) +
  ggtitle("Cluster distribution according to Plate") | DimPlot(object = samples.integrated, pt.size = 1,reduction = "umap", group.by = "seurat_clusters", label = T) +
  ggtitle("Clustering")

```


```{r cell distribution per cluster, message=FALSE, warning=FALSE}

Idents(samples.integrated) <- "seurat_clusters"

# use same colors for clusters as in plots
require(scales)
identities <- levels(samples.integrated$seurat_clusters) # Create vector with levels of object@ident
cluster_colors <- hue_pal()(length(identities)) # Create vector of default ggplot2 colors


# number of cells in each cluster 
cluster_nCell <- as.data.frame.matrix(table(samples.integrated$seurat_clusters,
                                            samples.integrated$Genotype_Age))
cluster_nCell["Total" ,] = colSums(cluster_nCell)
cluster_nCell


# % of cells in each cluster , grouped by genotype_Age
cluster_percent_Cell <- data.frame(round((prop.table(x = table(samples.integrated$seurat_clusters, 
                              samples.integrated$Genotype_Age), margin = 2)*100),2))
colnames(cluster_percent_Cell) <- c("Genotype_Age", "Celltype", "Frequency")
cluster_percent_Cell

# Grouped barchart of cell proportions
ggplot(cluster_percent_Cell, aes(fill=Genotype_Age, y=Frequency, x=Celltype)) + 
  geom_bar(position="dodge", stat="identity")+
  ggtitle("Cell distribution according to Genotype_Age [%]") +
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=1)) +
  xlab("")


# % of cells in each cluster , grouped by genotype_treatment
cluster_percent_Cell <- data.frame(round((prop.table(x = table(samples.integrated$seurat_clusters, 
                              samples.integrated$Genotype_Treatment), margin = 2)*100),2))
colnames(cluster_percent_Cell) <- c("Genotype_Treatment", "Celltype", "Frequency")
cluster_percent_Cell

# Grouped barchart of cell proportions
ggplot(cluster_percent_Cell, aes(fill=Genotype_Treatment, y=Frequency, x=Celltype)) + 
  geom_bar(position="dodge", stat="identity")+
  ggtitle("Cell distribution according to Genotype_Treatment [%]") +
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=1)) +
  xlab("")
```


# **Cell type identification**

## **Finding differentially expressed features (cluster biomarkers)**

```{r Marker genes, message=FALSE, warning=FALSE}

Idents(samples.integrated) <- samples.integrated$seurat_clusters

DoHeatmap(samples.integrated, features = top20$gene) + NoLegend() + ggtitle("Top20 cluster marker genes")

```


## **SingleR**
SingleR is an automatic annotation method for (scRNAseq) data (Aran et al. 2019). Given a reference dataset of samples (single-cell or bulk) with known labels, it labels new cells from a test dataset based on similarity to the reference set. Here we use the built-in references "Immgen" (830 microarray samples of sorted hematopoetic and immune cell populations) 	 and "Mouse RNA-Seq" (358 non-specific mouse RNA-seq samples). 

**Immgen reference**

Section Skipped as pred.immgen object is not available 

```{r SingleR: Immgen, echo=FALSE, results='hide',message=FALSE,warning=FALSE, eval=FALSE}

# Based on clusters

table(pred.immgen$first.labels) #main cell types
table(pred.immgen$labels) #fine tuned cell types
#table(pred.immgen$pruned.labels) #cell type after pruning

plotScoreHeatmap(pred.immgen, show.labels = TRUE,
                 annotation_col=data.frame(cluster=unique(samples.integrated@meta.data$seurat_clusters),
                                           row.names=rownames(pred.immgen)))

# Based on single cells
  
table(pred.immgen.sc$first.labels)
table(pred.immgen.sc$labels)

plotScoreHeatmap(pred.immgen.sc)

#Visualization
DimPlot(samples.integrated,group.by ="Immgen_cluster_labels" ,label = F) + NoLegend() | DimPlot(samples.integrated, reduction = "umap",group.by = "seurat_clusters", label = T) + NoLegend() 
DimPlot(samples.integrated,group.by ="Immgen_sc_labels", label = F ) + NoLegend() | DimPlot(samples.integrated, reduction = "umap",group.by = "seurat_clusters", label = T) + NoLegend() 


```

**MouseRNA-Seq reference**
Section Skipped as pred.mouseRNA object is not available 

```{r SingleR: MouseRNA-Seq, echo=FALSE, results='hide',message=FALSE,warning=FALSE, eval=FALSE}

# Based on clusters

table(pred.mouseRNA$first.labels) #main cell types
table(pred.mouseRNA$labels) #fine tuned cell types
#table(pred.mouseRNA$pruned.labels) #cell type after pruning

plotScoreHeatmap(pred.mouseRNA, show.labels = TRUE,
                 annotation_col=data.frame(cluster=unique(samples.integrated$seurat_clusters),
                                           row.names=rownames(pred.mouseRNA)))

# Based on single cells
  
table(pred.mouseRNA.sc$first.labels)
table(pred.mouseRNA.sc$labels)

plotScoreHeatmap(pred.mouseRNA.sc)

#Visualization

DimPlot(samples.integrated,group.by ="MouseRNASeq_cluster_labels" ,label = T) + NoLegend() | DimPlot(samples.integrated, reduction = "umap",group.by = "seurat_clusters", label=T)+ NoLegend()
DimPlot(samples.integrated,group.by ="MouseRNASeq_sc_labels", label=F) # + NoLegend() | DimPlot(samples.integrated, reduction = "umap",group.by = "seurat_clusters", label=T)+ NoLegend()

```


```{r cell distribution MouseRNASeq, message=FALSE, warning=FALSE}

# number of cells in each cluster 
cluster_nCell <- data.frame(table(samples.integrated$MouseRNASeq_sc_labels,samples.integrated$Genotype_Treatment))
colnames(cluster_nCell) <- c("Genotype_Treatment", "MouseRNASeq_sc_labels", "Number")

#cluster_nCell["Total" ,] = colSums(cluster_nCell)
# cluster_nCell

# % of cells in each cluster 
cluster_percent_Cell <- data.frame(round((prop.table(x = table(samples.integrated$MouseRNASeq_sc_labels, 
                              samples.integrated$Genotype_Treatment), margin = 2)*100),2))
colnames(cluster_percent_Cell) <- c("Genotype_Treatment", "MouseRNASeq_sc_labels", "Frequency")
# cluster_percent_Cell


# Grouped barchart of absolute cell numbers
ggplot(cluster_nCell, aes(fill=Genotype_Treatment, y=Number, x=MouseRNASeq_sc_labels)) + 
  geom_bar(position="dodge", stat="identity") +
  ggtitle("Cell distribution according to MouseRNASeq reference (absolut values)") +
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=0.5)) +
  xlab("")

# Grouped barchart of cell proportions
ggplot(cluster_percent_Cell, aes(fill=Genotype_Treatment, y=Frequency, x=MouseRNASeq_sc_labels)) + 
  geom_bar(position="dodge", stat="identity")+
  ggtitle("Cell distribution according to MouseRNASeq reference [%]") +
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=0.5)) +
  xlab("")

```

```{r cell distribution immgen, message=FALSE, warning=FALSE}

# number of cells in each cluster 
cluster_nCell <- data.frame(table(samples.integrated$Immgen_sc_labels,samples.integrated$Treatment))
colnames(cluster_nCell) <- c("Treatment", "Immgen_sc_labels", "Number")

#cluster_nCell["Total" ,] = colSums(cluster_nCell)
# cluster_nCell

# % of cells in each cluster 
cluster_percent_Cell <- data.frame(round((prop.table(x = table(samples.integrated$Immgen_sc_labels, 
                              samples.integrated$Treatment), margin = 2)*100),2))
colnames(cluster_percent_Cell) <- c("Treatment", "Immgen_sc_labels", "Frequency")
# cluster_percent_Cell


# Grouped barchart of absolute cell numbers
ggplot(cluster_nCell, aes(fill=Treatment, y=Number, x=Immgen_sc_labels)) + 
  geom_bar(position="dodge", stat="identity") +
  ggtitle("Cell distribution according to Immgen reference (absolut values)") +
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=0.5)) +
  xlab("")

# Grouped barchart of cell proportions
ggplot(cluster_percent_Cell, aes(fill=Treatment, y=Frequency, x=Immgen_sc_labels)) + 
  geom_bar(position="dodge", stat="identity")+
  ggtitle("Cell distribution according to Immgen reference [%]") +
  theme(axis.text.x = element_text(angle = 45, hjust=1, vjust=0.5)) +
  xlab("")

```
Renaming Clusters

```{r}
DimPlot(samples.integrated, reduction = "umap", label=T) + NoLegend()
```


# **Zoom into Microglia clusters**

A second Seurat cluster is generated, where all non-microglial immune cells are removed according to the mouseRNASeq_sc_labels.

```{r subclustering microglia, message=FALSE, warning=FALSE}

p3 + p3a

```


## **DAMs (Cell. 2017 Jun 15;169(7):1276-1290.e17. doi: 10.1016/j.cell.2017.05.018)**

**DAM plotting**

```{r Stage 2 DAMs, echo=FALSE, results='hide',message=FALSE,warning=FALSE}

p4 | p4a

```

```{r}
p5 | p5a
```


```{r echo=FALSE, message=FALSE,results='hide',warning=FALSE}
Idents(microglia) <- "seurat_clusters"

VlnPlot(microglia, features = DAM2_up_1) + ggtitle("Genes upregulated in Stage 2 DAMs")
VlnPlot(microglia, features = DAM2_up_2) + ggtitle("Genes upregulated in Stage 2 DAMs")
VlnPlot(microglia, features = DAM2_up_3) + ggtitle("Genes upregulated in Stage 2 DAMs")

VlnPlot(microglia, features = DAM2_down_1) + ggtitle("Genes downregulated in Stage 2 DAMs")
VlnPlot(microglia, features = DAM2_down_2) + ggtitle("Genes downregulated in Stage 2 DAMs")
VlnPlot(microglia, features = DAM2_down_3) + ggtitle("Genes downregulated in Stage 2 DAMs")

```


## **Renaming clusters**

```{r cluster renaming, message=FALSE,warning=FALSE}
Idents(microglia) <- "Celltype"
DimPlot(microglia, reduction = "umap", label = TRUE, pt.size = 1)+ NoLegend()

```


## **Signature enrichment**

### **DEGs and GSEA between microglia of different genotypes**

**WT vs App**

```{r general DEGs, echo=FALSE, message=FALSE,warning=FALSE, eval=FALSE}

Idents(microglia) <- "Genotype"

# WT vs App

EnhancedVolcano(markers_WtvsApp,
                lab = rownames(markers_WtvsApp),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-1.5, 1.5), ylim=c(0,15), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Wt vs APPPS1+",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")


GSEA_output_wtvsapp$GOdown_plot
GSEA_output_wtvsapp$GOup_plot

GSEA_output_wtvsapp$cannonicalPathwaysdown_plot
GSEA_output_wtvsapp$cannonicalPathwaysup_plot

GSEA_output_wtvsapp$Motifdown_plot
GSEA_output_wtvsapp$Motifup_plot

GSEA_output_wtvsapp$ImmSigdown_plot
GSEA_output_wtvsapp$ImmSigup_plot
```

**WT vs MX04+**

```{r echo=FALSE, message=FALSE,warning=FALSE , eval=FALSE}

EnhancedVolcano(markers_WtvsMe,
                lab = rownames(markers_WtvsMe),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-2, 2), ylim=c(0,30), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Wt vs MX04+",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")


GSEA_output_wtvsme$GOdown_plot
GSEA_output_wtvsme$GOup_plot

GSEA_output_wtvsme$KEGGdown_plot
GSEA_output_wtvsme$KEGGup_plot

GSEA_output_wtvsme$cannonicalPathwaysdown_plot
GSEA_output_wtvsme$cannonicalPathwaysup_plot

GSEA_output_wtvsme$Motifdown_plot
GSEA_output_wtvsme$Motifup_plot

GSEA_output_wtvsme$ImmSigdown_plot
GSEA_output_wtvsme$ImmSigup_plot
```

**MX04+ vs App**
```{r echo=FALSE, message=FALSE,warning=FALSE, eval=FALSE}

EnhancedVolcano(markers_MevsAPP,
                lab = rownames(markers_MevsAPP),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-2, 2), ylim=c(0,30), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "MX04+ vs APPPS1+",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")

GSEA_output_mevsapp$GOdown_plot
GSEA_output_mevsapp$GOup_plot

GSEA_output_mevsapp$KEGGdown_plot
GSEA_output_mevsapp$KEGGup_plot

GSEA_output_mevsapp$cannonicalPathwaysdown_plot
GSEA_output_mevsapp$cannonicalPathwaysup_plot

GSEA_output_mevsapp$Motifdown_plot
GSEA_output_mevsapp$Motifup_plot

GSEA_output_mevsapp$ImmSigdown_plot
GSEA_output_mevsapp$ImmSigup_plot
```



### **DEGs and GSEA of microglia of different genotype and treatment**

**WT_Ctrl vs WT_Stroke**

```{r echo=FALSE, message=FALSE,warning=FALSE, eval=FALSE}

Idents(microglia) <- "Genotype_Treatment"

EnhancedVolcano(markers_WtCtrlvsStroke,
                lab = rownames(markers_WtCtrlvsStroke),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-2, 2), ylim=c(0,20), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Wt_Ctrl vs Wt_Stroke",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")


GSEA_output_wtCtrlvsStroke$GOdown_plot
GSEA_output_wtCtrlvsStroke$GOup_plot

GSEA_output_wtCtrlvsStroke$KEGGdown_plot
GSEA_output_wtCtrlvsStroke$KEGGup_plot

GSEA_output_wtCtrlvsStroke$cannonicalPathwaysdown_plot
GSEA_output_wtCtrlvsStroke$cannonicalPathwaysup_plot

GSEA_output_wtCtrlvsStroke$Motifdown_plot
GSEA_output_wtCtrlvsStroke$Motifup_plot

GSEA_output_wtCtrlvsStroke$ImmSigdown_plot
GSEA_output_wtCtrlvsStroke$ImmSigup_plot
```


**APP_Ctrl vs APP_Stroke**

```{r echo=FALSE, message=FALSE,warning=FALSE, eval=FALSE}

EnhancedVolcano(markers_AppCtrlvsStroke,
                lab = rownames(markers_AppCtrlvsStroke),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-1.5, 1.5), ylim=c(0,17), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "APPPS1+_Ctrl vs APPPS1+_Stroke",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")


GSEA_output_AppCtrlvsStroke$GOdown_plot
GSEA_output_AppCtrlvsStroke$GOup_plot

GSEA_output_AppCtrlvsStroke$KEGGdown_plot
GSEA_output_AppCtrlvsStroke$KEGGup_plot

GSEA_output_AppCtrlvsStroke$cannonicalPathwaysdown_plot
GSEA_output_AppCtrlvsStroke$cannonicalPathwaysup_plot

GSEA_output_AppCtrlvsStroke$Motifdown_plot
GSEA_output_AppCtrlvsStroke$Motifup_plot

GSEA_output_AppCtrlvsStroke$ImmSigdown_plot
GSEA_output_AppCtrlvsStroke$ImmSigup_plot
```


**WT_Ctrl vs APP_Ctrl**

```{r echo=FALSE, message=FALSE,warning=FALSE, eval=FALSE}

EnhancedVolcano(markers_Ctrl,
                lab = rownames(markers_Ctrl),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-2.5, 2.5), ylim=c(0,40), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Wt_Ctrl vs APPPS1+_Ctrl",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")


GSEA_output_ctrl$GOdown_plot
GSEA_output_ctrl$GOup_plot

GSEA_output_ctrl$KEGGdown_plot
GSEA_output_ctrl$KEGGup_plot

GSEA_output_ctrl$cannonicalPathwaysdown_plot
GSEA_output_ctrl$cannonicalPathwaysup_plot

GSEA_output_ctrl$Motifdown_plot
GSEA_output_ctrl$Motifup_plot

GSEA_output_ctrl$ImmSigdown_plot
GSEA_output_ctrl$ImmSigup_plot
```


**WT_Stroke vs APP_Stroke**


```{r echo=FALSE, message=FALSE,warning=FALSE, eval=FALSE} 

EnhancedVolcano(markers_Stroke,
                lab = rownames(markers_Stroke),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-2, 2), ylim=c(0,15), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Wt_Stroke vs APPPS1+_Stroke",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")


GSEA_output_Stroke$GOdown_plot
GSEA_output_Stroke$GOup_plot

GSEA_output_Stroke$KEGGdown_plot
GSEA_output_Stroke$KEGGup_plot

GSEA_output_Stroke$cannonicalPathwaysdown_plot
GSEA_output_Stroke$cannonicalPathwaysup_plot

GSEA_output_Stroke$Motifdown_plot
GSEA_output_Stroke$Motifup_plot

GSEA_output_Stroke$ImmSigdown_plot
GSEA_output_Stroke$ImmSigup_plot
```


### **DEGs and GSEA of each cluster**

**Microglia_0**

Only works if ReloadAllData_Harmony_Hefendehl_Stroke.RData is loaded.
Otherwise, an error message will occur.

```{r DEGs GSEA_output_ctrl cluster, echo=FALSE, message=FALSE,warning=FALSE, eval=FALSE}

EnhancedVolcano(zero,
                lab = rownames(zero),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-3, 3), ylim=c(0,55), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Microglia_0",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")

GSEA_output_zero$GOdown_plot
GSEA_output_zero$GOup_plot

GSEA_output_zero$KEGGdown_plot
GSEA_output_zero$KEGGup_plot

GSEA_output_zero$cannonicalPathwaysdown_plot
GSEA_output_zero$cannonicalPathwaysup_plot

GSEA_output_zero$Motifdown_plot
GSEA_output_zero$Motifup_plot

GSEA_output_zero$ImmSigdown_plot
GSEA_output_zero$ImmSigup_plot
```


**Microglia_1**

```{r DEGs GSEA_output_ctrl cluster 2, echo=FALSE, message=FALSE,warning=FALSE, eval=FALSE}

Idents(microglia) <- "Celltype"

EnhancedVolcano(one,
                lab = rownames(one),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-3, 3), ylim=c(0,75), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Microglia_1",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")

GSEA_output_one$GOdown_plot
GSEA_output_one$GOup_plot

GSEA_output_one$KEGGdown_plot
GSEA_output_one$KEGGup_plot

GSEA_output_one$cannonicalPathwaysdown_plot
GSEA_output_one$cannonicalPathwaysup_plot

GSEA_output_one$Motifdown_plot
GSEA_output_one$Motifup_plot

GSEA_output_one$ImmSigdown_plot
GSEA_output_one$ImmSigup_plot
```


**Microglia_2**

```{r echo=FALSE, message=FALSE,warning=FALSE, eval=FALSE}

EnhancedVolcano(two,
                lab = rownames(two),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-1.5, 1.5), ylim=c(0,20), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Microglia_2",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")

GSEA_output_two$GOdown_plot
GSEA_output_two$GOup_plot

GSEA_output_two$KEGGdown_plot
GSEA_output_two$KEGGup_plot

GSEA_output_two$cannonicalPathwaysdown_plot
GSEA_output_two$cannonicalPathwaysup_plot

GSEA_output_two$Motifdown_plot
GSEA_output_two$Motifup_plot

GSEA_output_two$ImmSigdown_plot
GSEA_output_two$ImmSigup_plot
```


**Microglia_4**

```{r echo=FALSE, message=FALSE,warning=FALSE, eval=FALSE}

EnhancedVolcano(four,
                lab = rownames(four),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-3, 3), ylim=c(0,15), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Microglia_4",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")

GSEA_output_four$GOdown_plot
GSEA_output_four$GOup_plot

GSEA_output_four$KEGGdown_plot
GSEA_output_four$KEGGup_plot

GSEA_output_four$cannonicalPathwaysdown_plot
GSEA_output_four$cannonicalPathwaysup_plot

GSEA_output_four$Motifdown_plot
GSEA_output_four$Motifup_plot

GSEA_output_four$ImmSigdown_plot
GSEA_output_four$ImmSigup_plot
```

**DAM**
 
```{r echo=FALSE, message=FALSE,warning=FALSE, eval=FALSE}

EnhancedVolcano(dam,
                lab = rownames(dam),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-3, 3), ylim=c(0,60), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "DAM",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")

GSEA_output_dam$GOdown_plot
GSEA_output_dam$GOup_plot

GSEA_output_dam$KEGGdown_plot
GSEA_output_dam$KEGGup_plot

GSEA_output_dam$cannonicalPathwaysdown_plot
GSEA_output_dam$cannonicalPathwaysup_plot

GSEA_output_dam$Motifdown_plot
GSEA_output_dam$Motifup_plot

GSEA_output_dam$ImmSigdown_plot
GSEA_output_dam$ImmSigup_plot
```




**Microglia_5**

Only works if ReloadAllData_Harmony_Hefendehl_Stroke.RData is loaded.
Otherwise, an error message will occur.

```{r echo=FALSE, message=FALSE,warning=FALSE, eval=FALSE}

EnhancedVolcano(five,
                lab = rownames(five),
                x = 'avg_log2FC', 
                y = 'p_val', xlim = c(-4.5, 4.5), ylim=c(0,45), pCutoff = 0.01,
                FCcutoff = log(1.5,2),
                xlab = bquote(~Log[2]~ 'fold change'),ylab = bquote(~-Log[10]~italic(P)),
                title = "DEGs", subtitle = "Microglia_5",
                caption = "p-value cutoff  0.01 \n FC cutoff  1.5")

GSEA_output_five$GOdown_plot
GSEA_output_five$GOup_plot

GSEA_output_five$KEGGdown_plot
GSEA_output_five$KEGGup_plot

GSEA_output_five$cannonicalPathwaysdown_plot
GSEA_output_five$cannonicalPathwaysup_plot

GSEA_output_five$Motifdown_plot
GSEA_output_five$Motifup_plot

GSEA_output_five$ImmSigdown_plot
GSEA_output_five$ImmSigup_plot
```

## Genotype + Treatment comparison in each cluster

**Microglia_0**

```{r echo=FALSE, message=FALSE,warning=FALSE}
Idents(mg0) <- "Genotype_Treatment"

DoHeatmap(mg0, features = mg0_top20$gene) + NoLegend() +
  ggtitle("Top 20 Marker genes for Genotype_Treatment in Microglia_0")

```

**Microglia_1**

```{r echo=FALSE, message=FALSE,warning=FALSE}
Idents(mg1) <- "Genotype_Treatment"

DoHeatmap(mg1, features = mg1_top20$gene) + NoLegend() +
  ggtitle("Top 20 Marker genes for Genotype_Treatment in Microglia_1")

```

**Microglia_2**

```{r echo=FALSE, message=FALSE,warning=FALSE}
Idents(mg2) <- "Genotype_Treatment"

DoHeatmap(mg2, features = mg2_top20$gene) + NoLegend() +
  ggtitle("Top 20 Marker genes for Genotype_Treatment in Microglia_2")

```

**DAM?**

```{r echo=FALSE, message=FALSE,warning=FALSE}
Idents(dams) <- "Genotype_Treatment"

DoHeatmap(dams, features = dams_top20$gene) + NoLegend() +
  ggtitle("Top 20 Marker genes for Genotype_Treatment in DAMs?")

```

**Microglia_4**

```{r echo=FALSE, message=FALSE,warning=FALSE}
Idents(mg4) <- "Genotype_Treatment"

DoHeatmap(mg4, features = mg4_top20$gene) + NoLegend() +
  ggtitle("Top 20 Marker genes for Genotype_Treatment in Microglia_4")

```


**Microglia_5**

```{r echo=FALSE, message=FALSE,warning=FALSE}
Idents(mg5) <- "Genotype_Treatment"

DoHeatmap(mg5, features = mg5_top20$gene) + NoLegend() +
  ggtitle("Top 20 Marker genes for Genotype_Treatment in Microglia_5")

```









