---
title: "Analyses for Publication"
author: "AGC"
date: "2023-09-27"
output: html_document
---

```{r setup, include=FALSE}

## don't update matrixStats
knitr::opts_chunk$set(echo = TRUE)

library(limma)
library(DESeq2)
library(tidyverse)
library(scCustomize)
library(Seurat)
library(harmony)
library(SingleR)
library(data.table)
library(compareGroups)
library(brms)
library(xlsx)
library(viridis)
library(DT)
library(slingshot)
library(SingleCellExperiment)
library(ggpubr)
library(clustree)
library(RCurl)
library(AnnotationHub)
library(WGCNA)
library(plotly)
library(org.Mm.eg.db)

```

```{r, include=FALSE}
samples.integrated = readRDS("/files/scSeq_Hefendehl/data/20230203/seu-SCT-harmony_02-02-23.rds")

samples.integrated$Genotype_Treatment = factor(samples.integrated$Genotype_Treatment, levels = c("WT_Ctrl", "WT_Stroke", "APPPS1_Ctrl", "APPPS1_Stroke"))
samples.integrated.orig <- samples.integrated



samples.integrated.orig@meta.data %>% dplyr::select(Mouse_ID, Genotype, Genotype_Treatment, Sample) %>% dplyr::distinct() -> Samplelist

openxlsx::write.xlsx(Samplelist, "/files/scSeq_Hefendehl/data/Samplelist.xlsx")


samples.integrated.orig@meta.data %>% dplyr::select(Mouse_ID, Genotype, Genotype_Treatment, Plate) %>% dplyr::distinct() -> Samplelist2




source ("/files/scSeq_Hefendehl/code/custom_functions.R")

cc_file <- getURL("https://raw.githubusercontent.com/hbc/tinyatlas/master/cell_cycle/Mus_musculus.csv") 
cell_cycle_genes <- read.csv(text = cc_file)

ah <- AnnotationHub::AnnotationHub()

# Access the Ensembl database for organism
ahDb <- query(ah, 
              pattern = c("Mus musculus", "EnsDb"), 
              ignore.case = TRUE)

# Acquire the latest annotation files
id <- ahDb %>%
  mcols() %>%
  rownames() %>%
  tail(n = 1)

# Download the appropriate Ensembldb database
edb <- ah[[id]]

# Extract gene-level information from database
annotations <- genes(edb, 
                     return.type = "data.frame")

# Select annotations of interest
annotations <- annotations %>%
  dplyr::select(gene_id, gene_name, seq_name, gene_biotype, description)
```



## Reclustering of dataset from Desiree

a)  Exclude sex-related genes -- maybe this will already resolve our issues concerning the clustering in terms of biological relevance.
b)  In case the new clusters are not matching the biological relevance, reclustering based on lists we provide for homeostatic and reactive microglial cells

if you want to save a plot use ggsave(filename = "./output/PNametheplot.svg", plotobject))

### Define Gene lists

```{r}

DAM_up <- c("Itgax","Cst7","Ccl4","Clec7a","Lpl","Siglec1","Spp1","Axl","Apoe","Trem2","Csf1","Lyz2","H2-D1","Tyrobp","Cd74","Gpnmb","B2m","Cd9","Ctsb","Ctsd") # genes up regulated in DAMs versus homestatics

# additions
DAM1_up <- c("Tyrobp", "Ctsb", "Ctsd", "Apoe", "B2m", "Fth1", "Lyz2") # genes up regulated in DAM1 vs Homeostasis
DAM1_down <- c("Cx3cr1", "P2ry12", "Tmem119")# genes down regulated in DAM1 vs homeostasis
DAM2_up <- c("Trem2", "Axl", "Cst7", "Ctsl", "Lpl", "Cd9", "Csf1", "Ccl6", "Itgax", "Clec7a", "Lilrb4", "Timp2")

DAM1vsDAM2 = c("Trem2", "Cst7", "Spp1", "Lpl", "Itgax", "Ctsl", "Ctsz", "Cd68", "Axl", "Cd9", "Ccl6", "Csf1")

homeostat.genes=c("P2ry12","P2ry13","Tmem119","Cx3cr1","Selplg","Cd33")


INF_resp_microglia <- c("Ifit2", "Ifit3", "Ifitm3", "Irf7", "Oasl2")
Cycling_microglia <- c("Top2a", "Mcm2", "Tubb5", "Mki67",  "Cdk1")
Act_resp_microglia <- c("Cd74", "H2-Ab1", "H2-Aa", "Ctsb",  "Ctsd")
Axon_tract_microglia <- c("Spp1", "Gpnmb", "Igf1", "Lgals3", "Fabp5", "Lpl", "Lgals1", "Ctsl", "Anxa5")



cluster.genes=unique(c(DAM_up, DAM1vsDAM2, homeostat.genes))
```

### Spatial cell types

```{r}
ngenes = 50
distal_dat = openxlsx::read.xlsx("/files/scSeq_Hefendehl/data/Copy of DEG_control_vs_stroke_vs_distal_original.xlsx", sheet = 1)
distal_dat <- distal_dat[order(distal_dat$score, decreasing = T)[1:ngenes],]
distal_genes <- distal_dat$X1

control_dat = openxlsx::read.xlsx("/files/scSeq_Hefendehl/data/Copy of DEG_control_vs_stroke_vs_distal_original.xlsx",sheet = 2)
control_dat <- control_dat[order(control_dat$score, decreasing = T)[1:ngenes],]
control_genes <- control_dat$X1

Stroke_dat = openxlsx::read.xlsx("/files/scSeq_Hefendehl/data/Copy of DEG_control_vs_stroke_vs_distal_original.xlsx",sheet = 3)
Stroke_dat <- Stroke_dat[order(Stroke_dat$score, decreasing = T)[1:ngenes],]
Stroke_genes <- Stroke_dat$X1
```

c)  Mapping cells of "spatially identified signature" onto the new plots

## Rechecking exclusion of cells

Some cells appear to have a low expression of Aif1 as can be seen in the purple shading. However, some of these cells with the low expression seem to be missing later on (see Clustering example). Please check if really only cells with an expression of 0 Aif1 have been excluded

### inspect unfiltered original RAW data

```{r, fig.width=8,fig.height=4}
# cluster all cells for visualization purpose prior to exclusion 

DefaultAssay(samples.integrated) = "RNA"
samples.integrated <- ScaleData(samples.integrated)
samples.integrated <- FindVariableFeatures(samples.integrated)
samples.integrated<- RunHarmony(samples.integrated, group.by.vars="Plate", npcs = 50)
samples.integrated<- RunPCA(samples.integrated, npcs = 50)

ElbowPlot(samples.integrated, ndims = 50)
samples.integrated <- JackStraw(samples.integrated, assay = "RNA")
samples.integrated <- ScoreJackStraw(samples.integrated, dims = 1:20)
JackStrawPlot(samples.integrated, dims = 1:20)
samples.integrated<- RunUMAP(samples.integrated, reduction = "harmony", dims = 1:20)

a = DimPlot(samples.integrated, reduction = "umap")
b = FeaturePlot_scCustom(samples.integrated, reduction = "umap", features = "Aif1", colors_use = viridis_dark_high)
b2 = FeaturePlot_scCustom(samples.integrated, reduction = "umap", features = "Ptprc", colors_use = viridis_dark_high)
c = RidgePlot(samples.integrated, features = "Aif1")
c2 = RidgePlot(samples.integrated, features = "Ptprc")

d = DimPlot(samples.integrated, reduction = "umap", split.by = "Genotype_Treatment", pt.size = 1)
e = FeaturePlot_scCustom(samples.integrated, reduction = "umap", features = "Aif1", split.by = "Genotype_Treatment", colors_use = viridis_dark_high)
e2 = FeaturePlot_scCustom(samples.integrated, reduction = "umap", features = "Ptprc", split.by = "Genotype_Treatment", colors_use = viridis_dark_high)
```

```{r, fig.width=8,fig.height=4}
a|b|b2
c|c2

getwd() # checks where you are run it directly in the console not here 
ggsave("/files/scSeq_Hefendehl/output/cluster_desiree_orginal.pdf", a)
```

```{r, fig.width=12,fig.height=4}
d
e
e2

```

## MIKE XIST CHECK

Some cells appear to have a low expression of Aif1 as can be seen in the purple shading. However, some of these cells with the low expression seem to be missing later on (see Clustering example). Please check if really only cells with an expression of 0 Aif1 have been excluded
We did in the end not fitler for Afip1 negative cells, see line 237 it is commented out 


### inspect unfiltered original RAW data FOR XIST - MIKE TEST

```{r, fig.width=8,fig.height=4}
# cluster all cells for visualization purpose prior to exclusion 

DefaultAssay(samples.integrated) = "RNA"
samples.integrated <- ScaleData(samples.integrated)
samples.integrated <- FindVariableFeatures(samples.integrated)
samples.integrated<- RunHarmony(samples.integrated, group.by.vars="Plate", npcs = 50)
samples.integrated<- RunPCA(samples.integrated, npcs = 50)

ElbowPlot(samples.integrated, ndims = 50)
samples.integrated <- JackStraw(samples.integrated, assay = "RNA")
samples.integrated <- ScoreJackStraw(samples.integrated, dims = 1:20)
JackStrawPlot(samples.integrated, dims = 1:20)
samples.integrated<- RunUMAP(samples.integrated, reduction = "harmony", dims = 1:20)

a = DimPlot(samples.integrated, reduction = "umap")
b = FeaturePlot_scCustom(samples.integrated, reduction = "umap", features = "Xist" , colors_use = viridis_dark_high)
b2 = FeaturePlot_scCustom(samples.integrated, reduction = "umap", features = "Xist", colors_use = viridis_dark_high)
c = RidgePlot(samples.integrated, features = "Xist")
c2 = RidgePlot(samples.integrated, features = "Xist")

d = DimPlot(samples.integrated, reduction = "umap", split.by = "Genotype_Treatment", pt.size = 1)
e = FeaturePlot_scCustom(samples.integrated, reduction = "umap", features = "Xist", split.by = "Genotype_Treatment", colors_use = viridis_dark_high)
e2 = FeaturePlot_scCustom(samples.integrated, reduction = "umap", features = "Xist", split.by = "Genotype_Treatment", colors_use = viridis_dark_high)
```

```{r, fig.width=8,fig.height=4}
a|b|b2
c|c2

getwd() # checks where you are run it directly in the console not here 
ggsave("/files/scSeq_Hefendehl/output/cluster_desiree_orginalXIST.pdf", a)
```

```{r, fig.width=12,fig.height=4}
d
e
e2

```


```{r, include=FALSE}

reads = samples.integrated@assays$RNA@counts %>% as.data.frame()
reads = t(reads) %>% as.data.frame()

metadata <- samples.integrated@meta.data
metadata$Aif1neg = NA
metadata$Aif1neg <- paste0("Aif1neg_",as.character(reads$Aif1 ==0))
metadata -> samples.integrated@meta.data


table(metadata$Genotype_Treatment, metadata$Aif1neg)

```

### find markers for the Aif1 neg

### exclude cells with low Aif1 expression and X or Y coded genes

#### Exclusion and data harmonization step

```{r, fig.width=4,fig.height=4}
samples.integrated.filt<- samples.integrated

# no filtering of Aif1 or Ptpcr uncomment section if you want to filter 
#samples.integrated.filt <- subset(x = samples.integrated.filt, subset = Aif1 > 0)
#samples.integrated.filt <- subset(x = samples.integrated.filt, subset = Ptprc > 0)


sex.genes <- rownames(samples.integrated.filt) %in%  annotations$gene_name[grepl("X|Y", annotations$seq_name)]
Gmgenes <- grepl("^Gm", rownames(samples.integrated.filt))
Rikgenes <- grepl(".*Rik$", rownames(samples.integrated.filt))

samples.integrated.filt<- samples.integrated.filt[!(sex.genes|Gmgenes|Rikgenes),]


DefaultAssay(samples.integrated.filt)<- "RNA"

ifnb.list <- SplitObject(samples.integrated, split.by = "Plate")
ifnb.list <- lapply(X = ifnb.list, FUN = function(x) {
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

features <- SelectIntegrationFeatures(object.list = ifnb.list)

feat.anchors <- FindIntegrationAnchors(object.list = ifnb.list, anchor.features = features)
samples.integrated <- IntegrateData(anchorset = feat.anchors)

samples.integrated.filt <- SCTransform(samples.integrated.filt, ncells = 3000)
samples.integrated.filt <- ScaleData(object = samples.integrated.filt)
samples.integrated.filt <- FindVariableFeatures(samples.integrated.filt)
samples.integrated.filt <- RunPCA(samples.integrated.filt, npcs = 50)
samples.integrated.filt <- RunHarmony(samples.integrated.filt, group.by.vars="Plate")
p <- ElbowPlot(samples.integrated.filt, ndims = 50, reduction = "harmony")
p

ggsave(filename = "/files/scSeq_Hefendehl/output/Ellbowplot.svg", p)

```

#### Re-optimize clustering

large difference in cell distribution between the genotypes

```{r, fig.width=8,fig.height=8}

reslist= seq(0.2,1.2,0.2)

DefaultAssay(samples.integrated.filt)<- "SCT"
samples.integrated.filt.target <- RunUMAP(samples.integrated.filt, reduction = "pca",  features = cluster.genes) %>% 
  FindNeighbors(features=cluster.genes) %>% FindClusters(features=cluster.genes, resolution = reslist)

clustreeplot <- clustree(samples.integrated.filt.target, prefix = "SCT_snn_res.")
clustreeplot

Idents(samples.integrated.filt.target) = "SCT_snn_res.0.6"
a<- DimPlot(samples.integrated.filt.target)
b <- DimPlot(samples.integrated.filt.target, split.by = "Genotype_Treatment")
c<- FeaturePlot_scCustom(samples.integrated.filt.target,features = homeostat.genes, colors_use = viridis_dark_high)

a
b
c
ggsave(filename = "/files/scSeq_Hefendehl/output/Gene_list_based_Clustertree.svg", clustreeplot)
ggsave(filename = "/files/scSeq_Hefendehl/output/Gene_list_based_Clusters.svg", a)
ggsave(filename = "/files/scSeq_Hefendehl/output/Gene_list_based_Clusters_byCondition.svg", b)
ggsave(filename = "/files/scSeq_Hefendehl/output/Gene_list_based_Clusters_byHomeostatgenes.svg", c)
```

#### All genes based clustering

```{r}

reslist= seq(0.2,1.2,0.2)


DefaultAssay(samples.integrated.filt)<- "SCT"
samples.integrated.filt <- RunUMAP(samples.integrated.filt, reduction = "pca", dims=1:20) %>% 
  FindNeighbors() %>% FindClusters(resolution = reslist)

clustreeplot<-clustree(samples.integrated.filt, prefix = "SCT_snn_res.")

Idents(samples.integrated.filt) = "SCT_snn_res.0.6"
a<-DimPlot(samples.integrated.filt)
b<-DimPlot(samples.integrated.filt, split.by = "Genotype_Treatment")
c<-FeaturePlot_scCustom(samples.integrated.filt,features = homeostat.genes, colors_use = viridis_dark_high)

clustreeplot
a
b
c

ggsave(filename = "/files/scSeq_Hefendehl/output/All_genes_Clustertree.svg", clustreeplot)
ggsave(filename = "/files/scSeq_Hefendehl/output/All_genes_Clusters.svg", a)
ggsave(filename = "/files/scSeq_Hefendehl/output/All_genes_Clusters_byCondition.svg", b)
ggsave(filename = "/files/scSeq_Hefendehl/output/All_genes_Clusters_byHomeostatgenes.svg", c)
```

```{r}
samples.integrated.filt <- RunUMAP(samples.integrated.filt, reduction = "pca", dims=1:20) %>% 
  FindNeighbors() %>% FindClusters(resolution = 0.6)

a<-DimPlot(samples.integrated.filt, group.by = "seurat_clusters", pt.size =1, label = T) + NoLegend()
b<-DimPlot(samples.integrated.filt, reduction = "umap", split.by = "Plate", pt.size = 1)
c<-DimPlot(samples.integrated.filt, reduction = "umap", split.by = "Genotype_Treatment", pt.size =1)

a
b
c


ggsave(filename = "/files/scSeq_Hefendehl/output/Opt_All_genes_all_cells_Clusters.svg", a)
ggsave(filename = "/files/scSeq_Hefendehl/output/Opt_All_genes_all_cells_Clusters_byPlate.svg", b)
ggsave(filename = "/files/scSeq_Hefendehl/output/Opt_All_genes_all_cells_Clusters_byCondition.svg", c)

```

#### Cell cycle scoring

```{r}
cell_cycle_markers <- dplyr::left_join(cell_cycle_genes, annotations, by = c("geneID" = "gene_id"))

# Acquire the S phase genes
s_genes <- cell_cycle_markers %>%
  dplyr::filter(phase == "S") %>%
  pull("gene_name")

# Acquire the G2M phase genes        
g2m_genes <- cell_cycle_markers %>%
  dplyr::filter(phase == "G2/M") %>%
  pull("gene_name")

samples.integrated.filt<-CellCycleScoring(samples.integrated.filt, s.features = s_genes, g2m.features = g2m_genes)

a<-DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Phase", pt.size =1)
b<-DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Phase", split.by = "Plate", pt.size =1)
c<-DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Phase", split.by = "Genotype_Treatment", pt.size =1)

a
b
c

ggsave(filename = "/files/scSeq_Hefendehl/output/G2M_Opt_All_genes_all_cells_Clusters.svg", a)
ggsave(filename = "/files/scSeq_Hefendehl/output/G2M_Opt_All_genes_all_cells_Clusters_byPlate.svg", b)
ggsave(filename = "/files/scSeq_Hefendehl/output/G2M_Opt_All_genes_all_cells_Clusters_byCondition.svg", c)

```

#### map cell types

```{r}
ref.sce.mm <- celldex::MouseRNAseqData()
sce.filt<- as.SingleCellExperiment(samples.integrated.filt, assay = "SCT")

pred.mouseRNA <- SingleR(test =sce.filt, ref = ref.sce.mm, assay.type.test=1,
                         labels = ref.sce.mm$label.fine)

table(pred.mouseRNA$labels) #fine tuned cell types

#table(pred.mouseRNA$pruned.labels) #cell type after pruning

samples.integrated.filt$Immgen_sc_labels_agc <- NA
samples.integrated.filt$Immgen_sc_labels_agc <- pred.mouseRNA$labels


plotScoreHeatmap(pred.mouseRNA, show.labels = TRUE,
                 annotation_col=data.frame(cluster=samples.integrated.filt$seurat_clusters,
                                           
                                           row.names=rownames(pred.mouseRNA)))

Idents(samples.integrated.filt)<- "seurat_clusters"

a<- DimPlot(samples.integrated.filt, reduction="umap", group.by="Immgen_sc_labels_agc", 
            label = F)
b<- DimPlot(samples.integrated.filt, reduction="umap", group.by="Immgen_sc_labels_agc", 
            label = F, split.by = "Plate")
c<- DimPlot(samples.integrated.filt, reduction="umap", group.by="Immgen_sc_labels_agc", split.by = "Genotype_Treatment",
            label = F)

a
b
c



ggsave(filename = "/files/scSeq_Hefendehl/output/Celltypes_Opt_All_genes_all_cells_Clusters.svg", a)
ggsave(filename = "/files/scSeq_Hefendehl/output/Celltypes_Opt_All_genes_all_cells_Clusters_byPlate.svg", b)
ggsave(filename = "/files/scSeq_Hefendehl/output/Celltypes_Opt_All_genes_all_cells_Clusters_byCondition.svg", c)
```

### filter and recluster only Microglia

```{r}

samples.integrated.filt = samples.integrated.filt[,samples.integrated.filt$Immgen_sc_labels_agc=="Microglia"]

samples.integrated.filt <- RunUMAP(samples.integrated.filt, reduction = "pca", dims=1:20) %>% 
  FindNeighbors() %>% FindClusters(resolution = 0.6)




n = length(unique(samples.integrated.filt$seurat_clusters))
samples.integrated.filt$labels<- factor(Idents(samples.integrated.filt), 
                                        levels = c(0:(n-1)), labels=paste0("Microglia", 0:(n-1))
)

Idents(samples.integrated.filt) = "labels"
a <- DimPlot(samples.integrated.filt, reduction="umap", label = T)
b <- DimPlot(samples.integrated.filt, reduction="umap", split.by = "Plate", label = F)
c <- DimPlot(samples.integrated.filt, reduction="umap", split.by = "Genotype_Treatment", label = F)
d <- DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Genotype_Treatment")
e <- DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Gender")
g <- DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Aif1neg")
h <- DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Immgen_sc_labels_agc")
i <- DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Methoxy")


a
b
c
d
e
g
h
i

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Opt_All_genes_Clusters.svg", a)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Opt_All_genes_Clusters_byPlate.svg", b)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Opt_All_genes_Clusters_byCondition.svg", c)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Opt_All_genes_Condition.svg", d)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Opt_All_genes_Gender.svg", e)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Opt_All_genes_Aif1neg.svg", g)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Opt_All_genes_Immgen_sc_labels_agc.svg", h)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Opt_All_genes_methoxy.svg", i)

```

```{r, fig.height=15, fig.width=12}
f <- FeaturePlot_scCustom(samples.integrated.filt, 
                          features = sort(cluster.genes),order =T, 
                          colors_use = viridis_dark_high)

f
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Opt_all_genes_MicrogliaGenes.svg", f)

```

```{r, fig.width=6, fig.height=6}
genelist <- c("Alpk1", "Mmp12", "Igf1", "S100a6", "Apoc4", "Dock10", "Nanog", "Olfr344")
f <- FeaturePlot_scCustom(samples.integrated.filt, 
                          features = sort(genelist),order =T, 
                          colors_use = viridis_dark_high)

f
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Opt_all_genes_AddGenelist.svg", f)



```



## Markers for Clusters 

```{r}
metadata<-samples.integrated.filt@meta.data

res = compareGroups(Genotype_Treatment~labels, data = metadata)

compareGroups::createTable(res)

compareGroups::createTable(res) %>% export2xls("/files/scSeq_Hefendehl/output/Cluster_distribution_genotype.xlsx") 


res = compareGroups(Genotype_Treatment~Mouse_ID, data = metadata)

compareGroups::createTable(res)

compareGroups::createTable(res) %>% export2xls("/files/scSeq_Hefendehl/output/Cellcount_distribution_MouseBygenotype.xlsx") 

res = compareGroups(labels~Mouse_ID, data = metadata, max.ylev = 15, max.xlev = 12)

compareGroups::createTable(res)

compareGroups::createTable(res) %>% export2xls("/files/scSeq_Hefendehl/output/Cellcount_distribution_MouseByCluster.xlsx") 

```


```{r, fig.height=18, fig.width=24}
f <- RidgePlot(samples.integrated.filt, features = sort(cluster.genes), layer = "data", combine = F)


f <- ggarrange(plotlist = f, common.legend = T, legend = "right")
f
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Opt_all_genes_MicrogliaGenesRidgePlot.svg", f)

```



### all clusters

```{r}
Idents(samples.integrated.filt) <- "labels"
# remove not annotated genes 
geneunivers<-rownames(samples.integrated.filt@assays$SCT@scale.data)
geneuniverse_annot <- geneunivers[!grepl("^AC[0-9]{4}|Rik|^BC[1-9]{4}|^CT[1-9]{4}", geneunivers)]

Markers=FindAllMarkers(samples.integrated.filt, 
                       logfc.threshold = 1, 
                       assay = "SCT", slot="scale.data",
                       features = geneuniverse_annot,
                       only.pos = T)

display_tab(Markers)
openxlsx::write.xlsx(Markers, "./output/Clusters_Markers_all.xlsx")

genelist = list()
for ( i in unique(Markers$cluster))  genelist[[i]] <- Markers$gene[Markers$cluster==i]

GOterm=getGOresults(genelist, genereference = geneunivers, organism = "mmusculus")
display_tab(GOterm$result)
openxlsx::write.xlsx(GOterm$result, "./output/Clusters_GOterms_Markers_all.xlsx")

```

```{r, fig.width=10, fig.height=12}
plotlist = list()

for (m in names(genelist)){
  #reads = rowSums(samples.integrated.filt@assays$RNA$counts)
  targets <- Markers[Markers$cluster==m, ] %>% arrange(desc(pct.1))
  targets <- targets$gene[1:10]
  
  p<-DoHeatmap(samples.integrated.filt, features = targets, 
               group.by = "labels", slot="scale.data",  
               label = F, combine=F) 
  p <- p[[1]] + scale_fill_viridis()
  ggsave(p, file=paste0("output/Top10GenesByExpr_",m,"_Heatmaps_byCluster.svg"))
  plotlist[[m]]<-p
}

p<- ggarrange(plotlist = plotlist, labels = names(genelist), common.legend = F)
p
ggsave(file="output/Top10Genes_Heatmaps_byCluster.svg", p)
```


### Add Microglia subtype scores

```{r, fig.width=6, fig.height=6}
samples.integrated.filt<- AddModuleScore(object = samples.integrated.filt, 
                                         assay="RNA", slot="counts",
                                         features = list(INF_resp_microglia), 
                                         name = "MG_score_INF_resp_microglia")

samples.integrated.filt<- AddModuleScore(object = samples.integrated.filt, 
                                         assay="RNA", slot="counts",
                                         features = list(Cycling_microglia), 
                                         name = "MG_score_Cycling_microglia")
samples.integrated.filt<- AddModuleScore(object = samples.integrated.filt, 
                                         assay="RNA", slot="counts",
                                         features = list(Act_resp_microglia), 
                                         name = "MG_score_Act_resp_microglia")
samples.integrated.filt<- AddModuleScore(object = samples.integrated.filt, 
                                         assay="RNA", slot="counts",
                                         features = list(Axon_tract_microglia), 
                                         name = "MG_score_Axon_tract_microglia")


featlist=grep("MG_score", colnames(samples.integrated.filt@meta.data), value=T)

p <- FeaturePlot_scCustom(samples.integrated.filt, 
                          na_cutoff = 0,
                          features = featlist,
                          colors_use = viridis_dark_high)

pcomb <- ggarrange(p,common.legend = T)
pcomb

ggsave("/files/scSeq_Hefendehl/output/MGscores_FeatMaps.svg", pcomb)

p2 <-RidgePlot(samples.integrated.filt,
               features = featlist, combine=F)

p2<-ggarrange(plotlist=p2, nrow = 2, ncol = 2)

ggsave("/files/scSeq_Hefendehl/output/MGscores_RidgePlots.svg", p2)

```



## Comparison of DEGs in DAMs between treatments/genotypes
5)  Subcluster overlaps

We will provide gene lists for different previously described microglia subclusters including homeostatic, DAM 1/DAM2 and IRM identification list. Please generate a Venn diagram or similar to see how big the overlap of our cells of interest (identified in the scDataset via the spatial transcriptomics list) is with the individual subclusters,

#### define DAM

Profiles based on genelist provided by Jan Hoffmann

NOT done: Alterantive DAM profiles based on Cell. 2017 Jun 15;169(7):1276-1290.e17. doi: 10.1016/j.cell.2017.05.018)

##### DAM1

```{r, fig.width=12, fig.height=4}
#load("./data/DAM_genelists.RData")
DefaultAssay(samples.integrated.filt) = "SCT"

samples.integrated.filt.target<- AddModuleScore(object = samples.integrated.filt.target, 
                                                features = list(DAM1_up), name = "DAM_score")

#FeaturePlot_scCustom(samples.integrated.filt.target, features = "DAM_score1")+scale_color_viridis_c()

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          features = list(DAM_up), name = "DAM_score")

p4 <- FeaturePlot_scCustom(seurat_object = samples.integrated.filt, features = "DAM_score1", colors_use = viridis_dark_high)
p4a <- DimPlot(object = samples.integrated.filt, reduction = "umap")

DAMcutoff = 1.5
phist = ggplot(data = samples.integrated.filt@meta.data, aes(x=DAM_score1))+
  geom_density()+geom_vline(xintercept = DAMcutoff)+
  theme_classic()

samples.integrated.filt$DAM_binary = as.factor(samples.integrated.filt$DAM_score1>=DAMcutoff)
pbin <- DimPlot(object = samples.integrated.filt, group.by =  "DAM_binary")

comb = phist | p4 | pbin| p4a
comb

prop.table(table(DAMs=samples.integrated.filt$DAM_binary, Clusters=samples.integrated.filt$labels), margin = 2)

dam_GT_p <-DimPlot(object = samples.integrated.filt, group.by =  "DAM_binary", split.by = "Genotype_Treatment")


ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAM1_calling.svg", comb)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAM1_bygenotype.svg", dam_GT_p)
```


```{r}
phist
p4
pbin
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAM1_thresholding.svg", phist)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAM1_scoring.svg", p4)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAM1_binary.svg", pbin)


```

```{r, fig.width=12, fig.height=4}
samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          features = list(DAM2_up), name = "DAM2_score")
p4 <- FeaturePlot_scCustom(seurat_object = samples.integrated.filt, features = "DAM2_score1", colors_use = viridis_dark_high)
p4a <- DimPlot(object = samples.integrated.filt, reduction = "umap")

DAM2cutoff = 1.0
phist = ggplot(data = samples.integrated.filt@meta.data, aes(x=DAM2_score1))+
  geom_density()+geom_vline(xintercept = DAM2cutoff)+
  theme_classic()

samples.integrated.filt$DAM2_binary = as.factor(samples.integrated.filt$DAM2_score1>=DAM2cutoff)
pbin <- DimPlot(object = samples.integrated.filt, group.by =  "DAM2_binary")
comb = phist | p4 | pbin| p4a


dam_GT_p <-DimPlot(object = samples.integrated.filt, group.by =  "DAM2_binary", split.by = "Genotype_Treatment")

comb
dam_GT_p

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAM2_calling.svg", comb)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAM2_bygenotype.svg", dam_GT_p)


prop.table(table(DAM2=samples.integrated.filt$DAM2_binary, Clusters=samples.integrated.filt$labels), margin = 2)
prop.table(table(DAM2=samples.integrated.filt$DAM2_binary, DAMs=samples.integrated.filt$DAM_binary), margin = 2)
```

```{r, fig.width=12}
a <- ggplot(samples.integrated.filt@meta.data, aes(x=DAM2_score1, y=DAM_score1))+
  geom_density_2d_filled()+geom_point(col="#FFFFFF55")+theme_classic()+ggtitle("all cells")

b<- ggplot(samples.integrated.filt@meta.data[samples.integrated.filt$DAM_binary==TRUE,], aes(x=DAM2_score1, y=DAM_score1))+geom_density_2d_filled()+geom_point(col="#FFFFFF55")+theme_classic()+ggtitle("DAMs only")

combo <- a|b
combo
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAMvsDAM2_contourplot.svg", combo)

```

proportion tables show percebtages of DAM cells per condition

```{r}
metadata=samples.integrated.filt@meta.data
restab = table(GT_Treat = metadata$Genotype_Treatment, DAMs=metadata$DAM_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot <- ggplot(as.data.frame(restab), 
                  aes(fill=DAMs, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()+ggtitle("DAM proportions")

proplot

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAMproportions_per_condition_all.svg", proplot)

library(chisq.posthoc.test)
chisq.posthoc.test(restab)

#To Do Pairwise 

```

### a) DAMs in APPPS1 vs. DAMs in APPPS1 + Stroke -- Venn diagram, or similar

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "APPPS1", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), DAMs=metadata$DAM_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=DAMs, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

proplot

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAMproportions_per_condition_APPPS1.svg", proplot)

```

### b) DAMs in WT+Stroke vs. DAMs in APPPS1 + Stroke -- Venn diagram, or similar

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "WT", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), DAMs=metadata$DAM_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=DAMs, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

proplot
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAMproportions_per_condition_WT.svg", proplot)


```

### c) Separation of DAMs into DAM1 and DAM2 (with a provided gene list), repetition of a) and b) to see if our cells on interest overlap more with one or the other profile

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$DAM_binary==TRUE,]
restab = table(GT_Treat = metadata$Genotype_Treatment, DAM2=metadata$DAM2_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=DAM2, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

chisq.posthoc.test(restab)

proplot
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAM2proportionsofDAM1_per_condition.svg", proplot)
```

```{r}

metadata=samples.integrated.filt@meta.data[(samples.integrated.filt$DAM_binary==TRUE) & (samples.integrated.filt$Genotype=="APPPS1"),]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), DAM2=metadata$DAM2_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<- ggplot(as.data.frame(restab), 
                 aes(fill=DAM2, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()


proplot
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAM2proportionsofDAM1_per_condition_APPPS1.svg", proplot)

```

```{r}

metadata=samples.integrated.filt@meta.data[(samples.integrated.filt$DAM_binary==TRUE) & (samples.integrated.filt$Genotype=="WT"),]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), DAM2=metadata$DAM2_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=DAM2, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()


proplot
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAM2proportionsofDAM1_per_condition_WT.svg", proplot)

```

## Homeostatic Cells

```{r fig.width=8, fig.height=15}
# list from Jan Hofmann 
a <- RidgePlot(samples.integrated.filt, features = homeostat.genes, ncol = 2)
a

b<- FeaturePlot_scCustom(samples.integrated.filt, features = homeostat.genes, colors_use = viridis_dark_high)

b

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_HomeostatGenes_Cluster_Ridgeplot.svg", a)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_HomeostatGenes_Featureplot.svg", a)
```

```{r, fig.width=12, fig.height=4}

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          features = list(homeostat.genes), name = "Homeo_score")

p4 <- FeaturePlot_scCustom(seurat_object = samples.integrated.filt, features = "Homeo_score1", colors_use = viridis_dark_high)
p4a <- DimPlot(object = samples.integrated.filt, reduction = "umap")

Homeocutff = 2
phist = ggplot(data = samples.integrated.filt@meta.data, aes(x=Homeo_score1))+
  geom_density()+geom_vline(xintercept = Homeocutff)+theme_classic()

samples.integrated.filt$Homeo_binary = as.factor(samples.integrated.filt$Homeo_score1>=Homeocutff)
pbin <- DimPlot(object = samples.integrated.filt, group.by =  "Homeo_binary")
comb = phist | p4 | pbin| p4a

Hom_GT_p <-DimPlot(object = samples.integrated.filt, group.by =  "Homeo_binary", split.by = "Genotype_Treatment")

comb
Hom_GT_p

prop.table(table(Homeo=samples.integrated.filt$Homeo_binary, Clusters=samples.integrated.filt$labels), margin = 2)


ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Homeo_calling.svg", comb)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Homeo_bygenotype.svg", Hom_GT_p)

```

```{r}
phist
p4
pbin
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Homeo_thresholding.svg", phist)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Homeo_scoring.svg", p4)
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Homeo_binary.svg", pbin)

```

#### Homeo vs DAM


```{r}

metadata = samples.integrated.filt@meta.data

metadata$Homeo_DAM <- factor(1*as.logical(metadata$Homeo_binary) + 2*as.logical(metadata$DAM_binary), levels = c(0,1,2,3), labels = c("Neither", "Homeo", "DAM", "Unknown"))

samples.integrated.filt@meta.data <- metadata


a = DimPlot(samples.integrated.filt, reduction="umap", group.by = "Homeo_DAM")
a
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Homeo_DAM.svg", a)
```


```{r}
restab = table(GT_Treat = metadata$Genotype_Treatment, Homeo=metadata$Homeo_DAM)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=Homeo, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

chisq.posthoc.test(restab)

proplot
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_HomeoDAMproportions_per_condition.svg", proplot)
```


```{r}
restab = table(Clust = metadata$labels, Homeo=metadata$Homeo_DAM)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=Homeo, x=Clust, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

chisq.posthoc.test(restab)

proplot
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_HomeoDAMproportions_per_ckuster.svg", proplot)
```


```{r}
restab = table(Clust = metadata$labels, Mex=metadata$Methoxy)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=Mex, x=Clust, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

chisq.posthoc.test(restab)

proplot
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_mex.svg", proplot)
```

```{r}
restab = table(Clust = metadata$labels, Celltype=metadata$MouseRNASeq_sc_labels)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=Celltype, x=Clust, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

chisq.posthoc.test(restab)

proplot
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_celltype_labels.svg", proplot)
```


```{r}
restab = table(Clust = metadata$labels, gender=metadata$Gender)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=gender, x=Clust, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

chisq.posthoc.test(restab)

proplot
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_HomeoDAMproportions_per_gender.svg", proplot)
```

```{r, fig.width=12}
a <- ggplot(samples.integrated.filt@meta.data, aes(x=Homeo_score1, y=DAM_score1))+
  geom_density_2d_filled()+geom_point(col="#FFFFFF55")+theme_classic()+ggtitle("all cells")

b<- ggplot(samples.integrated.filt@meta.data[samples.integrated.filt$DAM_binary==TRUE,], aes(x=Homeo_score1, y=DAM_score1))+geom_density_2d_filled()+geom_point(col="#FFFFFF55")+theme_classic()+ggtitle("DAMs only")

a|b

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_DAMvsHomeo.svg", a|b)

```

### d) Homeostatic microglia (with provided gene list) in APPPS1 vs. Homeostatic microglia in APPPS1 + Stroke -- Venn diagram, or similar

proportion tables show percentages of Homeostatic cells per condition

```{r}
metadata=samples.integrated.filt@meta.data
restab = table(GT_Treat = metadata$Genotype_Treatment, Homeo=metadata$Homeo_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=Homeo, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()


chisq.posthoc.test(restab)


proplot
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Homeoproportions_per_condition.svg", proplot)

```

### e) Homeostatic microglia in WT+Stroke vs. Homeostatic microglia in APPPS1 + Stroke -- Venn diagram, or similar

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype=="APPPS1",]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Homeo=metadata$Homeo_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<- ggplot(as.data.frame(restab), 
                 aes(fill=Homeo, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

proplot
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Homeoproportions_per_condition_APPPS1.svg", proplot)

```

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype=="WT",]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Homeo=metadata$Homeo_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=Homeo, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

proplot
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Homeoproportions_per_condition_WT.svg", proplot)

```

```{r}
### Spatial cell types 


Idents(samples.integrated.filt) = "Genotype_Treatment"
DoHeatmap(samples.integrated.filt, slot = "data", features = c(distal_genes, control_genes, Stroke_genes))+ggtitle("spatial genes all")
DoHeatmap(samples.integrated.filt, slot = "data",features = c(distal_genes))+ggtitle("spatial genes distal")
DoHeatmap(samples.integrated.filt, slot = "data",features = c(control_genes))+ggtitle("spatial genes control")
DoHeatmap(samples.integrated.filt, slot = "data",features = c(Stroke_genes))+ggtitle("spatial genes peri-ictus")
```

##### DISTAL

```{r, fig.width=12, fig.height=4}
#load("./data/DAM_genelists.RData")
DefaultAssay(samples.integrated.filt) = "SCT"

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          features = list(distal_genes), name = "Distal")


p4 <- FeaturePlot_scCustom(seurat_object = samples.integrated.filt, features = "Distal1", colors_use = viridis_dark_high)
p4a <- DimPlot(object = samples.integrated.filt, reduction = "umap")

Distalcutoff = 0.5
phist = ggplot(data = samples.integrated.filt@meta.data, aes(x=Distal1))+
  geom_density()+geom_vline(xintercept = Distalcutoff)+
  theme_classic()

samples.integrated.filt$Distal_binary = as.factor(samples.integrated.filt$Distal1>=Distalcutoff)
pbin <- DimPlot(object = samples.integrated.filt, group.by =  "Distal_binary")
phist | p4 | pbin| p4a

prop.table(table(Distals=samples.integrated.filt$Distal_binary, Clusters=samples.integrated.filt$labels), margin = 2)

DimPlot(object = samples.integrated.filt, group.by =  "Distal_binary", split.by = "Genotype_Treatment")

```

```{r, fig.width=12, fig.height=4}
#load("./data/DAM_genelists.RData")
DefaultAssay(samples.integrated.filt) = "SCT"

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          features = list(control_genes), name = "Control")


p4 <- FeaturePlot_scCustom(seurat_object = samples.integrated.filt, features = "Control1", colors_use = viridis_dark_high)
p4a <- DimPlot(object = samples.integrated.filt, reduction = "umap")

Controlcutoff = 0.25
phist = ggplot(data = samples.integrated.filt@meta.data, aes(x=Control1))+
  geom_density()+geom_vline(xintercept = Controlcutoff)+
  theme_classic()

samples.integrated.filt$Control_binary = as.factor(samples.integrated.filt$Control1>=Controlcutoff)
pbin <- DimPlot(object = samples.integrated.filt, group.by =  "Control_binary")
phist | p4 | pbin| p4a

prop.table(table(Controls=samples.integrated.filt$Control_binary, Clusters=samples.integrated.filt$labels), margin = 2)

DimPlot(object = samples.integrated.filt, group.by =  "Control_binary", split.by = "Genotype_Treatment")

```

```{r, fig.width=12, fig.height=4}
#load("./data/DAM_genelists.RData")
DefaultAssay(samples.integrated.filt) = "SCT"

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          features = list(Stroke_genes), name = "Peri_ictus")


p4 <- FeaturePlot_scCustom(seurat_object = samples.integrated.filt, features = "Peri_ictus1", colors_use = viridis_dark_high)
p4a <- DimPlot(object = samples.integrated.filt, reduction = "umap")

Strokecutoff = 2.5
phist = ggplot(data = samples.integrated.filt@meta.data, aes(x=Peri_ictus1))+
  geom_density()+geom_vline(xintercept = Strokecutoff)+
  theme_classic()+ggtitle("Peri_ictus")

samples.integrated.filt$Peri_ictus_binary = as.factor(samples.integrated.filt$Peri_ictus1>=Strokecutoff)
pbin <- DimPlot(object = samples.integrated.filt, group.by =  "Peri_ictus_binary")
phist | p4 | pbin| p4a

prop.table(table(Strokes=samples.integrated.filt$Peri_ictus_binary, Clusters=samples.integrated.filt$labels), margin = 2)

DimPlot(object = samples.integrated.filt, group.by =  "Peri_ictus_binary", split.by = "Genotype_Treatment")

```

```{r}
ggplot(samples.integrated.filt@meta.data, aes(x=Peri_ictus1, y=Control1))+
  geom_density_2d_filled()+geom_point(col="#FFFFFF55")+theme_classic()+ggtitle("all cells")

ggplot(samples.integrated.filt@meta.data[samples.integrated.filt$Peri_ictus_binary==TRUE,], aes(x=Peri_ictus1, y=Control1))+geom_density_2d_filled()+geom_point(col="#FFFFFF55")+theme_classic()+ggtitle("Peri_ictus only")
```

proportion tables show percebtages of DAM cells per condition

### Distals binary proportions

```{r}
metadata=samples.integrated.filt@meta.data
restab = table(GT_Treat = metadata$Genotype_Treatment, Distals=metadata$Distal_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Distals, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

library(chisq.posthoc.test)
chisq.posthoc.test(restab)

#To Do Pairwise 


```

### a) Distals in APPPS1 vs. DAMs in APPPS1 + Stroke -- Venn diagram, or similar

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "APPPS1", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Distals=metadata$Distal_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Distals, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()



```

### b) Distals in WT+Stroke vs. DAMs in APPPS1 + Stroke -- Venn diagram, or similar

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "WT", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Distals=metadata$Distal_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Distals, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

```

### Controls binary proportions

```{r}
metadata=samples.integrated.filt@meta.data
restab = table(GT_Treat = metadata$Genotype_Treatment, Controls=metadata$Control_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Controls, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

library(chisq.posthoc.test)
chisq.posthoc.test(restab)

#To Do Pairwise 


```

### a) Controls in APPPS1 vs. DAMs in APPPS1 + Stroke -- Venn diagram, or similar

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "APPPS1", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Controls=metadata$Control_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Controls, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()



```

### b) Controls in WT+Stroke vs. DAMs in APPPS1 + Stroke -- Venn diagram, or similar

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "WT", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Controls=metadata$Control_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Controls, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()
```

### Strokes binary proportions

```{r}
metadata=samples.integrated.filt@meta.data
restab = table(GT_Treat = metadata$Genotype_Treatment, Peri_Ictal=metadata$Peri_ictus_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Peri_Ictal, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

library(chisq.posthoc.test)
chisq.posthoc.test(restab)

#To Do Pairwise 


```

### a) Strokes in APPPS1 vs. DAMs in APPPS1 + Stroke -- Venn diagram, or similar

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "APPPS1", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Peri_Ictal=metadata$Peri_ictus_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Peri_Ictal, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

```

### b) Strokes in WT+Stroke vs. DAMs in APPPS1 + Stroke -- Venn diagram, or similar

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "WT", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Peri_Ictal=metadata$Peri_ictus_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Peri_Ictal, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity")+theme_classic()

```

### Alternative appraoch to identify cell types (assuming the cell types are mutaully exclusive)

```{r}
library(scSorter)

annodf = data.frame(Type = "Periictus", Marker = Stroke_genes)
annodf = rbind(annodf, data.frame(Type = "Control", Marker = control_genes))
annodf = rbind(annodf, data.frame(Type = "Distal", Marker = distal_genes))


spatialtypes = scSorter(samples.integrated.filt@assays$SCT@counts, annodf) #makes the calling 

metadata = samples.integrated.filt@meta.data 
metadata = metadata[,! colnames(metadata) %in% c("PredType")]

samples.integrated.filt@meta.data <- cbind(metadata, PredType = spatialtypes$Pred_Type)

DimPlot(samples.integrated.filt, group.by = "PredType", split.by = "Genotype_Treatment")

samples.integrated.filt.comorbid <- samples.integrated.filt[,metadata$Genotype_Treatment == "APPPS1_Stroke"]
DimPlot(samples.integrated.filt.comorbid, group.by = "PredType", )

crosstab = table(Condition = samples.integrated.filt$Genotype_Treatment, Spatial_Type = samples.integrated.filt$PredType)


ggplot(as.data.frame(crosstab), aes(fill=Spatial_Type, x=Condition, y=Freq)) + geom_bar(position="fill", stat="identity")+theme_classic()
crosstab
chisq.test(crosstab)

ggplot(as.data.frame(crosstab[1:2,]), aes(fill=Spatial_Type, x=Condition, y=Freq)) + geom_bar(position="fill", stat="identity")+theme_classic()
crosstab[1:2,]
chisq.test(crosstab[1:2,])

ggplot(as.data.frame(crosstab[3:4,]), aes(fill=Spatial_Type, x=Condition, y=Freq)) + geom_bar(position="fill", stat="identity")+theme_classic()
crosstab[3:4,]
chisq.test(crosstab[3:4,])

crosstab/rowSums(crosstab)
```

### DEG comparing conditions overall

##### WT Stroke vs no Stroke

avg_logFC: log fold-chage of the average expression between the two groups. Positive values indicate that the gene is more highly expressed in the first group

```{r, fig.width=6, fig.height=6}
genes_to_label <- openxlsx::read.xlsx("/files/scSeq_Hefendehl/data/Volcano pot labels on graph.xlsx", 
                                      sheet = "Wildtype control vs Wildtype st")
genes_to_label <- c(genes_to_label$Up, genes_to_label$Down)

Idents(samples.integrated.filt) <- "Genotype_Treatment"
DimPlot(samples.integrated.filt)
DEG_Wt_CtrlvsStroke <-FindMarkers(samples.integrated.filt, 
                                  ident.1 = "WT_Stroke", ident.2 = "WT_Ctrl", 
                                  test.use = "LR", logfc.threshold = 0)

genes_to_label[!genes_to_label %in% rownames(DEG_Wt_CtrlvsStroke)]

p = EnhancedVolcano::EnhancedVolcano(DEG_Wt_CtrlvsStroke, 
                                     x = "avg_log2FC", 
                                     y = "p_val_adj", 
                                     pCutoff=0.05,
                                     lab=rownames(DEG_Wt_CtrlvsStroke),
                                     selectLab = genes_to_label,
                                     title = "WT_Ctrl vs WT_Stroke")



display_tab(DEG_Wt_CtrlvsStroke)
p

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Volcano_WTCtrlvsWTStroke.svg", p)
```

##### APPPS1 Stroke vs APPPS1 no Stroke

avg_logFC: log fold-chage of the average expression between the two groups. Positive values indicate that the gene is more highly expressed in the first group

```{r fig.width=6, fig.height=6}
genes_to_label <- openxlsx::read.xlsx("/files/scSeq_Hefendehl/data/Volcano pot labels on graph.xlsx", 
                                      sheet = "APPPS1 control Vs APPPS1 stroke")
genes_to_label <- c(genes_to_label$Up, genes_to_label$Down)



DEG_APPPS1_CtrlvsStroke <-FindMarkers(samples.integrated.filt, 
                                      ident.1 = "APPPS1_Stroke", ident.2 = "APPPS1_Ctrl", 
                                      test.use = "LR", logfc.threshold = 0)
display_tab(DEG_APPPS1_CtrlvsStroke)

genes_to_label[!genes_to_label %in% rownames(DEG_APPPS1_CtrlvsStroke)]

p<- EnhancedVolcano::EnhancedVolcano(DEG_APPPS1_CtrlvsStroke, 
                                     x = "avg_log2FC", 
                                     y = "p_val_adj",
                                     lab=rownames(DEG_APPPS1_CtrlvsStroke), 
                                     selectLab = genes_to_label,
                                     pCutoff = 0.05,
                                     title = "APPPS1_Ctrl vs APPPS1_Stroke")


p

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Volcano_APPPS1CtrlvsAPPPS1Stroke.svg", p)

```


##### WT Stroke vs APPPS1Stroke

```{r fig.width=6, fig.height=6}


# plot here only DAM markers 

genes_to_label <- unique(c(DAM_up, DAM1_down, DAM1_up, DAM1vsDAM2, DAM2_up))

DEG_Stroke_APPPS1vsWT <-FindMarkers(samples.integrated.filt, 
                                    ident.1 = "APPPS1_Stroke", ident.2 = "WT_Stroke", 
                                    test.use = "LR", logfc.threshold = 0)
display_tab(DEG_Stroke_APPPS1vsWT)

genes_to_label[!genes_to_label %in% rownames(DEG_Stroke_APPPS1vsWT)]

p<- EnhancedVolcano::EnhancedVolcano(DEG_Stroke_APPPS1vsWT, 
                                     x = "avg_log2FC", 
                                     y = "p_val_adj",
                                     lab=rownames(DEG_Stroke_APPPS1vsWT), 
                                     selectLab = genes_to_label,
                                     pCutoff = 0.05,
                                     title = "WT_stroke vs APPPS1_Stroke")

p

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Volcano_WTstrokevsAPPPS1StrokeDAMgenes.svg", p)

```

##### WT ctr; vs APPPS1 ctrl

avg_logFC: log fold-chage of the average expression between the two groups. Positive values indicate that the gene is more highly expressed in the first group

```{r fig.width=6, fig.height=6}

genes_to_label <- openxlsx::read.xlsx("/files/scSeq_Hefendehl/data/Volcano pot labels on graph.xlsx", 
                                      sheet = "Wild type control Vs APPPS1 con")
genes_to_label <- c(genes_to_label$Up, genes_to_label$Down)

DEG_APPPS1_CtrlvsWT_Ctrl <-FindMarkers(samples.integrated.filt, 
                                       ident.1 = "APPPS1_Ctrl", ident.2 = "WT_Ctrl", 
                                       test.use = "LR", logfc.threshold = 0)
display_tab(DEG_APPPS1_CtrlvsWT_Ctrl)
genes_to_label[!genes_to_label %in% rownames(DEG_APPPS1_CtrlvsWT_Ctrl)]

p<- EnhancedVolcano::EnhancedVolcano(DEG_APPPS1_CtrlvsWT_Ctrl, 
                                     x = "avg_log2FC", 
                                     y = "p_val_adj", 
                                     lab=rownames(DEG_APPPS1_CtrlvsWT_Ctrl), 
                                     selectLab = genes_to_label,
                                     pCutoff = 0.05,
                                     title = "WT_Ctrl vs APPPS1_Ctrl")


p

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Volcano_WTCtrlvsAPPPS1ctrl.svg", p)

```




#### Scatterplot of DEGs of all four conditions, see <https://www.nature.com/articles/s41593-023-01355-y> ; see Fig 1e


```{r}
genelist_both= intersect(rownames(DEG_Wt_CtrlvsStroke), rownames(DEG_APPPS1_CtrlvsStroke))

plotdata=data.frame(log2FC_WTCtrl_vs_WTStroke=DEG_Wt_CtrlvsStroke[genelist_both,]$avg_log2FC, log2FC_APPPS1Ctrl_vs_APPPS1Stroke = DEG_APPPS1_CtrlvsStroke[genelist_both,]$avg_log2FC)
rownames(plotdata) = genelist_both

plotdata$label = rownames(plotdata)

# the next lines plot the top 20 genes of either analyses 
n=20
plotdata$label[! c(1:nrow(plotdata)) %in% c(
  order(abs(plotdata$log2FC_WTCtrl_vs_WTStroke), decreasing = T)[1:n], 
  order(abs(plotdata$log2FC_APPPS1Ctrl_vs_APPPS1Stroke), decreasing = T)[1:n])]=""


#if you execute the next two lines the labels you choose will be plotted
# to execute set FALSE to TRUE and change the list of gene names 
if(TRUE){
  genestoplot = c("S100a6", "Apoc4", "C4b", "Atp1a2", "Apoc1", "Fn1","Ntrk2")
  print(paste0(genestoplot[! genestoplot %in% rownames(plotdata)], " cannot be found"))
  plotdata$label[! rownames(plotdata) %in% genestoplot] = ""
}
```


```{r}
p <- ggplot(plotdata, aes(x=log2FC_WTCtrl_vs_WTStroke, y=log2FC_APPPS1Ctrl_vs_APPPS1Stroke, label=label))+geom_point(alpha=0.7)+
  geom_text(vjust=-0.7, color="black")+
  #geom_abline(slope=-1, intercept=0, lty=2)+geom_abline(slope=1, intercept=0, lty=2)+
  theme_classic()+ggtitle("log_FC")
p
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_logFCCtrlvsStroke_APPPS1vsWT_all.svg", p)

```


```{r}

# sig only

genelist_both= unique(c(rownames(DEG_Wt_CtrlvsStroke[DEG_Wt_CtrlvsStroke$p_val<0.05,]), rownames(DEG_APPPS1_CtrlvsStroke[DEG_APPPS1_CtrlvsStroke$p_val>0.05,])))

plotdata=data.frame(log2FC_WTCtrl_vs_WTStroke=DEG_Wt_CtrlvsStroke[genelist_both,]$avg_log2FC, log2FC_APPPS1Ctrl_vs_APPPS1Stroke = DEG_APPPS1_CtrlvsStroke[genelist_both,]$avg_log2FC)
rownames(plotdata) = genelist_both

plotdata$label = rownames(plotdata)



#if you execute the next lines the labels you choose will be plotted
# if set to TRUE, change the list of gene names and they will be plotted
# if set to FALSE the top n=15 of both will be plotted 

mode = "both" # alternative "selected" or "topN" or "both"
genestoplot = c("S100a6", "Apoc4", "C4b", "Atp1a2", "Apoc1", "Fn1","Ntrk2") # name here your gene of interest
topn = 15 # name the top number of genes to be plotted 


### don't change anything below the line here 

plotdata$label = rownames(plotdata)

if(mode=="selected"){
  
  print(paste0(genestoplot[! genestoplot %in% rownames(plotdata)], " cannot be found"))
  plotdata$label[which(! rownames(plotdata) %in% genestoplot)] = ""
} else if (mode=="topN"){
  n=topn
  genestoplotn15 <-c(1:nrow(plotdata)) %in% 
    c(
      order(abs(plotdata$log2FC_WTCtrl_vs_WTStroke), decreasing = T)[1:n], 
      order(abs(plotdata$log2FC_APPPS1Ctrl_vs_APPPS1Stroke), decreasing = T)[1:n])
  plotdata$label[!genestoplotn15]=""
} else if (mode =="both") {
  print(paste0(genestoplot[! genestoplot %in% rownames(plotdata)], " cannot be found"))
  genestoplotcmbn <-(c(1:nrow(plotdata)) %in% 
                       c(
                         order(abs(plotdata$log2FC_WTCtrl_vs_WTStroke), decreasing = T)[1:n], 
                         order(abs(plotdata$log2FC_APPPS1Ctrl_vs_APPPS1Stroke), decreasing = T)[1:n])) | (rownames(plotdata) %in% genestoplot)
  plotdata$label[which(! genestoplotcmbn)] = ""
}



p <- ggplot(plotdata, aes(x=log2FC_WTCtrl_vs_WTStroke, y=log2FC_APPPS1Ctrl_vs_APPPS1Stroke, label=label))+geom_point(alpha=0.7)+
  geom_text(vjust=-0.7, color="blue")+
  #geom_abline(slope=-1, intercept=0, lty=2)+geom_abline(slope=1, intercept=0, lty=2)+
  theme_classic()+ggtitle("log_FC")

p
ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_logFCCtrlvsStroke_APPPS1vsWT_sig_only.svg", p)




```


```{r fig.width=8, fig.height=8}
reads_data = samples.integrated.filt@assays$RNA@data # loaded data as used above 
#reads_data = samples.integrated.filt@assays$RNA@counts # Raw counts/reads  
#reads_data = samples.integrated.filt@assays$RNA@scale.data  #scaled/reads

barplot_genes = function(gene, split.by =T){
  require(ggbeeswarm)
  
  plotdata = data.frame(log2Reads = log2(reads_data[gene,]+1), 
                        Genotype_Treatment = samples.integrated.filt@meta.data$Genotype_Treatment, 
                        Cluster = samples.integrated.filt@meta.data$labels)
  
  if(split.by=="none" | split.by=="cluster") {
    p=ggplot(plotdata, aes(x=Genotype_Treatment, y=log2Reads, fill=Genotype_Treatment))+
      geom_violin()+
      geom_beeswarm(alpha=0.4, method = "hex")+ggtitle(gene)+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
    if(split.by=="cluster"){p=p+facet_wrap(~Cluster)}
  } else if (split.by=="Microglia") {
    p=ggplot(plotdata, aes(x=Cluster, y=log2Reads, fill=Cluster))+
      geom_violin()+
      geom_beeswarm(alpha=0.4, method = "hex")+
      ggtitle(gene)+
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  }
  p
}

barplot_genes(gene="Ptprc", split.by ="cluster") # alternative split.by = "none" oder " Micorglia
barplot_genes(gene="Ptprc", split.by ="none") # just across the genotype_treatments 
barplot_genes(gene="Ptprc", split.by ="Microglia") # add +coord_cartesian(ylim=c(0, 7)) if you want to change the ylims
barplot_genes(gene="Ptprc", split.by ="Microglia")+coord_cartesian(ylim=c(0, 2))
barplot_genes(gene="Ptprc", split.by ="Microglia") + facet_wrap(~Genotype_Treatment)

```

```{r, eval=FALSE}

DEG_Strokebins_Wt_CtrlvsStroke <-FindMarkers(samples.integrated.filt, subset.ident = "Stroke_bin", subset="TRUE", 
                                             ident.1 = "WT_Stroke", ident.2 = "WT_Ctrl", 
                                             test.use = "LR", logfc.threshold = 0)

display_tab(DEG_Strokebins_Wt_CtrlvsStroke)

DEG_Strokebins_Appps1_CtrlvsStroke <-FindMarkers(samples.integrated.filt, subset.ident = "Stroke_bin", subset="TRUE", 
                                                 ident.1 = "APPPS1_Stroke", ident.2 = "APPPS1_Ctrl", 
                                                 test.use = "LR", logfc.threshold = 0)

display_tab(DEG_Strokebins_Appps1_CtrlvsStroke)

genelist_both= intersect(rownames(DEG_Strokebins_Appps1_CtrlvsStroke), rownames(DEG_Strokebins_Wt_CtrlvsStroke))

plotdata=data.frame(WT=DEG_Strokebins_Wt_CtrlvsStroke[genelist_both,]$avg_log2FC, APPPS1 = DEG_Strokebins_Appps1_CtrlvsStroke[genelist_both,]$avg_log2FC)
rownames(plotdata) = genelist_both

plotdata$label = rownames(plotdata)
n=5
plotdata$label[! c(1:nrow(plotdata)) %in% 
                 c(
                   order(abs(plotdata$WT), decreasing = T)[1:n], 
                   order(abs(plotdata$APPPS1), decreasing = T)[1:n])]=""

plotdata$col="WT"
apps1=abs(plotdata$APPPS1)>abs(plotdata$WT)
plotdata$col[ apps1]="APPPS1"
plotdata$col[(abs(plotdata$APPPS1)<0.25) & (abs(plotdata$WT)<0.25)] <- NA
ggplot(plotdata, aes(x=WT, y=APPPS1, color=col, label=label))+geom_point(alpha=0.7)+
  geom_text(vjust=-0.7, color="black")+geom_abline(slope=-1, intercept=0, lty=2)+geom_abline(slope=1, intercept=0, lty=2)+theme_classic()+ggtitle("log_FC_Stroke_bins")

```

### DAMS in APPS1 Wt vs Stroke

to do selection of DAMS only compare APPPS1CTRL vs APPS1 Stroke to do selection of DAMs only compare APPPS1Stroke vs WT_Stroke

### Dot plot showing top markers for condition clusters, same paper Fig 2f

```{r, fig.width=18, fig.height=5}

reads = samples.integrated.filt@assays$SCT@counts

reads.filt= reads[cluster.genes,]

datatoplot_mean = aggregate(t(reads.filt), 
                            list(paste0(samples.integrated.filt$labels, "/",samples.integrated.filt$Genotype_Treatment)), 
                            mean, na.rm=T) %>% 
  as.data.frame() %>% 
  column_to_rownames("Group.1") %>% scale()



datatoplot_mean=reshape2::melt(datatoplot_mean)
datatoplot_mean$Microglia = factor(gsub("/.*", "", datatoplot_mean$Var1), levels=rev(levels(samples.integrated.filt$labels)))
datatoplot_mean$Genotype_Treatment = factor(gsub(".*/", "", datatoplot_mean$Var1), levels=c("WT_Ctrl", "WT_Stroke", "APPPS1_Ctrl","APPPS1_Stroke")  )                                

colnames(datatoplot_mean) = c("Cluster_Condition", "Gene", "Scaled_mean", "Microglia", "Genotype_Treatment")

datatoplot_perc = aggregate(t(reads.filt), list(paste0(samples.integrated.filt$labels, "/",samples.integrated.filt$Genotype_Treatment)), 
                            function(x){length(which(x>0))/length(x)})

datatoplot_perc=reshape2::melt(datatoplot_perc, value.name = "perc")

datatoplot = cbind(datatoplot_mean, perc=datatoplot_perc$perc)

p <- ggplot(datatoplot, aes(x=Gene, y=Microglia, col=Scaled_mean, size=perc))+geom_point()+scale_size_continuous(range = c(0,3))+
  theme_classic()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+facet_grid(~Genotype_Treatment)

p

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Dotplot_ClsuterCondition_Genelists.svg", p)


```

### TEST - Mike

```{r}

Spatial_mutual <- c("S100a6", "Apoc4", "C4b", "Atp1a2", "Apoc1", "Fn1", "Ntrk2") # genes up in both spatial and APPPS1_stroke vs APPPS1

### Dot plot showing top markers for condition clusters, same paper Fig 2f


reads = samples.integrated.filt@assays$SCT@counts

reads.filt= reads[Spatial_mutual,]

datatoplot_mean = aggregate(t(reads.filt), 
                            list(paste0(samples.integrated.filt$labels, "/",samples.integrated.filt$Genotype_Treatment)), 
                            mean, na.rm=T) %>% 
  as.data.frame() %>% 
  column_to_rownames("Group.1") %>% scale()



datatoplot_mean=reshape2::melt(datatoplot_mean)
datatoplot_mean$Microglia = factor(gsub("/.*", "", datatoplot_mean$Var1), levels=rev(levels(samples.integrated.filt$labels)))
datatoplot_mean$Genotype_Treatment = factor(gsub(".*/", "", datatoplot_mean$Var1), levels=c("WT_Ctrl", "WT_Stroke", "APPPS1_Ctrl","APPPS1_Stroke")  )                                

colnames(datatoplot_mean) = c("Cluster_Condition", "Gene", "Scaled_mean", "Microglia", "Genotype_Treatment")

datatoplot_perc = aggregate(t(reads.filt), list(paste0(samples.integrated.filt$labels, "/",samples.integrated.filt$Genotype_Treatment)), 
                            function(x){length(which(x>0))/length(x)})

datatoplot_perc=reshape2::melt(datatoplot_perc, value.name = "perc")

datatoplot = cbind(datatoplot_mean, perc=datatoplot_perc$perc)

p <- ggplot(datatoplot, aes(x=Gene, y=Microglia, col=Scaled_mean, size=perc))+geom_point()+scale_size_continuous(range = c(0,3))+
  theme_classic()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+facet_grid(~Genotype_Treatment)

p

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Dotplot_Up_spatial_and_APPPS1vsAPPPS1stroke.svg", p)


```
### TEST - Mike

```{r}

Test_genes <- c("S100a6", "S100a9", "S100a8", "S100a10", "S100a11", "S100b", "S100a7a") # genes up in both spatial and APPPS1_stroke vs APPPS1

### Dot plot showing top markers for condition clusters, same paper Fig 2f


reads = samples.integrated.filt@assays$SCT@counts

reads.filt= reads[Test_genes,]

datatoplot_mean = aggregate(t(reads.filt), 
                            list(paste0(samples.integrated.filt$labels, "/",samples.integrated.filt$Genotype_Treatment)), 
                            mean, na.rm=T) %>% 
  as.data.frame() %>% 
  column_to_rownames("Group.1") %>% scale()



datatoplot_mean=reshape2::melt(datatoplot_mean)
datatoplot_mean$Microglia = factor(gsub("/.*", "", datatoplot_mean$Var1), levels=rev(levels(samples.integrated.filt$labels)))
datatoplot_mean$Genotype_Treatment = factor(gsub(".*/", "", datatoplot_mean$Var1), levels=c("WT_Ctrl", "WT_Stroke", "APPPS1_Ctrl","APPPS1_Stroke")  )                                

colnames(datatoplot_mean) = c("Cluster_Condition", "Gene", "Scaled_mean", "Microglia", "Genotype_Treatment")

datatoplot_perc = aggregate(t(reads.filt), list(paste0(samples.integrated.filt$labels, "/",samples.integrated.filt$Genotype_Treatment)), 
                            function(x){length(which(x>0))/length(x)})

datatoplot_perc=reshape2::melt(datatoplot_perc, value.name = "perc")

datatoplot = cbind(datatoplot_mean, perc=datatoplot_perc$perc)

p <- ggplot(datatoplot, aes(x=Gene, y=Microglia, col=Scaled_mean, size=perc))+geom_point()+scale_size_continuous(range = c(0,3))+
  theme_classic()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+facet_grid(~Genotype_Treatment)

p

ggsave(filename = "/files/scSeq_Hefendehl/output/Microglia_Dotplot_test.svg", p)


```

### DEG comparing Cluster 
Change the Cluster A and B Names as you like and run (samll geern arrow on the right) this and the subsequent box (GORes might take a while)

```{r}

Idents(samples.integrated.filt) <- "labels"

compareCluster <- function(ClusterA = "Microglia6", ClusterB = "Microglia0", 
                           seuratObject = samples.integrated.filt, 
                           output ="/files/scSeq_Hefendehl/output/", 
                           genes_to_label=genes_to_label, FCcutoff=1, 
                           comorbid = F){
  if(comorbid){
    resname=paste0(ClusterA,"vs",ClusterB,"_comorbid")
    DEG_tmp <-FindMarkers(seuratObject, assay="SCT", # slot="scale.data",
                          subset.ident = "Genotype_Treatment", 
                          subset = "APPPS1_Stroke",
                          ident.1 = ClusterA, ident.2 = ClusterB, 
                          test.use = "LR", logfc.threshold = 0)
  }else {
    resname=paste0(ClusterA,"vs",ClusterB)
    DEG_tmp <-FindMarkers(seuratObject, assay="SCT", #slot="scale.data",
                               ident.1 = ClusterA, ident.2 = ClusterB, 
                               test.use = "LR", logfc.threshold = 0)
    }
  display_tab(DEG_tmp)
  #genes_to_label[!genes_to_label %in% rownames(DEG_tmp)]
  q <- gg_qqplot(DEG_tmp$p_val)
  p<- EnhancedVolcano::EnhancedVolcano(DEG_tmp, 
                                       x = "avg_log2FC", 
                                       y = "p_val_adj", 
                                       lab=rownames(DEG_tmp), 
                                       #selectLab = genes_to_label,
                                       pCutoff = 0.05,
                                       title = "Microglia6 vs Microglia0")
  ggsave(filename = paste0(output,"DEG_Vulcano_",resname,".svg"), p|q, 
         width = 8,height = 4)
  gene_univers = rownames(seuratObject)
  gogenes=list(
    genesigall=rownames(DEG_tmp)[DEG_tmp$p_val_adj<0.05&abs(DEG_tmp$avg_log2FC)>FCcutoff],
    genesigup=rownames(DEG_tmp)[DEG_tmp$p_val_adj<0.05&DEG_tmp$avg_log2FC>FCcutoff],
    genesigdown=rownames(DEG_tmp)[DEG_tmp$p_val_adj<0.05&DEG_tmp$avg_log2FC< -FCcutoff])
  
  Gores = getGOresults(geneset = gogenes, domain_scope = "known",
                       genereference = gene_univers, 
                       return_genelist = T,
                       organism = "mmusculus")
  openxlsx::write.xlsx(Gores$result%>% dplyr::select(-parents),
                       file=paste0(output,"DEG_Genelist_",resname,".xlsx"))
  
  plotlist <- list()
  for(m in unique(Gores$result$query)){
    idx=Gores$result$query ==m & grepl("GO",Gores$result$source)
    if(sum(idx)>=1){
      
    tmp <- GOplot(Gores$result[idx,], N = 10, Title = m)
    plotlist[[m]] <- tmp
    } else {
      plotlist[[m]] <-  ggplot() + annotate("text", x = 10, y = 10, size = 6,
		label = "no significant GO term\ncheck excel sheet for other enrichments") + theme_void() +ggtitle(m)

    }
  }
  
  
  ggsave(filename = paste0(output,"DEG_GO_plots",resname,".svg"), 
         ggarrange(plotlist=plotlist), width = 8, height = 4)
         
  return(list(Markers = DEG_tmp, 
              VulcanoPlot=p, 
              qqplot =q, 
              Goresults=Gores$result, 
              GOplots=plotlist))
}

```


##### Cluster 6 vs 0

```{r, fig.width=8, fig.height=4}

res <- compareCluster(ClusterA = "Microglia6" , 
                        ClusterB = "Microglia0", FCcutoff = 0.5)

res$VulcanoPlot | res$qqplot
ggarrange(plotlist=res$GOplots)

res$Markers

## all results are saved as DEG_*_ClusterAvsClusterB
```

##### Cluster 6 vs 0 comorbid 

```{r, fig.width=8, fig.height=4}

res <- compareCluster(ClusterA = "Microglia6" , 
                        ClusterB = "Microglia0", FCcutoff = 0.5, comorbid = T)

res$VulcanoPlot | res$qqplot
ggarrange(plotlist=res$GOplots)
res$Markers

## all results are saved as DEG_*_ClusterAvsClusterB
```

##### Cluster 5 vs 0
```{r, fig.width=8, fig.height=4}

res <- compareCluster(ClusterA = "Microglia5" , 
                        ClusterB = "Microglia0", FCcutoff = 0.5)

res$VulcanoPlot | res$qqplot
ggarrange(plotlist=res$GOplots)
res$Markers

## all results are saved as DEG_*_ClusterAvsClusterB
``` 

##### Cluster 5 vs 0 comorbid 
```{r, fig.width=8, fig.height=4}

res <- compareCluster(ClusterA = "Microglia5" , 
                        ClusterB = "Microglia0", FCcutoff = 0.5, comorbid = T)

res$VulcanoPlot | res$qqplot
ggarrange(plotlist=res$GOplots)
res$Markers

## all results are saved as DEG_*_ClusterAvsClusterB
```

##### Cluster 3 vs 2

```{r, fig.width=8, fig.height=4}

res <- compareCluster(ClusterA = "Microglia3" , 
                        ClusterB = "Microglia2", FCcutoff = 0.5)

res$VulcanoPlot | res$qqplot
ggarrange(plotlist=res$GOplots)
res$Markers

## all results are saved as DEG_*_ClusterAvsClusterB
```

##### Cluster 3 vs 2 comorbid
```{r, fig.width=8, fig.height=4}

res <- compareCluster(ClusterA = "Microglia3" , 
                        ClusterB = "Microglia2", FCcutoff = 0.5, comorbid = T)

res$VulcanoPlot | res$qqplot
ggarrange(plotlist=res$GOplots)
res$Markers

## all results are saved as DEG_*_ClusterAvsClusterB
```

##### Cluster 5 vs 6
```{r, fig.width=8, fig.height=4}

res <- compareCluster(ClusterA = "Microglia5" , 
                        ClusterB = "Microglia6", FCcutoff = 0.5)

res$VulcanoPlot | res$qqplot
ggarrange(plotlist=res$GOplots)
res$Markers

## all results are saved as DEG_*_ClusterAvsClusterB
```

##### Cluster 5 vs 6 comorbid 
```{r, fig.width=8, fig.height=4}

res <- compareCluster(ClusterA = "Microglia5" , 
                        ClusterB = "Microglia6", FCcutoff = 0.5, comorbid = T)

res$VulcanoPlot | res$qqplot
ggarrange(plotlist=res$GOplots)
res$Markers

## all results are saved as DEG_*_ClusterAvsClusterB
```

##### Cluster 2 vs 4 
```{r, fig.width=8, fig.height=4}

res <- compareCluster(ClusterA = "Microglia2" , 
                        ClusterB = "Microglia4", FCcutoff = 0.5)

res$VulcanoPlot | res$qqplot
ggarrange(plotlist=res$GOplots)
res$Markers

## all results are saved as DEG_*_ClusterAvsClusterB
```

##### Cluster 2 vs 4 comorbid 
```{r, fig.width=8, fig.height=4}

res <- compareCluster(ClusterA = "Microglia2" , 
                        ClusterB = "Microglia4", FCcutoff = 0.5, comorbid = T)

res$VulcanoPlot | res$qqplot
ggarrange(plotlist=res$GOplots)
res$Markers

## all results are saved as DEG_*_ClusterAvsClusterB
```


## Pseudotime analysis

### slinghsot  starting Microglia 0 (DAM) for reviewer
```{r}
samples.integrated.filt.meta = samples.integrated.filt@meta.data
sce.microglia=as.SingleCellExperiment(samples.integrated.filt)
sce.microglia <- slingshot(sce.microglia, clusterLabels = 'labels', reducedDim = 'UMAP', start.clus="Microglia0")

samples.integrated.filt.meta$Pseudotime_1 <- sce.microglia$slingPseudotime_1
samples.integrated.filt.meta$Pseudotime_2 <- sce.microglia$slingPseudotime_2
samples.integrated.filt.meta$Pseudotime_3 <- sce.microglia$slingPseudotime_3

samples.integrated.filt.meta -> samples.integrated.filt@meta.data

dataset = data.frame(reducedDim(sce.microglia, "UMAP"), 
                     cluster = as.character(sce.microglia$labels), 
                     DAMs = as.character(sce.microglia$DAM_binary))

curves <- slingCurves(sce.microglia, as.df=T)

p = ggplot(dataset, aes(x=umap_1, y=umap_2))+
  geom_point(aes(col = cluster), size=2)+
  geom_point(data=subset(dataset, dataset$DAMs==T), aes(fill=DAMs), pch=21, size=0.7)+
  scale_fill_manual(values="black")
curves = curves %>% arrange(Lineage, Order)

p = p + geom_path(data = curves, aes(group = Lineage), lwd=1, lineend = "butt",)+theme_classic()

curves_arrows = curves %>% group_by(Lineage) %>%  mutate(xend = lead(umap_1), yend = lead(umap_2))

curves_arrows = curves_arrows[curves_arrows$Order %in% seq(10,150, by=floor((150-10)/10)),]

p = p + geom_segment(data = curves_arrows,
                     aes(x = umap_1, y = umap_2, xend = xend, yend = yend),
                     linewidth=0.5, 
                     arrow = arrow(length=unit(0.3, "cm"), type="closed"))
p

ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime_linages_startDAM.svg", p)
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime_linages_startDAM.pdf", p)
```


```{r, fig.width=12, fig.height=4}
p1 <- FeaturePlot_scCustom(samples.integrated.filt, features = "Pseudotime_1", colors_use = viridis_dark_high)
p2 <- FeaturePlot_scCustom(samples.integrated.filt, features = "Pseudotime_2", colors_use = viridis_dark_high)
p3 <- FeaturePlot_scCustom(samples.integrated.filt, features = "Pseudotime_3", colors_use = viridis_dark_high)


p1|p2|p3
```


### slinghsot  starting Microglia 2
```{r}

samples.integrated.filt.meta = samples.integrated.filt@meta.data
sce.microglia=as.SingleCellExperiment(samples.integrated.filt)
sce.microglia <- slingshot(sce.microglia, clusterLabels = 'labels', reducedDim = 'UMAP', start.clus="Microglia2")

samples.integrated.filt.meta$Pseudotime_1 <- sce.microglia$slingPseudotime_1
samples.integrated.filt.meta$Pseudotime_2 <- sce.microglia$slingPseudotime_2
samples.integrated.filt.meta$Pseudotime_3 <- sce.microglia$slingPseudotime_3

samples.integrated.filt.meta -> samples.integrated.filt@meta.data

dataset = data.frame(reducedDim(sce.microglia, "UMAP"), 
                     cluster = as.character(sce.microglia$labels), 
                     DAMs = as.character(sce.microglia$DAM_binary))

curves <- slingCurves(sce.microglia, as.df=T)

p = ggplot(dataset, aes(x=umap_1, y=umap_2))+
  geom_point(aes(col = cluster), size=2)+
  geom_point(data=subset(dataset, dataset$DAMs==T), aes(fill=DAMs), pch=21, size=0.7)+
  scale_fill_manual(values="black")
curves = curves %>% arrange(Lineage, Order)

p = p + geom_path(data = curves, aes(group = Lineage), lwd=1, lineend = "butt",)+theme_classic()

curves_arrows = curves %>% group_by(Lineage) %>%  mutate(xend = lead(umap_1), yend = lead(umap_2))

curves_arrows = curves_arrows[curves_arrows$Order %in% seq(10,150, by=floor((150-10)/10)),]

p = p + geom_segment(data = curves_arrows,
                     aes(x = umap_1, y = umap_2, xend = xend, yend = yend),
                     linewidth=0.5, 
                     arrow = arrow(length=unit(0.3, "cm"), type="closed"))
p

ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime_linages.svg", p)
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime_linages.pdf", p)

library(TSCAN)

n.top.gene = 10

sce.microglia.PT1<- testPseudotime(sce.microglia, pseudotime = sce.microglia$slingPseudotime_1)
sce.microglia.PT2<- testPseudotime(sce.microglia, pseudotime = sce.microglia$slingPseudotime_2)
sce.microglia.PT3<- testPseudotime(sce.microglia, pseudotime = sce.microglia$slingPseudotime_3)


```

Look for genes that are up/down-regulated along pseudotime, for the re-clustered cells (Same paper Fig 2G/H)

#### Trajectory 1 genes

By default, estimates of the spline coefficients are not returned as they are difficult to interpret. Rather, a log-fold change of expression along each path is estimated to provide some indication of the overall magnitude and direction of any change.

```{r}

sce.microglia.PT1 <- sce.microglia.PT1[order(sce.microglia.PT1$FDR, decreasing = F),]
display_tab(as.data.frame(sce.microglia.PT1))
write.xlsx(as.data.frame(sce.microglia.PT1), file = "/files/scSeq_Hefendehl/output/Pseudotime1_genes.xlsx")
PT1_up = sce.microglia.PT1[sce.microglia.PT1$logFC>0,]
PT1_low = sce.microglia.PT1[sce.microglia.PT1$logFC<0,]

p <- FeaturePlot_scCustom(samples.integrated.filt, features = "Pseudotime_1", colors_use = viridis_dark_high)
p
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime1.svg", p)

```

```{r, fig.width=8, fig.height=5}
a=scater::plotExpression(sce.microglia, features=rownames(PT1_up)[1:n.top.gene], x="slingPseudotime_1", colour_by="labels")+scale_color_hue()
b=scater::plotExpression(sce.microglia, features=rownames(PT1_low)[1:n.top.gene], x="slingPseudotime_1", colour_by="labels")+scale_color_hue()
a|b
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime1_Top10up_down_reg_Genes-correct.svg", a|b)

```

#### Trajectory 2 genes

```{r}
sce.microglia.PT2 <- sce.microglia.PT2[order(sce.microglia.PT2$FDR, decreasing = F),]
display_tab(as.data.frame(sce.microglia.PT2))
write.xlsx(as.data.frame(sce.microglia.PT2), 
           file = "/files/scSeq_Hefendehl/output/Pseudotime2_genes.xlsx")

PT2_up = sce.microglia.PT2[sce.microglia.PT2$logFC>0,]
PT2_low = sce.microglia.PT2[sce.microglia.PT2$logFC<0,]
p <- FeaturePlot_scCustom(samples.integrated.filt, features = "Pseudotime_2", colors_use = viridis_dark_high)
p
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime2.svg", p)
```


```{r, fig.width=8, fig.height=5}
a<-scater::plotExpression(sce.microglia, features=rownames(PT2_up)[1:n.top.gene], x="slingPseudotime_2", colour_by="labels")+scale_color_hue()
b<-scater::plotExpression(sce.microglia, features=rownames(PT2_low)[1:n.top.gene], x="slingPseudotime_2", colour_by="labels")+scale_color_hue()

a|b

ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime2_Top10up_down_reg_Genes.svg", a|b)


```

#### Trajectory 3 genes

```{r}
sce.microglia.PT3 <- sce.microglia.PT3[order(sce.microglia.PT3$FDR, decreasing = F),]
display_tab(as.data.frame(sce.microglia.PT3))
write.xlsx(as.data.frame(sce.microglia.PT3), 
           file = "/files/scSeq_Hefendehl/output/Pseudotime3_genes.xlsx")

PT3_up = sce.microglia.PT3[sce.microglia.PT3$logFC>0,]
PT3_low = sce.microglia.PT3[sce.microglia.PT3$logFC<0,]

p <- FeaturePlot_scCustom(samples.integrated.filt, features = "Pseudotime_3",  colors_use = viridis_dark_high)
p
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime3.svg", p)
```




```{r, fig.width=8, fig.height=5}
a<-scater::plotExpression(sce.microglia, features=rownames(PT3_up)[1:n.top.gene], x="slingPseudotime_3", colour_by="labels")+scale_color_hue()
b<-scater::plotExpression(sce.microglia, features=rownames(PT3_low)[1:n.top.gene], x="slingPseudotime_3", colour_by="labels")+scale_color_hue()
a|b
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime3_Top10up_down_reg_Genes.svg", a|b)
```


### Pseudotime starting cluster 4


### slinghsot 

```{r}

samples.integrated.filt.meta = samples.integrated.filt@meta.data
sce.microglia=as.SingleCellExperiment(samples.integrated.filt)
sce.microglia <- slingshot(sce.microglia, clusterLabels = 'labels', reducedDim = 'UMAP', start.clus="Microglia4")

samples.integrated.filt.meta$Pseudotime_1_Start4 <- sce.microglia$slingPseudotime_1
samples.integrated.filt.meta$Pseudotime_2_Start4 <- sce.microglia$slingPseudotime_2
samples.integrated.filt.meta$Pseudotime_3_Start4 <- sce.microglia$slingPseudotime_3

samples.integrated.filt.meta -> samples.integrated.filt@meta.data

dataset = data.frame(reducedDim(sce.microglia, "UMAP"), 
                     cluster = as.character(sce.microglia$labels), 
                     DAMs = as.character(sce.microglia$DAM_binary))

curves <- slingCurves(sce.microglia, as.df=T)

p = ggplot(dataset, aes(x=umap_1, y=umap_2))+
  geom_point(aes(col = cluster), size=2)+
  geom_point(data=subset(dataset, dataset$DAMs==T), aes(fill=DAMs), pch=21, size=0.7)+
  scale_fill_manual(values="black")
curves = curves %>% arrange(Lineage, Order)

p = p + geom_path(data = curves, aes(group = Lineage), lwd=1, lineend = "butt",)+theme_classic()

curves_arrows = curves %>% group_by(Lineage) %>%  mutate(xend = lead(umap_1), yend = lead(umap_2))

curves_arrows = curves_arrows[curves_arrows$Order %in% seq(10,150, by=floor((150-10)/10)),]

p = p + geom_segment(data = curves_arrows,
                     aes(x = umap_1, y = umap_2, xend = xend, yend = yend),
                     linewidth=0.5, 
                     arrow = arrow(length=unit(0.3, "cm"), type="closed"))
p

ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime_linages_Start4.svg", p)
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime_linages_Start4.pdf", p)

library(TSCAN)

n.top.gene = 10

sce.microglia.PT1<- testPseudotime(sce.microglia, pseudotime = sce.microglia$slingPseudotime_1)
sce.microglia.PT2<- testPseudotime(sce.microglia, pseudotime = sce.microglia$slingPseudotime_2)
sce.microglia.PT3<- testPseudotime(sce.microglia, pseudotime = sce.microglia$slingPseudotime_3)


```

Look for genes that are up/down-regulated along pseudotime, for the re-clustered cells (Same paper Fig 2G/H)

#### Trajectory 1 genes

By default, estimates of the spline coefficients are not returned as they are difficult to interpret. Rather, a log-fold change of expression along each path is estimated to provide some indication of the overall magnitude and direction of any change.

```{r}

sce.microglia.PT1 <- sce.microglia.PT1[order(sce.microglia.PT1$FDR, decreasing = F),]
display_tab(as.data.frame(sce.microglia.PT1))

PT1_up = sce.microglia.PT1[sce.microglia.PT1$logFC>0,]
PT1_low = sce.microglia.PT1[sce.microglia.PT1$logFC<0,]

p <- FeaturePlot_scCustom(samples.integrated.filt, features = "Pseudotime_1_Start4", colors_use = viridis_dark_high)
p
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime1_Start4.svg", p)

```

```{r, fig.width=8, fig.height=5}
a=scater::plotExpression(sce.microglia, features=rownames(PT1_up)[1:n.top.gene], x="slingPseudotime_1", colour_by="labels")+scale_color_hue()
b=scater::plotExpression(sce.microglia, features=rownames(PT1_low)[1:n.top.gene], x="slingPseudotime_1", colour_by="labels")+scale_color_hue()
a|b
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime1_Top10up_down_reg_Genes-correct_Start4.svg", a|b)

```

#### Trajectory 2 genes

```{r}
sce.microglia.PT2 <- sce.microglia.PT2[order(sce.microglia.PT2$FDR, decreasing = F),]
display_tab(as.data.frame(sce.microglia.PT2))

PT2_up = sce.microglia.PT2[sce.microglia.PT2$logFC>0,]
PT2_low = sce.microglia.PT2[sce.microglia.PT2$logFC<0,]
p <- FeaturePlot_scCustom(samples.integrated.filt, features = "Pseudotime_2_Start4", colors_use = viridis_dark_high)
p
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime2Start4.svg", p)
```


```{r, fig.width=8, fig.height=5}
a<-scater::plotExpression(sce.microglia, features=rownames(PT2_up)[1:n.top.gene], x="slingPseudotime_2", colour_by="labels")+scale_color_hue()
b<-scater::plotExpression(sce.microglia, features=rownames(PT2_low)[1:n.top.gene], x="slingPseudotime_2", colour_by="labels")+scale_color_hue()

a|b

ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime2_Top10up_down_reg_GenesStart4.svg", a|b)


```

#### Trajectory 3 genes

```{r}
sce.microglia.PT3 <- sce.microglia.PT3[order(sce.microglia.PT3$FDR, decreasing = F),]
display_tab(as.data.frame(sce.microglia.PT3))


PT3_up = sce.microglia.PT3[sce.microglia.PT3$logFC>0,]
PT3_low = sce.microglia.PT3[sce.microglia.PT3$logFC<0,]

p <- FeaturePlot_scCustom(samples.integrated.filt, features = "Pseudotime_3_Start4",  colors_use = viridis_dark_high)
p
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime3Start4.svg", p)
```






```{r, fig.width=8, fig.height=5}
a<-scater::plotExpression(sce.microglia, features=rownames(PT3_up)[1:n.top.gene], x="slingPseudotime_3", colour_by="labels")+scale_color_hue()
b<-scater::plotExpression(sce.microglia, features=rownames(PT3_low)[1:n.top.gene], x="slingPseudotime_3", colour_by="labels")+scale_color_hue()
a|b
ggsave(filename = "/files/scSeq_Hefendehl/output/Pseudotime3_Top10up_down_reg_Genes.svg", a|b)
```


## monocle 

```{r}
#library(monocle3)


```


### Marker for DAM onyl DEG for APPPS1_Ctrl vs APPPS_1 Stroke; WT_Stroke vs APPPS1_Stroke

### WGCNA

In the paper <https://www.nature.com/articles/s41586-018-0023-4> this analysis was used to generate different modules of gene lists that can then be mapped onto different treatment/genotypes within one experiment (see Figure 4). This is an analysis that can be done in Seurat <https://smorabit.github.io/tutorials/9_scWGCNA_tutorial/> Which could be very cool for our data. In the paper they also have 3 different conditions in 2 genotypes.

avoid running this section as it takes very long

```{r}
nclusters <- length(unique(samples.integrated.filt$labels))
genes.use <- rownames(samples.integrated.filt)
targets <- samples.integrated.filt@meta.data

group <- as.factor(samples.integrated.filt$Genotype_Treatment)
```

```{r fig.width=8, fig.height=6}
force.recalc = T
powers = c(seq(1,10,by=1), seq(12,20, by=2));

# skips recalculation of net object as it takes long... 

datExpr <- as.data.frame(GetAssayData(samples.integrated.filt, assay='RNA', layer ='data')[genes.use,])
datExpr <- as.data.frame(t(datExpr))
datExpr <- datExpr[,goodGenes(datExpr)]

# expressed in more than 10% of cells 
exprgenes = colSums(datExpr>0)>nrow(datExpr)*0.1
datExpr <- datExpr[,exprgenes]

#annotated only 


mmacc=as.list(org.Mm.egALIAS2EG)
annotatedgenes = colnames(datExpr) %in% names(mmacc)
datExpr <- datExpr[,annotatedgenes]

powers = c(seq(1,10,by=1), seq(12,20, by=2));


enableWGCNAThreads()

if(force.recalc | (! file.exists(file="/files/scSeq_Hefendehl/output/powertable.rds"))){  
  # Call the network topology analysis function for each set in turn
  powerTable = list(
    data = pickSoftThreshold(
      datExpr,
      powerVector=powers,
      verbose = 100,
      networkType="unsigned",
      corFnc="cor"
    )[[2]]
  );
  saveRDS(powerTable, file="/files/scSeq_Hefendehl/output/powertable.rds")
} else {
  powerTable = readRDS(file="/files/scSeq_Hefendehl/output/powertable.rds")
}

colors = c("blue", "red","black")
# Will plot these columns of the returned scale free analysis tables
plotCols = c(2,5,6,7)
colNames = c("Scale Free Topology Model Fit", "Mean connectivity", "mean connectivity",
             "Max connectivity");

# Get the minima and maxima of the plotted points
ylim = matrix(NA, nrow = 2, ncol = 4);
for (col in 1:length(plotCols)){
  ylim[1, col] = min(ylim[1, col], powerTable$data[, plotCols[col]], na.rm = TRUE);
  ylim[2, col] = max(ylim[2, col], powerTable$data[, plotCols[col]], na.rm = TRUE);
}

# Plot the quantities in the chosen columns vs. the soft thresholding power

par(mfcol = c(2,2));
par(mar = c(4.2, 4.2 , 2.2, 0.5))
cex1 = 0.7;

for (col in 1:length(plotCols)){
  plot(powerTable$data[,1], -sign(powerTable$data[,3])*powerTable$data[,2],
       xlab="Soft Threshold (power)",ylab=colNames[col],type="n", ylim = ylim[, col],
       main = colNames[col]);
  addGrid();
  
  if (col==1){
    text(powerTable$data[,1], -sign(powerTable$data[,3])*powerTable$data[,2],
         labels=powers,cex=cex1,col=colors[1]);
  } else
    text(powerTable$data[,1], powerTable$data[,plotCols[col]],
         labels=powers,cex=cex1,col=colors[1]);
  if (col==1){
    legend("bottomright", legend = 'Metacells', col = colors, pch = 20) ;
  } else
    legend("topright", legend = 'Metacells', col = colors, pch = 20) ;
}

softPower=3

nSets = 1
setLabels = 'OCS'
shortLabels = setLabels

multiExpr <- list()
multiExpr[['ODC']] <- list(data=datExpr)

checkSets(multiExpr) # check data size

# construct network
if(force.recalc | (! file.exists(file="/files/scSeq_Hefendehl/output/net.rds"))){
  net=blockwiseConsensusModules(multiExpr, blocks = NULL,
                                maxBlockSize = 30000, ## This should be set to a smaller size if the user has limited RAM
                                randomSeed = 12345,
                                corType = "pearson",
                                power = softPower,
                                consensusQuantile = 0.3,
                                networkType = "unsigned",
                                TOMType = "signed",
                                TOMDenom = "min",
                                scaleTOMs = TRUE, scaleQuantile = 0.8,
                                sampleForScaling = TRUE, sampleForScalingFactor = 1000,
                                useDiskCache = TRUE, chunkSize = NULL,
                                deepSplit = 4,
                                pamStage=FALSE,
                                detectCutHeight = 0.995, minModuleSize = 50,
                                mergeCutHeight = 0.2,
                                saveConsensusTOMs = TRUE,
                                consensusTOMFilePattern = "ConsensusTOM-block.%b.rda")
  saveRDS(net, file="/files/scSeq_Hefendehl/output/net.rds")
  
} else {
  net = readRDS(file="/files/scSeq_Hefendehl/output/net.rds")
}





consMEs = net$multiMEs;
moduleLabels = net$colors;


# Convert the numeric labels to color labels
moduleColors = as.character(moduleLabels)
consTree = net$dendrograms[[1]];



# module eigengenes
MEs=moduleEigengenes(multiExpr[[1]]$data, colors = moduleColors, nPC=1)$eigengenes
MEs=orderMEs(MEs)
meInfo<-data.frame(rownames(datExpr), MEs)
colnames(meInfo)[1]= "SampleID"

# intramodular connectivity
KMEs<-signedKME(datExpr, MEs,outputColumnName = "kME",corFnc = "bicor")

# compile into a module metadata table
geneInfo=as.data.frame(cbind(colnames(datExpr),moduleColors, KMEs))

# how many modules did we get?
nmodules <- length(unique(moduleColors))

# merged gene symbol column
colnames(geneInfo)[1]= "GeneSymbol"
colnames(geneInfo)[2]= "Initially.Assigned.Module.Color"

# save info
write.csv(geneInfo,file=paste0('/files/scSeq_Hefendehl/output/geneInfoSigned.csv'))

PCvalues=MEs
```


```{r}
p = WGCNA::plotDendroAndColors(consTree, moduleColors[net$goodGenes], "Module colors", dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05,
                               main = paste0("Microglia  gene dendrogram and module colors"))

ggsave(file="/files/scSeq_Hefendehl/output/Microglia_WGCNA_Dendrocolplot.svg")

```
##### WGCNA by Genotype_treatment

```{r, fig.width=7, fig.height=4}

plot_df <- cbind(dplyr::select(targets, c(Genotype, Treatment, labels)), PCvalues)
plot_df <- reshape2::melt(plot_df, id.vars = c('Genotype', 'Treatment', "labels"))
plot_df$Genotype <- factor(plot_df$Genotype, levels=c('WT','APPPS1'))
plot_df$Genotype_Treatment <- paste0(plot_df$Genotype, "_", plot_df$Treatment)
plot_df$Genotype_Treatment <- factor(plot_df$Genotype_Treatment, levels=c('WT_Ctrl','WT_Stroke', "APPPS1_Ctrl", "APPPS1_Stroke"))

colors <- sub('ME', '', as.character(levels(plot_df$variable)))
p <- ggplot(plot_df, aes(x=variable, y=value, fill=Treatment)) +
  geom_boxplot(notch=FALSE) +
  RotatedAxis() + ylab('Module Eigengene') + xlab('') +
  theme(
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
  )

p <- p + facet_wrap(~variable+Genotype, scales='free_x', ncol=nmodules)+theme_classic()

p
```

##### WGCNA violin by condition
```{r, fig.height=2, fig.width=8}
p2 <- ggplot(plot_df, aes(x=variable, y=value, fill=Genotype_Treatment)) +
  geom_violin() +
  # RotatedAxis() + 
  ylab('Module Eigengene') + xlab('') +
  theme(
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
  )

p2 <- p2 + facet_wrap(~variable, scales='free', ncol=nmodules)+theme_classic()
p2
ggsave(file="/files/scSeq_Hefendehl/output/Microgila_WGCNA_Eigenscore.svg", p2)

```
##### WGCNA violin by cluster
```{r, fig.height=2, fig.width=10}
p2 <- ggplot(plot_df, aes(x=variable, y=value, fill=labels)) +
  geom_violin() +
  # RotatedAxis() + 
  ylab('Module Eigengene') + xlab('') +
  theme(
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
  )

p2 <- p2 + facet_wrap(~variable, scales='free', ncol=nmodules)+theme_classic()
p2
ggsave(file="/files/scSeq_Hefendehl/output/Microgila_WGCNA_Eigenscore_byCluster.svg", p2)

```

##### WGCNA Heatmap  by condition

```{r, fig.width=6, fig.height=4}
reslist = list()
for(M in unique(plot_df$variable)){
  
  Mdata = plot_df[plot_df$variable==M,]
  fit= lm(value~Genotype*Treatment, data = Mdata)
  resfit = summary(fit)
  reslist[[M]] <- c(Genotype_Treatment = resfit$coefficients["GenotypeAPPPS1:TreatmentStroke", "Pr(>|t|)"])
}

ME_avg <- plot_df %>% dplyr::group_by(variable, Treatment, Genotype) %>% summarise(avg_Eigenmodule = mean(value),
                                                                                   sd_Eigenvalue = sd(value),
                                                                                   med_Eigenvalue=median(value)) 
ME_avg$pval=unlist(reslist[ME_avg$variable])
ME_avg$pval_ast=convertpvaltostars(ME_avg$pval)
ME_avg$Genotype_Treatment = paste0(ME_avg$Genotype, "_",ME_avg$Treatment)
ME_avg$Genotype_Treatment <- factor(ME_avg$Genotype_Treatment, levels=c('WT_Ctrl','WT_Stroke', "APPPS1_Ctrl", "APPPS1_Stroke"))
ME_avg$Module = paste0(ME_avg$pval_ast, ME_avg$variable)

ME_avg
openxlsx::write.xlsx(ME_avg, file="/files/scSeq_Hefendehl/output/WGCNA_LinModStat_GTxTreatm.xlsx")

```


```{r}
Meandata = MEs %>% group_by(samples.integrated.filt$Genotype_Treatment) %>% summarise_all(mean) %>% column_to_rownames("samples.integrated.filt$Genotype_Treatment") %>% t()
SDdata = MEs %>% group_by(samples.integrated.filt$Genotype_Treatment) %>% summarise_all(sd)%>% column_to_rownames("samples.integrated.filt$Genotype_Treatment") %>% t()
colLabels=colnames(Meandata)
rowLabels=rownames(Meandata)
rowLabels=paste0(convertpvaltostars(unlist(reslist[rowLabels])), " ",rowLabels=rownames(Meandata))


textlab = paste0(signif(Meandata,2), "\n", "(", signif(SDdata, 2), ")")

svg("/files/scSeq_Hefendehl/output/WGCNA_labeled_Heatmap.svg")
par(mar=c(5,8,5,5))
labeledHeatmap(Meandata, xLabels = colLabels, 
               yLabels = rowLabels,
               yColorLabels = T,
               colors = viridis(50),
               setStdMargins = FALSE, textMatrix = textlab,
               cex.col="white",
               main = "Mean(SD) Eigengene-expression");
dev.off()
```
##### WGCNA Heatmap  by cluster

```{r, fig.width=6, fig.height=4}
reslist = list()
for(M in unique(plot_df$variable)){
  
  Mdata = plot_df[plot_df$variable==M,]
  fit= aov(value~labels, data = Mdata)
  resfit = summary(fit)
  reslist[[M]] <- c(Clusters = resfit[[1]]["labels", "Pr(>F)"])
}

ME_avg <- plot_df %>% dplyr::group_by(variable, labels) %>% summarise(avg_Eigenmodule = mean(value),
                                                                      sd_Eigenvalue = sd(value),
                                                                      med_Eigenvalue=median(value)) 
ME_avg$pval=unlist(reslist[ME_avg$variable])
ME_avg$pval_ast=convertpvaltostars(ME_avg$pval)
ME_avg$Module = paste0(ME_avg$pval_ast, ME_avg$variable)

ME_avg
openxlsx::write.xlsx(ME_avg, file="/files/scSeq_Hefendehl/output/WGCNA_LinModStat_ByCluster.xlsx")

```

```{r}
Meandata = MEs %>% group_by(samples.integrated.filt$labels) %>% summarise_all(mean) %>% column_to_rownames("samples.integrated.filt$labels") %>% t()
SDdata = MEs %>% group_by(samples.integrated.filt$labels) %>% summarise_all(sd)%>% column_to_rownames("samples.integrated.filt$labels") %>% t()
colLabels=colnames(Meandata)
rowLabels=rownames(Meandata)
rowLabels=rownames(Meandata)
rowLabels=paste0(convertpvaltostars(unlist(reslist[rowLabels])), " ",rowLabels=rownames(Meandata))
textlab = paste0(signif(Meandata,2), "\n", "(", signif(SDdata, 2), ")")

svg("/files/scSeq_Hefendehl/output/WGCNA_labeled_Heatmap_byCluster.svg")
par(mar=c(5,8,5,5))
labeledHeatmap(Meandata, xLabels = colLabels, 
               yLabels = rowLabels,
               colorLabels = TRUE,
               colors = viridis(50),
               setStdMargins = FALSE, textMatrix = textlab,
               cex.col="white",
               main = "Mean(SD) Eigengene-expression");
dev.off()
```

##### Heatmaps Genewise per module 
```{r}

Modules = unique(moduleColors)
Modules = Modules[! Modules %in% "grey"] # drop grey one

for(m in Modules){
  
  p = DoHeatmap(samples.integrated.filt, features = rownames(samples.integrated.filt)[moduleColors==m], slot="counts",
                group.by = "Genotype_Treatment")+scale_fill_viridis()
  
  ggsave(paste0("/files/scSeq_Hefendehl/output/WGCNAGeneWise_ExpressionModule", m, ".svg"), p)
}


```


##### Heatmpas top 50 genes per module Conditions wise
```{r, fig.width=14, fig.height=14}
genelist = split(names(net$colors), net$colors)
gene_univers = names(net$colors)
plotlist = list()

reads = rowSums(samples.integrated.filt@assays$RNA$counts)


for (m in names(genelist)){
  
  targets = genelist[[m]]
  targets <- reads[targets] %>% sort(., decreasing = T)
  targets <- names(targets)[1:50]
  
  p<-DoHeatmap(samples.integrated.filt,features = targets, 
               group.by = "Genotype_Treatment", slot = "data")
  ggsave(p, file=paste0("output/WGCNA_Module_",m,"_allgenes_Heatmaps.svg"))
  plotlist[[m]]<-p
}

p<- ggarrange(plotlist = plotlist, labels = names(genelist))
p
ggsave(file="output/WGCNA_Module_Heatmaps.svg", p)

names(plotlist)

```

```{r, fig.width=5, fig.height=5}
module = "green"
featureslist=c("Apoe", "Axl", "Cst7", "Tyrobp","Cd63", "Clec7a", "H2-D1", "Lyz2", "Npc2")
svg(paste0("./output/WGCNA_", module, "selectedgenes_heatmap.svg"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "Genotype_Treatment", slot = "data")
dev.off()
```

```{r, fig.width=5, fig.height=5}
module = "brown"
featureslist=c("Trem2", "Csf1r", "C1qa", "C1qb", "C1qc", "Ctss",
               "Ctsl", "Ctsh", "Cyba", "Lamp1")
svg(paste0("./output/WGCNA_", module, "selectedgenes_heatmap.svg"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "Genotype_Treatment", slot = "data")
dev.off()
```

```{r, fig.width=5, fig.height=5}
module = "turqouise"
featureslist=c("Olfr1307", "Olfr635", "Olfr1336", "Olfr46", 
               "Olfr344", "Fgd4", "Pkib", "Shisa5", "Senp7", "Snhg14")
svg(paste0("./output/WGCNA_", module, "selectedgenes_heatmap.svg"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "Genotype_Treatment", slot = "data")
dev.off()
```

```{r, fig.width=5, fig.height=5}
module = "grey"
featureslist=c("Adam10", "Abca1", "CD74","Arpc1b", "Eef2",
               "Itgb2", "Pdia3", "Aldoa", "Cfh", "Ssh2")
svg(paste0("./output/WGCNA_", module, "selectedgenes_heatmap.svg"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "Genotype_Treatment", slot = "data")
dev.off()
```


##### Heatmpas top 50 genes per module Cluster wise
```{r, fig.width=14, fig.height=14}
plotlist = list()

reads = rowSums(samples.integrated.filt@assays$RNA$counts)


for (m in names(genelist)){
  
  targets = genelist[[m]]
  targets <- reads[targets] %>% sort(., decreasing = T)
  targets <- names(targets)[1:50]
  
  p<-DoHeatmap(samples.integrated.filt, features = targets, 
               group.by = "labels", slot = "data")
  ggsave(p, file=paste0("output/WGCNA_Module_",m,"_Heatmaps_byCluster.svg"))
  plotlist[[m]]<-p
}

p<- ggarrange(plotlist = plotlist, labels = names(genelist))
p
ggsave(file="output/WGCNA_Module_Heatmaps_byCluster.svg", p)

names(plotlist)

```


```{r, fig.width=5, fig.height=5}
module = "green"
featureslist=c("Apoe", "Axl", "Cst7", "Tyrobp","Cd63", "Clec7a", "H2-D1", "Lyz2", "Npc2")
svg(paste0("./output/WGCNA_", module, "selectedgenes_heatmap_byCluster.svg"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "labels", slot = "data")
dev.off()
```

```{r, fig.width=5, fig.height=5}
module = "brown"
featureslist=c("Trem2", "Csf1r", "C1qa", "C1qb", "C1qc", "Ctss",
               "Ctsl", "Ctsh", "Cyba", "Lamp1")
svg(paste0("./output/WGCNA_", module, "selectedgenes_heatmap_byCluster.svg"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "labels", slot = "data")
dev.off()
```

```{r, fig.width=5, fig.height=5}
module = "turqouise"
featureslist=c("Olfr1307", "Olfr635", "Olfr1336", "Olfr46", 
               "Olfr344", "Fgd4", "Pkib", "Shisa5", "Senp7", "Snhg14")
svg(paste0("./output/WGCNA_", module, "selectedgenes_heatmap_byCluster.svg"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "labels", slot = "data")
dev.off()
```

```{r, fig.width=5, fig.height=5}
module = "grey"
featureslist=c("Adam10", "Abca1", "CD74","Arpc1b", "Eef2",
               "Itgb2", "Pdia3", "Aldoa", "Cfh", "Ssh2")
svg(paste0("./output/WGCNA_", module, "selectedgenes_heatmap_byCluster.svg"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "labels", slot = "data")
dev.off()
```


##### WGCNA GO terms

```{r, out.width='130%',out.height='100%'}


pvallist = unlist(reslist)
MOI = gsub("ME", "", names(pvallist)) %>% gsub(".Clusters", "", .)
# MOI = MOI[pvallist<0.05]

# if you want to have the affected genes set return_genelist = T; this takes roughly 4-8 hours 

Gores = getGOresults(geneset = genelist[MOI], domain_scope = "known",
                     genereference = gene_univers, 
                     return_genelist = T,
                     organism = "mmusculus")

display_tab(Gores$result)
openxlsx::write.xlsx(Gores$result%>% dplyr::select(-parents), file="output/WGCNA_GO_terms.xlsx")
write.csv2(Gores$result %>% dplyr::select(-parents), file="output/WGCNA_GO_terms.csv")

plotlist <- list()
for(m in unique(Gores$result$query)){
  p <- GOplot(Gores$result[Gores$result$query ==m,], N = 10, Title = m)
  plotlist[[m]] <- p
}

```

```{r, fig.width=8, fig.height=5}
p<- ggarrange(plotlist = plotlist)
p 
ggsave(file="output/WGCNA_GO_terms.svg", p)
```


## cell cylce scoring plots 

```{r}
p <- DimPlot(samples.integrated.filt, group.by = "Phase")
p
ggsave(file="output/CClustering_CellPhase.svg", p)

p <- FeaturePlot_scCustom(samples.integrated.filt, features =  "S.Score",  colors_use = viridis_dark_high)
p
ggsave(file="output/CClustering_CellPhaseScore.svg", p)

p <- FeaturePlot_scCustom(samples.integrated.filt, features =  "G2M.Score",  colors_use = viridis_dark_high)
p
ggsave(file="output/CClustering_CellPhaseG2MScore.svg", p)


metadata <- samples.integrated.filt@meta.data

p<- ggplot(metadata, aes(x=labels, y= S.Score))+geom_violin(aes(fill=labels))+geom_boxplot(aes(alpha=1))+theme_classic()
p
ggsave(file="output/CClustering_CellPhaseScoreViolin.svg", p)
p<- ggplot(metadata, aes(x=labels, y= G2M.Score))+geom_violin(aes(fill=labels))+geom_boxplot(aes(alpha=1))+theme_classic()
p
ggsave(file="output/CClustering_CellPhaseG2MScoreViolin.svg", p)

pairwise.t.test(metadata$S.Score,  metadata$labels)
pairwise.t.test(metadata$G2M.Score,  metadata$labels)


```

## Rdigplot custom genes BAM CAM

BAM = "Cd36", "CD38","Lyve1","Cd206","CD163", "Cd169"
CAM = "Emilin2", "PF4", "Ms4a", "Hp","F5","Mki67"

```{r, fig.width=12, fig.height=5}


DefaultAssay(samples.integrated.filt) <- "RNA"

BAM = c("Cd36", "Cd38","Lyve1","Cd206","Cd163", "Cd169")
CAM = c("Emilin2", "Pf4", "Msa4a", "Hp","F5","Mki67")

# Cd206, Cd169 not found 

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, assay = "RNA", layer="counts",
                                          features = list(BAM), name = "BAM_score")

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, assay = "RNA", layer="counts",
                                          features = list(CAM), name = "CAM_score")

p <- RidgePlot(samples.integrated.filt, features = c("BAM_score1", "CAM_score1") , group.by = "labels")

p
  
ggsave("/files/scSeq_Hefendehl/output/Ridge_Custom_Genes.svg", p)

```
#QC plots final set 
```{r fig.width=12, fig.height=4}
DefaultAssay(samples.integrated.filt) <- "RNA"

metadata <- samples.integrated.filt@meta.data

metadata$nCount_RNA_finset <- colSums(samples.integrated.filt)
metadata$nFeature_RNA_finset <- colSums(samples.integrated.filt@assays$RNA@counts>0)

metadata -> samples.integrated.filt@meta.data
p <- VlnPlot(samples.integrated.filt, features = c("nCount_RNA_finset") , group.by = "labels") + ylim(c(0,4000)) +geom_hline(yintercept = 450)
p
p2 <- VlnPlot(samples.integrated.filt, features = c("nFeature_RNA_finset")  , group.by = "labels") + ylim(c(0,4000)) +geom_hline(yintercept = 150)
p2

p3 <- VlnPlot(samples.integrated.filt, features = "percent.mt" , group.by = "labels") + ylim(c(0,10)) +geom_hline(yintercept = 7)
p3


pcomb <- ggarrange(p,p2,p3, common.legend = T, ncol = 3, legend = "right")
pcomb
ggsave("/files/scSeq_Hefendehl/output/Qualityplots.svg", pcomb)


```



```{r}
saveRDS(samples.integrated.filt, "/files/scSeq_Hefendehl/output/SeuratObjectafterProcessing.rds")




```


