---
title: "Analyses for Publication"
author: "AGC"
date: "2023-09-27"
output:
  #pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}

#options(Seurat.object.assay.version = "v5")

outputdir="./output/Res_202602/"
if(!dir.exists(outputdir)) dir.create(outputdir, recursive = T)

## don't update matrixStats
knitr::opts_chunk$set(  
  echo    = FALSE,
  warning = FALSE,
  message = FALSE)

#library(limma)
#library(DESeq2)
library(tidyverse)
library(scCustomize)
library(Seurat)
library(harmony)
library(SingleR)
library(data.table)
library(compareGroups)
library(openxlsx)
library(viridis)
library(DT)
library(slingshot)
library(SingleCellExperiment)
library(ggpubr)
library(clustree)
library(RCurl)
library(AnnotationHub)
library(WGCNA)
library(plotly)
library(org.Mm.eg.db)
library(knitr)
library(ggrepel)

set.seed(1324567)
getwd()

```


```{r load_annotations, cache=F, include=FALSE}
source ("./code/custom_functions.R")

cc_file <- getURL("https://raw.githubusercontent.com/hbc/tinyatlas/master/cell_cycle/Mus_musculus.csv") 
cell_cycle_genes <- read.csv(text = cc_file)

ah <- AnnotationHub::AnnotationHub()

# Access the Ensembl database for organism
ahDb <- query(ah, 
              pattern = c("Mus musculus", "EnsDb"), 
              ignore.case = TRUE)

# Acquire the latest annotation files
id <- ahDb %>%
  mcols() %>%
  rownames() %>%
  tail(n = 1)

# Download the appropriate Ensembldb database
edb <- ah[[id]]

# Extract gene-level information from database
annotations <- genes(edb, 
                     return.type = "data.frame")

# Select annotations of interest
annotations <- annotations %>%
  dplyr::select(gene_id, gene_name, seq_name, gene_biotype, description)
```


# Load batch 2026 
## Metadata additional samples 


```{r load_add_meta, cache=F}
# Read Sheet2
df_raw <- read.xlsx("/files/scSeq_Hefendehl/data/20260126/534_plates_wSexAge.xlsx", sheet = "Sheet2")

# Clean column names
colnames(df_raw) <- c("User_Sample_Name", "Plate", "Container_Type", 
                      "Well_Location", "Protocol", "Species", "Mouse_ID",
                      "Tissue_or_Organ", "Cell_Type", "Genotype", "Treatment", 
                      "Gender", "Age")

# Filter to actual sample entries (remove header rows / template rows)
df <- df_raw %>%
  dplyr::filter(!is.na(User_Sample_Name), 
                !grepl("^<", User_Sample_Name),   # remove <TABLE HEADER> etc.
                Container_Type == "386 well plate") # keep only plate entries

# Function to expand a well range like "A1 - P12" into all individual wells
expand_wells <- function(well_range) {
  # Parse "A1 - P12" format
  parts <- trimws(strsplit(well_range, "-")[[1]])
  start <- trimws(parts[1])
  end   <- trimws(parts[2])
  
  # Extract row letters and column numbers
  start_row <- substr(start, 1, 1)
  start_col <- as.integer(substring(start, 2))
  end_row   <- substr(end, 1, 1)
  end_col   <- as.integer(substring(end, 2))
  
  rows <- LETTERS[which(LETTERS == start_row):which(LETTERS == end_row)]
  cols <- start_col:end_col
  
  # Generate all combinations
  expand.grid(Row = rows, Col = cols, stringsAsFactors = FALSE) %>%
    mutate(Well = paste0(Row, Col)) %>%
    pull(Well)
}

# Expand each sample entry to individual well positions
df_expanded <- df %>%
  rowwise() %>%
  mutate(Well = list(expand_wells(Well_Location))) %>%
  unnest(Well) %>%
  dplyr::select(Plate,
                Well,
                Species,
                Mouse_ID,
                Tissue_or_Organ,
                Cell_Type,
                Genotype,
                Treatment, 
                Age, Gender)

df_expanded <- df_expanded %>% 
  mutate(
    Plate = gsub("_011", "_11", Plate), 
    Plate = gsub("_012", "_12", Plate), 
    Cell_ID = paste0(Plate, "_", Well), 
    Gender = ifelse(Gender=="female", "f", "m"),
    Genotype_Treatment = paste0(Genotype, "_", Treatment), 
  ) %>% 
  dplyr::select(
    Plate,
    Cell_ID,
    Mouse_ID,
    Genotype,
    Treatment,
    Genotype_Treatment,
    Age,
    Gender) %>% as.data.frame()

rownames(df_expanded) <- df_expanded$Cell_ID
# View result
display_tab(df_expanded)

# Optional: save to CSV
write.csv(df_expanded, 
          "/files/scSeq_Hefendehl/data/20260126/534_Plate_positions_all.csv", 
          row.names = FALSE)

```




## load additional samples second GT 

```{r load_add_samples 2nd, cache=F, dependson="load_add_meta"}

files <- list.files("/files/scSeq_Hefendehl/data/20260126/", 
                    pattern = ".*Kallisto.*.csv", 
                    full.names = T)

countlist = list()
for(f in files){
  counts <- read.csv(f, row.names = 1, header = TRUE)
  matched <- annotations$gene_name[match(
    gsub("\\.[0-9]+", "", rownames(counts)), 
    annotations$gene_id)]
  gene_names <- ifelse(is.na(matched), rownames(counts), matched)
  dups <- duplicated(gene_names) | duplicated(gene_names, fromLast = TRUE)
  dup_counts <- counts[dups, ] %>%
    as.data.frame() %>%
    mutate(gene = matched[dups]) %>%
    group_by(gene) %>%
    summarise(across(everything(), sum)) %>%
    column_to_rownames("gene")
  
  rownames(counts)[!dups] <- gene_names[!dups]
  counts <- rbind(counts[!dups, ], dup_counts)
  
  counts <- counts[rownames(counts) != "", ]
  counts <- as((as.matrix(counts)), "sparseMatrix")
  
  meta_counts <- df_expanded[colnames(counts), ]
  
  counts <- CreateSeuratObject(counts = counts, 
                               project = unique(meta_counts$Plate), 
                               min.cells = 3, 
                               min.features = 150, 
                               meta.data =meta_counts)
  Notannotgenes <- grepl("^ENSMUSG", rownames(counts))
  counts <- counts[!Notannotgenes,]
  
  print(unique(counts$Plate))
  counts[["orig.ident"]] <- counts$Cell_ID
  countlist[[unique(counts$Plate)]] <- counts
}
```

# Load batch 2025

```{r load_add_samples, cache=F, dependson="load_add_samples 2nd"}

Sample_files <- list.files("/files/scSeq_Hefendehl/data/20230203/Sample_Tables/", 
                           pattern = "Sample.*.csv", 
                           full.names = T)

counts_run1 <- read.csv("/files/scSeq_Hefendehl/data/20230203/Alignments/Run1/Kallisto/counts.csv", row.names = 1, header = TRUE) %>% rownames_to_column(var="GENE")
counts_run2 <- read.csv("/files/scSeq_Hefendehl/data/20230203/Alignments/Run2/Kallisto/counts.csv", row.names = 1, header = TRUE) %>% rownames_to_column(var="GENE")

counts <- counts_run1 %>% full_join(counts_run2, by = "GENE")


Platenames= c("025"="SS2_20_025_9", 
              "022"="SS2_20_022_7", 
              "024"="SS2_20_024_8",
              "048"="SS2_21_048_6")


Samples_meta = data.frame()
for(s in Sample_files){
  P = substr(basename(s), 14,16)
  P = Platenames[P]
  Samples <- read.csv(s, header = TRUE) %>% 
    mutate(
      Plate=as.character(P), 
      Genotype = ifelse(Genotype=="WT", "WT", "APPPS1"), 
      Genotype_Treatment = paste0(Genotype, "_", Treatment),
      Cell_ID=paste0(P,"_", Cell_ID), 
      Age = as.numeric(gsub(" weeks", "", Age))) %>% 
    dplyr::select(
      "Plate",
      "Cell_ID",
      "Mouse_ID",
      "Genotype",
      "Treatment",
      "Genotype_Treatment",
      "Gender",
      "Age")
  Samples_meta <- rbind(Samples_meta, Samples)
}

counts_sel <- counts %>% dplyr::select(any_of(Samples_meta$Cell_ID), "GENE") %>% as.data.frame() %>% column_to_rownames("GENE")
counts_sel[is.na(counts_sel)]<-0

matched <- annotations$gene_name[match(
  gsub("\\.[0-9]+", "", rownames(counts_sel)), 
  annotations$gene_id)]
gene_names <- ifelse(is.na(matched), rownames(counts_sel), matched)
dups <- duplicated(gene_names) | duplicated(gene_names, fromLast = TRUE)
dup_counts <- counts_sel[dups, ] %>%
  as.data.frame() %>%
  mutate(gene = matched[dups]) %>%
  group_by(gene) %>%
  summarise(across(everything(), sum)) %>%
  column_to_rownames("gene")

rownames(counts_sel)[!dups] <- gene_names[!dups]
counts_sel <- rbind(counts_sel[!dups, ], dup_counts)

counts_sel <- counts_sel[rownames(counts_sel) != "", ]
counts_sel <- as((as.matrix(counts_sel)), "sparseMatrix")

Cells <- intersect(colnames(counts_sel), Samples_meta$Cell_ID)
counts_sel <- counts_sel[,Cells]

rownames(Samples_meta) <- Samples_meta$Cell_ID
Samples_meta <- Samples_meta[Cells,]

counts <- CreateSeuratObject(counts = counts_sel, 
                             project = "Batch1", 
                             min.cells = 3, 
                             min.features = 150, 
                             meta.data =Samples_meta)
Notannotgenes <- grepl("^ENSMUSG", rownames(counts))
counts <- counts[!Notannotgenes,]
counts_list_firstbatch <- SplitObject(counts, split.by = "Plate")


```


```{r}
seurat_list = c(counts_list_firstbatch, countlist)

seurat_obj <- merge(
  seurat_list[[1]],
  y = seurat_list[-1],
  add.cell.ids = names(seurat_list))

# Exclusde not gene_name annotated genes 



```


```{r}
rm(list=c("meta_counts",
          "samples.integrated", 
          "seurat_list", 
          "tmp_counts", 
          "tmp_meta", 
          "tmp_seurat", 
          "ah", 
          "ahDb", 
          "countlist", 
          "counts", 
          "counts_list_firstbatch", 
          "edb",
          "df", "df_expanded", 
          "df_raw", 
          "dup_counts"))

gc()
```


```{r, fig.width=18, fig.height=5}
Idents(seurat_obj) <- "Plate"
DefaultAssay(seurat_obj) <- "RNA"

seurat_obj[["percent.mt"]] <- PercentageFeatureSet(seurat_obj, pattern = "^mt-")
seurat_obj <- JoinLayers(seurat_obj)
seurat_obj[["RNA"]] <- JoinLayers(seurat_obj[["RNA"]])

seurat_obj <- seurat_obj %>% 
  NormalizeData()  %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA() %>% 
  FindNeighbors(dims = 1:30, reduction = "pca") %>% 
  FindClusters(cluster.name = "unintegrated_clusters") %>% 
  RunUMAP(dims = 1:30, reduction = "pca", reduction.name = "umap.unintegrated")

p <- DimPlot(seurat_obj, reduction = "umap.unintegrated", group.by = c("Plate", "Mouse_ID", "unintegrated_clusters"))
p

ggsave(paste0(outputdir, "DataLoad_umap.unintegrated.pdf"), p, width = 18, height=5)
```


```{r fig.width=12, fig.height=4}
VlnPlot(seurat_obj, features = c("nCount_RNA", "nFeature_RNA", "percent.mt"), 
        group.by = "Plate")

seurat.obj.filt <- subset(seurat_obj, 
                          subset = nFeature_RNA > 100 & 
                            nFeature_RNA < 7000 & 
                            nCount_RNA <3e6 &
                            nCount_RNA > 400 &
                            percent.mt < 5)

counts <- JoinLayers(seurat.obj.filt, assay = "RNA") %>%    GetAssayData(assay = "RNA", layer="data") 
genes.percent.expression <- rowMeans(counts > 0)
keep.genes <- names(genes.percent.expression[genes.percent.expression >= 0.005])

# filter for expression in at least 0.5% of cells

seurat.obj.filt <- seurat.obj.filt[keep.genes, ]

VlnPlot(seurat.obj.filt, features = c("nCount_RNA", "nFeature_RNA", "percent.mt"), 
        group.by = "Plate")

# 7000 or less than 150 and cells with >7%  mitochondrial counts.

samples.integrated <- seurat.obj.filt

```



```{r}
#samples.integrated = readRDS("/files/scSeq_Hefendehl/data/20230203/seu-SCT-harmony_02-02-23.rds")

samples.integrated[["Genotype_Treatment"]] = factor(samples.integrated$Genotype_Treatment, 
                                                    levels = c("WT_Ctrl", 
                                                               "WT_Stroke", 
                                                               "APPPS1_Ctrl", 
                                                               "APPPS1_Stroke"))


samples.integrated@meta.data %>% 
  dplyr::select(Mouse_ID, Genotype_Treatment, Gender, Age) %>% 
  group_by(Mouse_ID) %>% mutate(
    nCells = n()
  ) %>% ungroup() %>% 
  dplyr::distinct() -> Samplelist

openxlsx::write.xlsx(Samplelist, "/files/scSeq_Hefendehl/data/Samplelist.xlsx",rowNames = TRUE)

display_tab(Samplelist)

```


## Reclustering of datasets

### Define Gene lists

```{r}

DAM_up <- c("Itgax","Cst7","Ccl4","Clec7a","Lpl","Siglec1","Spp1","Axl","Apoe","Trem2","Csf1","Lyz2","H2-D1","Tyrobp","Cd74","Gpnmb","B2m","Cd9","Ctsb","Ctsd") # genes up regulated in DAMs versus homestatics

# additions
DAM1_up <- c("Tyrobp", "Ctsb", "Ctsd", "Apoe", "B2m", "Fth1", "Lyz2") # genes up regulated in DAM1 vs Homeostasis
DAM1_down <- c("Cx3cr1", "P2ry12", "Tmem119")# genes down regulated in DAM1 vs homeostasis
DAM2_up <- c("Trem2", "Axl", "Cst7", "Ctsl", "Lpl", "Cd9", "Csf1", "Ccl6", "Itgax", "Clec7a", "Lilrb4", "Timp2")

DAM1vsDAM2 = c("Trem2", "Cst7", "Spp1", "Lpl", "Itgax", "Ctsl", "Ctsz", "Cd68", "Axl", "Cd9", "Ccl6", "Csf1")

homeostat.genes=c("P2ry12","P2ry13","Tmem119","Cx3cr1","Selplg","Cd33")


INF_resp_microglia <- c("Ifit2", "Ifit3", "Ifitm3", "Irf7", "Oasl2")
Cycling_microglia <- c("Top2a", "Mcm2", "Tubb5", "Mki67",  "Cdk1")
Act_resp_microglia <- c("Cd74", "H2-Ab1", "H2-Aa", "Ctsb",  "Ctsd")
Axon_tract_microglia <- c("Spp1", "Gpnmb", "Igf1", "Lgals3", "Fabp5", "Lpl", "Lgals1", "Ctsl", "Anxa5")


cluster.genes=unique(c(DAM_up, DAM1vsDAM2, homeostat.genes))
```

```{r}
ngenes = 50
distal_dat = openxlsx::read.xlsx(
  "/files/scSeq_Hefendehl/data/Copy of DEG_control_vs_stroke_vs_distal_original.xlsx", 
  sheet = 1)
distal_dat <- distal_dat[order(distal_dat$score, decreasing = T)[1:ngenes],]
distal_genes <- distal_dat$X1

control_dat = openxlsx::read.xlsx("/files/scSeq_Hefendehl/data/Copy of DEG_control_vs_stroke_vs_distal_original.xlsx",sheet = 2)
control_dat <- control_dat[order(control_dat$score, decreasing = T)[1:ngenes],]
control_genes <- control_dat$X1

Stroke_dat = openxlsx::read.xlsx("/files/scSeq_Hefendehl/data/Copy of DEG_control_vs_stroke_vs_distal_original.xlsx",sheet = 3)
Stroke_dat <- Stroke_dat[order(Stroke_dat$score, decreasing = T)[1:ngenes],]
Stroke_genes <- Stroke_dat$X1
```

## inspect data

### SCT CCA integration all cells

```{r, fig.width=8,fig.height=4}

DefaultAssay(samples.integrated) = "RNA"
samples.integrated <- UpdateSeuratObject(samples.integrated)

samples.integrated$Run <-substr(samples.integrated$Plate, 1, 6)

samples.integrated[["RNA"]] <- split(samples.integrated[["RNA"]], f= samples.integrated$Run)
samples.integrated <- SCTransform(samples.integrated, vst.flavor = "v2")
samples.integrated <- RunPCA(samples.integrated, npcs = 30, verbose = FALSE)

# one-liner to run Integration
DefaultAssay(samples.integrated) <- "RNA"
samples.integrated <- IntegrateLayers(object = samples.integrated, method = CCAIntegration,
                                      orig.reduction = "pca", new.reduction = 'integrated.cca')

samples.integrated <- FindNeighbors(samples.integrated, reduction = "integrated.cca", dims = 1:30)
samples.integrated <- FindClusters(samples.integrated, resolution = 0.6)
samples.integrated <- RunUMAP(samples.integrated, reduction = "integrated.cca", dims = 1:30, reduction.name = "umap.integrated.cca")

a = DimPlot(samples.integrated, reduction = "umap.integrated.cca", group.by = "Plate")
b = FeaturePlot_scCustom(samples.integrated, reduction = "umap.integrated.cca", features = "Aif1", colors_use = viridis_dark_high)
b2 = FeaturePlot_scCustom(samples.integrated, reduction = "umap.integrated.cca", features = "Ptprc", colors_use = viridis_dark_high)
c = RidgePlot(samples.integrated, features = "Aif1", group.by = "Mouse_ID")
c2 = RidgePlot(samples.integrated, features = "Ptprc", group.by = "Mouse_ID")

d = DimPlot(samples.integrated, reduction = "umap.integrated.cca", split.by = "Genotype_Treatment", pt.size = 1)
e = FeaturePlot_scCustom(samples.integrated, reduction = "umap.integrated.cca", features = "Aif1", split.by = "Genotype_Treatment", colors_use = viridis_dark_high)
e2 = FeaturePlot_scCustom(samples.integrated, reduction = "umap.integrated.cca", features = "Ptprc", split.by = "Genotype_Treatment", colors_use = viridis_dark_high)
```

```{r, fig.width=12,fig.height=5}
a|b|b2

ggarrange(c,c2, common.legend=T, legend = "none")

ggsave(paste0(outputdir,"cluster_cca_integrated_nofilter.pdf"), a)
```

```{r, fig.width=12,fig.height=4}
d
e
e2

```

###  Sex specific genes 

```{r, fig.width=12,fig.height=4}
Idents(samples.integrated) <- "Mouse_ID"
Sexgenes = c("Ddx3y", #y linked and microglia expressed
             "Xist") #X inactivation

# cluster all cells for visualization purpose prior to exclusion 

a = DimPlot(samples.integrated, reduction = "umap.integrated.cca")
b = FeaturePlot_scCustom(samples.integrated, reduction = "umap.integrated.cca", 
                         features = "Xist", 
                         colors_use = viridis_dark_high)
c = RidgePlot(samples.integrated, features = Sexgenes)

d = DimPlot(samples.integrated, 
            reduction = "umap.integrated.cca", 
            group.by = "Plate", 
            split.by = "Mouse_ID",
            pt.size = 1)
e = FeaturePlot_scCustom(samples.integrated, 
                         reduction = "umap.integrated.cca", 
                         features = Sexgenes,
                         split.by = "Mouse_ID", 
                         colors_use = viridis_dark_high)
```

```{r, fig.width=10,fig.height=4}
a|b
c

ggsave(paste0(outputdir, "cluster_Xist_harmony_integrated_nofilter.pdf"), a|b)
ggsave(paste0(outputdir, "cluster_Xist_Ridge_integrated_nofilter.pdf"), c)
```

```{r, fig.width=30,fig.height=4}
d

```

```{r fig.width=6, fig.height=6}

DimPlot(samples.integrated, reduction="umap.integrated.cca", group.by = "Genotype_Treatment")

```

```{r fig.width=80, fig.height=15}
e
ggsave(paste0(outputdir, "cluster_Sexgenes_integrated_nofilter.pdf"), e, limitsize = F)
```

```{r, include=FALSE}
reads = JoinLayers(samples.integrated, assay = "RNA") %>%    GetAssayData(assay = "RNA", layer="data") 
reads = t(reads) %>% as.data.frame()

metadata <- samples.integrated@meta.data
metadata$Aif1neg = NA
metadata$Aif1neg <- paste0("Aif1neg_",as.character(reads$Aif1 ==0))
metadata -> samples.integrated@meta.data

```

```{r}
rm( list=c("reads", "seurat_obj", "seurat.obj.filt"))
```

### remove sex specific and not-annotated genes

```{r, fig.width=4,fig.height=4}
samples.integrated.filt <- samples.integrated

sex.genes <- rownames(samples.integrated.filt) %in%  annotations$gene_name[grepl("X|Y", annotations$seq_name)]
Gmgenes <- grepl("^Gm", rownames(samples.integrated.filt))
Rikgenes <- grepl(".*Rik$", rownames(samples.integrated.filt))

samples.integrated.filt<- samples.integrated.filt[!(sex.genes|Gmgenes|Rikgenes),]
```

# Clean integration 

## Pre cell filtering integration

```{r, fig.width=4,fig.height=4}
DefaultAssay(samples.integrated.filt)<- "RNA"

samples.integrated.filt <- samples.integrated.filt %>% RunHarmony("Run")
SDs=Stdev(object = samples.integrated.filt, reduction =  "pca")
varexpl <- SDs/sum(SDs)
ndims <- which(cumsum(varexpl)>0.75)[1]


ElbowPlot(samples.integrated.filt, reduction = "pca", ndims = 30) + geom_vline(xintercept =ndims)

samples.integrated.filt <- samples.integrated.filt %>% 
  FindNeighbors(reduction = "integrated.cca", dims = 1:ndims) %>% 
  FindClusters(cluster.name = "integrated_clusters") %>% 
  RunUMAP(dims = 1:ndims, reduction = "integrated.cca")
```


```{r, fig.width=12,fig.height=6}
DimPlot(samples.integrated.filt, reduction = "umap", 
        group.by = c("Plate", "Mouse_ID", "integrated_clusters", "Genotype_Treatment"))

```


## map cell types

```{r}
ref.sce.mm <- celldex::MouseRNAseqData()
sce.filt<- as.SingleCellExperiment(JoinLayers(samples.integrated.filt, assay = "RNA"), assay = "RNA")

pred.mouseRNA <- SingleR(test =sce.filt, ref = ref.sce.mm, assay.type.test=1,
                         labels = ref.sce.mm$label.fine)

table(pred.mouseRNA$labels) #fine tuned cell types

#table(pred.mouseRNA$pruned.labels) #cell type after pruning

samples.integrated.filt$Immgen_sc_labels_agc <- NA
samples.integrated.filt$Immgen_sc_labels_agc <- pred.mouseRNA$labels
```


```{r}
pheat=plotScoreHeatmap(pred.mouseRNA, show.labels = TRUE,
                       annotation_col=data.frame(cluster=samples.integrated.filt$seurat_clusters,
                                                 
                                                 row.names=rownames(pred.mouseRNA)))

Idents(samples.integrated.filt)<- "seurat.integrated"

a<- DimPlot(samples.integrated.filt, reduction="umap", group.by="Immgen_sc_labels_agc", 
            label = F)
b<- DimPlot(samples.integrated.filt, reduction="umap", group.by="Immgen_sc_labels_agc", 
            label = F, split.by = "Plate")
c<- DimPlot(samples.integrated.filt, reduction="umap", group.by="Immgen_sc_labels_agc", split.by = "Genotype_Treatment",
            label = F)
```

```{r}
print(pheat)

ggsave(filename = paste0(outputdir, "CelltypesHeatmap_Opt_All_genes_all_cells_Clusters_integrated.pdf"), pheat)
```


```{r}
a
ggsave(filename = paste0(outputdir, "Celltypes_Opt_All_genes_all_cells_Clusters_integrated.pdf"), a)
```


```{r, fig.width=18, fig.height=4}
b

ggsave(filename = paste0(outputdir, "Celltypes_Opt_All_genes_all_cells_Clusters_byPlate_integrated.pdf"), b)
```


```{r, fig.width=12, fig.height=4}
c

ggsave(filename = paste0(outputdir, "Celltypes_Opt_All_genes_all_cells_Clusters_byCondition_integrated.pdf"), c)
```

## filter and recluster only Microglia

### Re-optimize clustering

```{r}
samples.integrated.filt <-  samples.integrated.filt[,grepl("Microglia",samples.integrated.filt$Immgen_sc_labels_agc)]


DefaultAssay(samples.integrated.filt)<- "RNA"
samples.integrated.filt <- samples.integrated.filt %>% 
  NormalizeData() %>% 
  ScaleData() %>% 
  RunPCA() %>% 
  RunHarmony("Run")
```

#### Layer integration across Runs using CCA Integration 

```{r}

reslist= seq(0.2,1.2,0.2)

samples.integrated.filt <- samples.integrated.filt %>% 
  IntegrateLayers(method = CCAIntegration,
                  orig.reduction = "pca", 
                  new.reduction = 'integrated.cca') %>% 
  FindNeighbors(reduction = "integrated.cca", dims = 1:ndims) %>% 
  FindClusters(resolution = reslist) %>% 
  RunUMAP(reduction = "integrated.cca", min.dist = 0.1, spread = 1, dims=1:ndims)

clustreeplot<-clustree(samples.integrated.filt, prefix = "RNA_snn_res.")


samples.integrated.filt <- samples.integrated.filt %>% RunUMAP(reduction = "integrated.cca", min.dist = 0.1, spread = 1, dims=1:ndims, reduction.name = "umap")

```

### Define resoultion for clustering 

```{r, fig.width=15, fig.height=12}
plotlist=list()
for( cl in (grep("RNA_snn", colnames(samples.integrated.filt@meta.data), value=T))){
  plotlist[[cl]] <- DimPlot(samples.integrated.filt, reduction="umap", group.by = cl)
}

plotlist <- plotlist[sort(names(plotlist))]

ggarrange(plotlist=plotlist)

```


```{r, fig.width=5, fig.height=6}
clustreeplot
ggsave(filename = paste0(outputdir, "All_genes_Clustertree_integrated.pdf"), clustreeplot)

```

```{r}
library(mclust)

samples.integrated.filt.meta <- samples.integrated.filt@meta.data

res_cols <- sort(grep("RNA_snn_res", colnames(samples.integrated.filt.meta), value = TRUE))

ari_results <- data.frame()

for (i in 1:(length(res_cols) - 1)) {
  col1 <- res_cols[i]
  col2 <- res_cols[i + 1]
  ari <- adjustedRandIndex(samples.integrated.filt.meta[[col1]], 
                           samples.integrated.filt.meta[[col2]])
  ari_results <- rbind(ari_results, data.frame(res1 = col1, res2 = col2, ARI = ari))
}

best_res <- ari_results$res1[which.max(ari_results$ARI)]
best_res <- as.numeric(gsub("RNA_snn_res.", "", best_res))
ari_results
best_res <- 0.6
```

clustering resolultion was set to `r best_res`

```{r, warning=F, }

DefaultAssay(samples.integrated.filt) <- "RNA"

samples.integrated.filt <- samples.integrated.filt %>% 
  FindNeighbors(reduction = "integrated.cca", dims = 1:ndims) %>% 
  FindClusters(resolution = best_res, cluster.name = "seurat.integrated") %>% 
  RunUMAP(reduction = "integrated.cca", min.dist = 0.1, spread = 1, dims=1:ndims)

n = length(unique(samples.integrated.filt$seurat.integrated))

Idents(samples.integrated.filt) <- "seurat.integrated"
samples.integrated.filt$labels<- factor(Idents(samples.integrated.filt), 
                                        levels = c(0:(n-1)), labels=paste0("Microglia", 0:(n-1))
)

Idents(samples.integrated.filt) = "labels"
DefaultAssay(samples.integrated.filt) <- "RNA"
samples.integrated.filt[["RNA"]] <- samples.integrated.filt[["RNA"]] %>% JoinLayers() 
samples.integrated.filt[["RNA"]]$scale.data <- NULL

samples.integrated.filt<- samples.integrated.filt %>% NormalizeData() %>%  ScaleData()


a<-DimPlot(samples.integrated.filt, , reduction = "umap",
           group.by = "seurat.integrated", pt.size =1, label = T) + NoLegend()
b<-DimPlot(samples.integrated.filt, reduction = "umap", split.by = "Plate", pt.size = 1)
c<-DimPlot(samples.integrated.filt, reduction = "umap", split.by = "Genotype_Treatment", pt.size =1)
d <- DimPlot(samples.integrated.filt, reduction = "umap", split.by = "Mouse_ID", pt.size =1)
e<-FeaturePlot_scCustom(samples.integrated.filt,reduction = "umap", layer = "scale.data",
                        features = homeostat.genes, colors_use = viridis_dark_high)

```


```{r, fig.width=5, fig.height=5}

a
ggsave(filename = 
         paste0(outputdir,
                "Opt_All_genes_all_cells_Clusters_integrated.pdf"), a)
```


```{r, fig.width=30, fig.height=5}
b

ggsave(filename = 
         paste0(outputdir,
                "Opt_All_genes_all_cells_Clusters_byPlate_integrated.pdf"),
       b)


```


```{r, fig.width=12, fig.height=4}
c
ggsave(filename = paste0(
  outputdir,
  "Opt_All_genes_all_cells_Clusters_byCondition_integrated.pdf"),
  c)
```


```{r, fig.width=30, fig.height=3}
d
ggsave(filename = paste0(
  outputdir,
  "Opt_All_genes_all_cells_Clusters_byMouseID_integrated.pdf"), 
  d)
```

```{r, fig.width=8, fig.height=12}
e
ggsave(filename = paste0(outputdir,
                         "All_genes_Clusters_byHomeostatgenes_integrated.pdf"), e)
```

```{r}
a <- DimPlot(samples.integrated.filt, reduction="umap", label = T)
b <- DimPlot(samples.integrated.filt, reduction="umap", split.by = "Plate", label = F)
c <- DimPlot(samples.integrated.filt, reduction="umap", split.by = "Genotype_Treatment", label = F)
d <- DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Genotype_Treatment")
e <- DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Gender")
g <- DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Aif1neg")
h <- DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Immgen_sc_labels_agc")

samples.integrated.filt[["Immgen_sc_label_short"]] <- ifelse(
  grepl("Microglia", samples.integrated.filt$Immgen_sc_labels_agc), 
  "Microglia", NA)

i <- DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Immgen_sc_label_short")


```


```{r, fig.width=8, fig.height=7}
a
ggsave(filename = paste0(outputdir, "Microglia_Opt_All_genes_Clusters_integrated.pdf"), a)


d
ggsave(filename = paste0(outputdir, "Microglia_Opt_All_genes_Condition_integrated.pdf"), d)

e
ggsave(filename = paste0(outputdir, "Microglia_Opt_All_genes_Gender_integrated.pdf"), e)

g
ggsave(filename = paste0(outputdir, "Microglia_Opt_All_genes_Aif1neg_integrated.pdf"), g)

h
ggsave(filename = paste0(outputdir, "Microglia_Opt_All_genes_Immgen_sc_labels_agc_integrated.pdf"), h)

i
ggsave(filename = paste0(outputdir, "Microglia_Opt_All_genes_Immgen_sc_labels_summed_integrated.pdf"), i)

```


```{r, fig.width=12, fig.height=4}
c
ggsave(filename = paste0(outputdir, "Microglia_Opt_All_genes_Clusters_byCondition.pdf"), c)
```

```{r, fig.width=18, fig.height=3}
b
ggsave(filename = paste0(outputdir, "Microglia_Opt_All_genes_Clusters_byPlate_integrated.pdf"), b)
```


```{r, fig.height=18, fig.width=10}
f <- FeaturePlot_scCustom(samples.integrated.filt, reduction ="umap",
                          features = sort(cluster.genes),order =T, 
                          colors_use = viridis_dark_high)

f
ggsave(filename = paste0(outputdir, "Microglia_Opt_all_genes_MicrogliaGenesintegrated.pdf"), f)

```

```{r, fig.width=8, fig.height=8}
genelist <- c("Alpk1", "Mmp12", "Igf1", "S100a6", "Apoc4", "Dock10", "Nanog", "Olfr344")
f <- FeaturePlot_scCustom(samples.integrated.filt, reduction = "umap",
                          layer = "scale.data", 
                          features = sort(genelist),order =T, 
                          colors_use = viridis_dark_high)

f
ggsave(filename = paste0(outputdir, "Microglia_Opt_all_genes_AddGenelistintegrated.pdf"), f)



```

# Map initial clusters batch 1

```{r, fig.width=8, fig.height=7}
samples.integrated.filt_batch1 <- readRDS("./output/SeuratObjectafterProcessing.rds")

meta <- samples.integrated.filt@meta.data
meta_B1 <- samples.integrated.filt_batch1@meta.data

lkup <- c(SS2_20_22="SS2_20_022", 
          SS2_20_24="SS2_20_024",
          SS2_20_25="SS2_20_025",
          SS2_21_48="SS2_21_048")

meta_B1$plate_new <- lkup[meta_B1$Plate]


meta$label_B1<- meta_B1[match(meta$Cell_ID, paste0(meta_B1$plate_new, "_",meta_B1$Cell_ID)), "labels"]

samples.integrated.filt@meta.data <- meta
p<- DimPlot(samples.integrated.filt, 
            group.by = "label_B1", 
            reduction = "umap", order = T) + scale_color_discrete(na.value = "#F1F1F1")

ggsave(paste0(outputdir, "Overlay_Initital_clusters.pdf"), p)
p
table(samples.integrated.filt$labels, samples.integrated.filt$label_B1)

ggsave(filename = paste0(outputdir, "Microglia_Opt_all_genes_Batch1_Old_labels_integrated.pdf"), p)
rm("samples.integrated.filt_batch1")
```

```{r, fig.width=7, fig.height=7}

DefaultAssay(samples.integrated.filt) <- "RNA"
FeaturePlot_scCustom(samples.integrated.filt, reduction="umap", features = "P2ry1", colors_use = viridis_dark_high, layer = "scale.data")

```


# Markers for Clusters 

```{r}
metadata<-samples.integrated.filt@meta.data

res = compareGroups(Genotype_Treatment~labels, data = metadata)

compareGroups::createTable(res, show.p.mul = T)

compareGroups::createTable(res, show.p.mul = T) %>% export2xls(paste0(outputdir, "Cluster_distribution_genotype.xlsx")) 


res = compareGroups(Genotype_Treatment~Mouse_ID, data = metadata, max.xlev = 70, max.ylev = 20)

compareGroups::createTable(res, show.p.mul = T)

compareGroups::createTable(res, show.p.mul = T) %>% export2xls(paste0(outputdir, "Cellcount_distribution_MouseBygenotype.xlsx"))

res = compareGroups(labels~Mouse_ID, data = metadata, max.xlev = 70, max.ylev = 20)

compareGroups::createTable(res, show.p.mul = T)

compareGroups::createTable(res, show.p.mul = T) %>% export2xls(paste0(outputdir, "Cellcount_distribution_MouseByCluster.xlsx"))


```


```{r, fig.height=18, fig.width=24}
DefaultAssay(samples.integrated.filt) <- "RNA"

f <- RidgePlot(samples.integrated.filt, 
               features = sort(cluster.genes),
               assay = "RNA",
               layer = "data", combine = F)


f <- ggarrange(plotlist = f, common.legend = T, legend = "right")
f
ggsave(filename = paste0(outputdir, "Microglia_Opt_all_genes_MicrogliaGenesRidgePlotintegrated.pdf"), f)

```



### all clusters

```{r}
Idents(samples.integrated.filt) <- "labels"
DefaultAssay(samples.integrated.filt)<- "RNA"
#samples.integrated.filt <- PrepSCTFindMarkers(samples.integrated.filt)

# remove not annotated genes 
geneunivers<-rownames(samples.integrated.filt)

Markers=FindAllMarkers(samples.integrated.filt, 
                       logfc.threshold = 1,
                       assay = "RNA", slot = "data",
                       features = geneunivers, 
                       only.pos = T)

display_tab(Markers)
openxlsx::write.xlsx(Markers, paste0(outputdir, "Clusters_Markers_all.xlsx"),rowNames = TRUE)

genelist = list()
for ( i in unique(Markers$cluster))  genelist[[i]] <- Markers$gene[Markers$cluster==i]

GOterm=getGOresults(genelist, genereference = geneunivers, organism = "mmusculus")
display_tab(GOterm$result)
openxlsx::write.xlsx(GOterm$result, paste0(outputdir, "Clusters_GOterms_Markers_all.xlsx"),rowNames = TRUE)

```



```{r}
metadata=samples.integrated.filt@meta.data
restab = table(GT_Treat = metadata$Genotype_Treatment, DAMs=metadata$labels)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot <- ggplot(as.data.frame(restab), 
                  aes(fill=DAMs, x=GT_Treat, y=Freq))+
  geom_bar(position="fill", stat="identity", color="black", linewidth=0.5)+theme_classic()+ggtitle("Micorglia cluster proportions")

proplot

ggsave(filename = paste0(outputdir, "Microglia_Clusterproportions_per_condition_all_integrated.pdf"), proplot)

library(chisq.posthoc.test)
chisq.posthoc.test(restab)

#To Do Pairwise 

```


```{r, fig.width=10, fig.height=12}
plotlist = list()

mat <- GetAssayData(samples.integrated.filt, assay = "RNA", slot = "data")
temp_obj <- samples.integrated.filt


for (m in names(genelist)){
  #reads = rowSums(samples.integrated.filt@assays$RNA$counts)
  
  targets <- Markers[Markers$cluster==m, ] %>% arrange(desc(pct.1))
  targets <- targets$gene[1:10]
  mat_sel <- mat[targets, , drop = FALSE]
  
  mat_norm <- t(apply(mat_sel, 1, function(x) (x - min(x)) / (max(x) - min(x) + 1e-9))) %>% as.matrix()
  
  temp_obj@assays$RNA@layers$data[match(targets, rownames(temp_obj)), ] <- mat_norm[targets, colnames(temp_obj)]
  
  p<-DoHeatmap(temp_obj, features = targets, disp.max = 1,
               group.by = "labels",  assay = "RNA", slot = "data",
               label = F, combine=F) 
  p <- p[[1]] + scale_fill_viridis()
  ggsave(p, file=paste0(outputdir, "Top10GenesByExpr_",m,"_Heatmaps_byCluster_integrated.pdf"))
  plotlist[[m]]<-p
}

p<- ggarrange(plotlist = plotlist, labels = names(genelist), common.legend = T)
p
ggsave(file=paste0(outputdir, "Top10Genes_Heatmaps_byCluster_Row_Scaled_integrated.pdf"), p)
```

```{r, fig.width=15, fig.height=12}
plotlist = list()

mat <- GetAssayData(samples.integrated.filt, assay = "RNA", slot = "data")

for (m in names(genelist)){
  #reads = rowSums(samples.integrated.filt@assays$RNA$counts)
  
  targets <- Markers[Markers$cluster==m, ] %>% arrange(desc(pct.1))
  targets <- targets$gene[1:10]
  mat_sel <- mat[targets, , drop = FALSE]
  max_plot <- mean(mat_sel)+2*sd(mat_sel)
  
  p<-DoHeatmap(samples.integrated.filt, features = targets, disp.max = ceiling(max_plot),
               group.by = "labels",  assay = "RNA", slot = "data",
               label = F, combine=F) 
  p <- p[[1]] + scale_fill_viridis()
  ggsave(p, file=paste0(outputdir, "Top10GenesByExpr_",m,"_Heatmaps_byCluster_integrated.pdf"))
  plotlist[[m]]<-p
}

p<- ggarrange(plotlist = plotlist, labels = names(genelist), common.legend = F)
p
ggsave(file=paste0(outputdir, "Top10Genes_Heatmaps_byCluster_integrated.pdf"), p)
```

```{r, fig.width=14, fig.height=7}

FeaturePlot_scCustom(samples.integrated.filt, features = c("Lipe1", "Rfx7", "Olfr344", "S100a8"), 
                     layer="data", colors_use = viridis_dark_high, , reduction = "umap")

```

```{r, fig.width=7, fig.height=7}
DimPlot(samples.integrated.filt, reduction = "umap")

```

## Comparison of DEGs in DAMs between treatments/genotypes

### define DAM

DAM profiles based on Cell. 2017 Jun 15;169(7):1276-1290.e17. doi: 10.1016/j.cell.2017.05.018)

##### DAM1

```{r, fig.width=12, fig.height=4}
#load("./data/DAM_genelists.RData")
DefaultAssay(samples.integrated.filt) = "RNA"

samples.integrated.filt<- AddModuleScore(object = samples.integrated.filt, 
                                         assay="RNA", slot="data",
                                         features = list(DAM1_up), name = "DAM_score")

#FeaturePlot_scCustom(samples.integrated.filt, features = "DAM_score1")+scale_color_viridis_c()

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          assay="RNA", slot="data",
                                          features = list(DAM_up), name = "DAM_score")

p4 <- FeaturePlot_scCustom(seurat_object = samples.integrated.filt, reduction = "umap",
                           features = "DAM_score1", colors_use = viridis_dark_high)
p4a <- DimPlot(object = samples.integrated.filt, reduction = "umap")

DAMcutoff = 1.5
phist = ggplot(data = samples.integrated.filt@meta.data, aes(x=DAM_score1))+
  geom_density()+geom_vline(xintercept = DAMcutoff)+
  theme_classic()

samples.integrated.filt$DAM_binary = as.factor(samples.integrated.filt$DAM_score1>=DAMcutoff)
pbin <- DimPlot(object = samples.integrated.filt, group.by =  "DAM_binary", reduction = "umap")

comb = phist | p4 | pbin| p4a
comb

prop.table(table(DAMs=samples.integrated.filt$DAM_binary, Clusters=samples.integrated.filt$labels), margin = 2)

dam_GT_p <-DimPlot(object = samples.integrated.filt, group.by =  "DAM_binary", split.by = "Genotype_Treatment")


ggsave(filename = paste0(outputdir, "Microglia_DAM1_callingintegrated.pdf"), comb)
ggsave(filename = paste0(outputdir, "Microglia_DAM1_bygenotype_integrated.pdf"), dam_GT_p)
```


```{r}
phist
p4
pbin
ggsave(filename = paste0(outputdir, "Microglia_DAM1_thresholding_integrated.pdf"), phist)
ggsave(filename = paste0(outputdir, "Microglia_DAM1_scoring_integrated.pdf"), p4)
ggsave(filename = paste0(outputdir, "Microglia_DAM1_binary_integrated.pdf"), pbin)


```

##### DAM2

```{r, fig.width=12, fig.height=4}
samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          features = list(DAM2_up), name = "DAM2_score")
p4 <- FeaturePlot_scCustom(seurat_object = samples.integrated.filt, reduction = "umap",
                           features = "DAM2_score1", colors_use = viridis_dark_high)
p4a <- DimPlot(object = samples.integrated.filt, reduction = "umap")

DAM2cutoff = 1.0
phist = ggplot(data = samples.integrated.filt@meta.data, aes(x=DAM2_score1))+
  geom_density()+geom_vline(xintercept = DAM2cutoff)+
  theme_classic()

samples.integrated.filt$DAM2_binary = as.factor(samples.integrated.filt$DAM2_score1>=DAM2cutoff)
pbin <- DimPlot(object = samples.integrated.filt, reduction = "umap",
                group.by =  "DAM2_binary")
comb = phist | p4 | pbin| p4a


dam_GT_p <-DimPlot(object = samples.integrated.filt,reduction = "umap", group.by =  "DAM2_binary", split.by = "Genotype_Treatment")

comb
dam_GT_p

ggsave(filename = paste0(outputdir, "Microglia_DAM2_calling_integrated.pdf"), comb)
ggsave(filename = paste0(outputdir, "Microglia_DAM2_bygenotype_integrated.pdf"), dam_GT_p)


prop.table(table(DAM2=samples.integrated.filt$DAM2_binary, Clusters=samples.integrated.filt$labels), margin = 2)
prop.table(table(DAM2=samples.integrated.filt$DAM2_binary, DAMs=samples.integrated.filt$DAM_binary), margin = 2)
```

```{r, fig.width=12}
a <- ggplot(samples.integrated.filt@meta.data, aes(x=DAM2_score1, y=DAM_score1))+
  geom_density_2d_filled()+geom_point(col="#FFFFFF55")+theme_classic()+ggtitle("all cells")

b<- ggplot(samples.integrated.filt@meta.data[samples.integrated.filt$DAM_binary==TRUE,], aes(x=DAM2_score1, y=DAM_score1))+geom_density_2d_filled()+geom_point(col="#FFFFFF55")+theme_classic()+ggtitle("DAMs only")

combo <- a|b
combo
ggsave(filename = paste0(outputdir, "Microglia_DAMvsDAM2_contourplot_integrated.pdf"), combo)

```


```{r}
metadata=samples.integrated.filt@meta.data
restab = table(GT_Treat = metadata$Genotype_Treatment, DAMs=metadata$DAM_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot <- ggplot(as.data.frame(restab), 
                  aes(fill=DAMs, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()+ggtitle("DAM proportions")

proplot

ggsave(filename = paste0(outputdir, "Microglia_DAMproportions_per_condition_all_integrated.pdf"), proplot)

library(chisq.posthoc.test)
chisq.posthoc.test(restab)

#To Do Pairwise 

```

#### DAMs in APPPS1 vs. DAMs in APPPS1 + Stroke 

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "APPPS1", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), DAMs=metadata$DAM_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=DAMs, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

proplot

ggsave(filename = paste0(outputdir, "Microglia_DAMproportions_per_condition_APPPS1_integrated.pdf"), proplot)

```

#### DAMs in WT vs. DAMs in WT + Stroke

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "WT", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), DAMs=metadata$DAM_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=DAMs, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

proplot
ggsave(filename = paste0(outputdir, "Microglia_DAMproportions_per_condition_WT_integrated.pdf"), proplot)


```

#### DAM1 vs DAM2

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$DAM_binary==TRUE,]
restab = table(GT_Treat = metadata$Genotype_Treatment, DAM2=metadata$DAM2_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=DAM2, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

chisq.posthoc.test(restab)

proplot
ggsave(filename = paste0(outputdir, "Microglia_DAM2proportionsofDAM1_per_condition_integrated.pdf"), proplot)
```

```{r}

metadata=samples.integrated.filt@meta.data[(samples.integrated.filt$DAM_binary==TRUE) & (samples.integrated.filt$Genotype=="APPPS1"),]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), DAM2=metadata$DAM2_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<- ggplot(as.data.frame(restab), 
                 aes(fill=DAM2, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()


proplot
ggsave(filename = paste0(outputdir, "Microglia_DAM2proportionsofDAM1_per_condition_APPPS1_integrated.pdf"), proplot)

```

```{r}

metadata=samples.integrated.filt@meta.data[(samples.integrated.filt$DAM_binary==TRUE) & (samples.integrated.filt$Genotype=="WT"),]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), DAM2=metadata$DAM2_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=DAM2, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()


proplot
ggsave(filename = paste0(outputdir, "Microglia_DAM2proportionsofDAM1_per_condition_WT_integrated.pdf"), proplot)

```

#### Homeostatic Cells

```{r fig.width=9, fig.height=15}
# list from Jan Hofmann 
a <- RidgePlot(samples.integrated.filt, features = homeostat.genes, ncol = 2)
a

b<- FeaturePlot_scCustom(samples.integrated.filt, reduction="umap",
                         features = homeostat.genes, colors_use = viridis_dark_high)

b

ggsave(filename = paste0(outputdir, "Microglia_HomeostatGenes_Cluster_Ridgeplot_integrated.pdf"), a)
ggsave(filename = paste0(outputdir, "Microglia_HomeostatGenes_Featureplot_integrated.pdf"), a)
```

```{r, fig.width=14, fig.height=4}

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, layer ="scaled.data",
                                          features = list(homeostat.genes), name = "Homeo_score")

p4 <- FeaturePlot_scCustom(seurat_object = samples.integrated.filt, 
                           reduction = "umap", 
                           features = "Homeo_score1", 
                           colors_use = viridis_dark_high)

p4a <- DimPlot(object = samples.integrated.filt, reduction = "umap")

Homeocutff = 1.5
phist = ggplot(data = samples.integrated.filt@meta.data, aes(x=Homeo_score1))+
  geom_density()+geom_vline(xintercept = Homeocutff)+theme_classic()

samples.integrated.filt$Homeo_binary = as.factor(samples.integrated.filt$Homeo_score1>=Homeocutff)
pbin <- DimPlot(object = samples.integrated.filt,  reduction = "umap", group.by =  "Homeo_binary")
comb = phist | p4 | pbin| p4a

Hom_GT_p <-DimPlot(object = samples.integrated.filt, reduction = "umap",
                   group.by =  "Homeo_binary", split.by = "Genotype_Treatment")

comb
Hom_GT_p

prop.table(table(Homeo=samples.integrated.filt$Homeo_binary, 
                 Clusters=samples.integrated.filt$labels), margin = 2)


ggsave(filename = paste0(outputdir, "Microglia_Homeo_calling_integrated.pdf"), comb)
ggsave(filename = paste0(outputdir, "Microglia_Homeo_bygenotype_integrated.pdf"), Hom_GT_p)

```

```{r}
phist
p4
pbin
ggsave(filename = paste0(outputdir, "Microglia_Homeo_thresholding_integrated.pdf"), phist)
ggsave(filename = paste0(outputdir, "Microglia_Homeo_scoring_integrated.pdf"), p4)
ggsave(filename = paste0(outputdir, "Microglia_Homeo_binary_integrated.pdf"), pbin)

```

#### Homeo vs DAM


```{r}

metadata = samples.integrated.filt@meta.data

metadata$Homeo_DAM <- factor(1*as.logical(metadata$Homeo_binary) + 2*as.logical(metadata$DAM_binary), levels = c(0,1,2,3), labels = c("Neither", "Homeo", "DAM", "Unknown"))

samples.integrated.filt@meta.data <- metadata


a = DimPlot(samples.integrated.filt, reduction="umap", group.by = "Homeo_DAM")
a
ggsave(filename = paste0(outputdir, "Microglia_Homeo_DAM_integrated.pdf"), a)
```


```{r}
restab = table(GT_Treat = metadata$Genotype_Treatment, Homeo=metadata$Homeo_DAM)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=Homeo, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

chisq.posthoc.test(restab)

proplot
ggsave(filename = paste0(outputdir, "Microglia_HomeoDAMproportions_per_condition_integrated.pdf"), proplot)
```


```{r}
restab = table(Clust = metadata$labels, Homeo=metadata$Homeo_DAM)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=Homeo, x=Clust, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

chisq.posthoc.test(restab)

proplot
ggsave(filename = paste0(outputdir, "Microglia_HomeoDAMproportions_per_ckuster_integrated.pdf"), proplot)
```


```{r}
restab = table(Clust = metadata$labels, Celltype=metadata$Mouse_ID)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=Clust, x=Celltype, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

chisq.posthoc.test(restab)

proplot <- proplot + theme(axis.text.x = element_text(angle=90))
proplot
ggsave(filename = paste0(outputdir, "Microglia_celltype_labels_integrated.pdf"), proplot)
```


```{r}
restab = table(gender=metadata$Gender, Clust = metadata$labels)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=gender, x=Clust, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

chisq.posthoc.test(restab)

proplot
ggsave(filename = paste0(outputdir, "Microglia_HomeoDAMproportions_per_gender_integrated.pdf"), proplot)
```

```{r, fig.width=12}
a <- ggplot(samples.integrated.filt@meta.data, aes(x=Homeo_score1, y=DAM_score1))+
  geom_density_2d_filled()+geom_point(col="#FFFFFF55")+theme_classic()+ggtitle("all cells")

b<- ggplot(samples.integrated.filt@meta.data[samples.integrated.filt$DAM_binary==TRUE,], aes(x=Homeo_score1, y=DAM_score1))+geom_density_2d_filled()+geom_point(col="#FFFFFF55")+theme_classic()+ggtitle("DAMs only")

a|b

ggsave(filename = paste0(outputdir, "Microglia_DAMvsHomeo_integrated.pdf"), a|b)

```

#### Homeo in Genotype Treatment

proportion tables show percentages of Homeostatic cells per condition

```{r}
metadata=samples.integrated.filt@meta.data
restab = table(GT_Treat = metadata$Genotype_Treatment, Homeo=metadata$Homeo_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=Homeo, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()


chisq.posthoc.test(restab)


proplot
ggsave(filename = paste0(outputdir, "Microglia_Homeoproportions_per_condition_integrated.pdf"), proplot)

```


```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype=="APPPS1",]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Homeo=metadata$Homeo_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<- ggplot(as.data.frame(restab), 
                 aes(fill=Homeo, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

proplot
ggsave(filename = paste0(outputdir, "Microglia_Homeoproportions_per_condition_APPPS1_integrated.pdf"), proplot)

```

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype=="WT",]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Homeo=metadata$Homeo_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

proplot<-ggplot(as.data.frame(restab), 
                aes(fill=Homeo, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

proplot
ggsave(filename = paste0(outputdir, "Microglia_Homeoproportions_per_condition_WT_integrated.pdf"), proplot)

```

### Spatial cell types

```{r, fig.height=12,fig.width=5}


Idents(samples.integrated.filt) = "Genotype_Treatment"
DoHeatmap(samples.integrated.filt, slot = "data", features = c(distal_genes, control_genes, Stroke_genes))+ggtitle("spatial genes all")
DoHeatmap(samples.integrated.filt, slot = "data",features = c(distal_genes))+ggtitle("spatial genes distal")
DoHeatmap(samples.integrated.filt, slot = "data",features = c(control_genes))+ggtitle("spatial genes control")
DoHeatmap(samples.integrated.filt, slot = "data",features = c(Stroke_genes))+ggtitle("spatial genes peri-ictus")
```

#### Distal Cell types 

```{r, fig.width=12, fig.height=4}
#load("./data/DAM_genelists.RData")
DefaultAssay(samples.integrated.filt) = "RNA"

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          features = list(distal_genes), name = "Distal")


p4 <- FeaturePlot_scCustom(seurat_object = samples.integrated.filt, reduction = "umap",
                           features = "Distal1", colors_use = viridis_dark_high)
p4a <- DimPlot(object = samples.integrated.filt, reduction = "umap")

Distalcutoff = 0.125
phist = ggplot(data = samples.integrated.filt@meta.data, aes(x=Distal1))+
  geom_density()+geom_vline(xintercept = Distalcutoff)+
  theme_classic()

samples.integrated.filt$Distal_binary = as.factor(samples.integrated.filt$Distal1>=Distalcutoff)
pbin <- DimPlot(object = samples.integrated.filt,  reduction = "umap",group.by =  "Distal_binary")
phist | p4 | pbin| p4a

prop.table(table(Distals=samples.integrated.filt$Distal_binary, Clusters=samples.integrated.filt$labels), margin = 2)

DimPlot(object = samples.integrated.filt, group.by =  "Distal_binary", reduction = "umap", split.by = "Genotype_Treatment")

```

#### Control Cell types

```{r, fig.width=12, fig.height=4}
#load("./data/DAM_genelists.RData")
DefaultAssay(samples.integrated.filt) = "RNA"

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          features = list(control_genes), name = "Control")


p4 <- FeaturePlot_scCustom(seurat_object = samples.integrated.filt,  reduction = "umap",
                           features = "Control1", colors_use = viridis_dark_high)
p4a <- DimPlot(object = samples.integrated.filt, reduction = "umap")

Controlcutoff = 0.1
phist = ggplot(data = samples.integrated.filt@meta.data, aes(x=Control1))+
  geom_density()+geom_vline(xintercept = Controlcutoff)+
  theme_classic()

samples.integrated.filt$Control_binary = as.factor(samples.integrated.filt$Control1>=Controlcutoff)
pbin <- DimPlot(object = samples.integrated.filt,  reduction = "umap",
                group.by =  "Control_binary")
phist | p4 | pbin| p4a

prop.table(table(Controls=samples.integrated.filt$Control_binary, Clusters=samples.integrated.filt$labels), margin = 2)

DimPlot(object = samples.integrated.filt, group.by =  "Control_binary", reduction = "umap",
        split.by = "Genotype_Treatment")

```
#### Peri_ictus Cell types

```{r, fig.width=12, fig.height=4}
#load("./data/DAM_genelists.RData")
DefaultAssay(samples.integrated.filt) = "RNA"

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          features = list(Stroke_genes), name = "Peri_ictus")


p4 <- FeaturePlot_scCustom(seurat_object = samples.integrated.filt,  reduction = "umap",features = "Peri_ictus1", colors_use = viridis_dark_high)
p4a <- DimPlot(object = samples.integrated.filt, reduction = "umap")

Strokecutoff = 1.75
phist = ggplot(data = samples.integrated.filt@meta.data, aes(x=Peri_ictus1))+
  geom_density()+geom_vline(xintercept = Strokecutoff)+
  theme_classic()+ggtitle("Peri_ictus")

samples.integrated.filt$Peri_ictus_binary = as.factor(samples.integrated.filt$Peri_ictus1>=Strokecutoff)
pbin <- DimPlot(object = samples.integrated.filt,  reduction = "umap",group.by =  "Peri_ictus_binary")
phist | p4 | pbin| p4a

prop.table(table(Strokes=samples.integrated.filt$Peri_ictus_binary, Clusters=samples.integrated.filt$labels), margin = 2)

DimPlot(object = samples.integrated.filt,  reduction = "umap",group.by =  "Peri_ictus_binary", split.by = "Genotype_Treatment")

```

```{r}
ggplot(samples.integrated.filt@meta.data, aes(x=Peri_ictus1, y=Control1))+
  geom_density_2d_filled()+geom_point(col="#FFFFFF55")+theme_classic()+ggtitle("all cells")

ggplot(samples.integrated.filt@meta.data[samples.integrated.filt$Peri_ictus_binary==TRUE,], aes(x=Peri_ictus1, y=Control1))+geom_density_2d_filled()+geom_point(col="#FFFFFF55")+theme_classic()+ggtitle("Peri_ictus only")
```

proportion tables show percebtages of DAM cells per condition

#### Distals binary proportions

```{r}
metadata=samples.integrated.filt@meta.data
restab = table(GT_Treat = metadata$Genotype_Treatment, Distals=metadata$Distal_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Distals, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

library(chisq.posthoc.test)
chisq.posthoc.test(restab)

#To Do Pairwise 


```

#### Distals in APPPS1 

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "APPPS1", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Distals=metadata$Distal_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Distals, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()



```

#### Distals in WT

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "WT", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Distals=metadata$Distal_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Distals, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

```

#### Controls binary proportions

```{r}
metadata=samples.integrated.filt@meta.data
restab = table(GT_Treat = metadata$Genotype_Treatment, Controls=metadata$Control_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Controls, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

library(chisq.posthoc.test)
chisq.posthoc.test(restab)

#To Do Pairwise 


```

#### Controls in APPPS1

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "APPPS1", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Controls=metadata$Control_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Controls, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()



```

#### Controls in WT

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "WT", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Controls=metadata$Control_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Controls, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()
```

#### Peri_ictus binary proportions

```{r}
metadata=samples.integrated.filt@meta.data
restab = table(GT_Treat = metadata$Genotype_Treatment, Peri_Ictal=metadata$Peri_ictus_binary)
restab
prop.table(restab, margin = 1)
chisq.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Peri_Ictal, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

library(chisq.posthoc.test)
chisq.posthoc.test(restab)

#To Do Pairwise 


```

#### Peri_ictus in APPPS1 

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "APPPS1", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Peri_Ictal=metadata$Peri_ictus_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Peri_Ictal, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

```

#### Peri_ictus WT

```{r}
metadata=samples.integrated.filt@meta.data[samples.integrated.filt$Genotype == "WT", ]
restab = table(GT_Treat = as.character(metadata$Genotype_Treatment), Peri_Ictal=metadata$Peri_ictus_binary)
restab
prop.table(restab, margin = 1)
fisher.test(restab)

ggplot(as.data.frame(restab), 
       aes(fill=Peri_Ictal, x=GT_Treat, y=Freq))+
  geom_bar(position="fill",  color="black",stat="identity")+theme_classic()

```

# DEG comparing conditions overall

## WT Stroke vs no Stroke

avg_logFC: log fold-chage of the average expression between the two groups. Positive values indicate that the gene is more highly expressed in the first group

```{r, fig.width=6, fig.height=6}
genes_to_label <- openxlsx::read.xlsx("/files/scSeq_Hefendehl/data/Volcano pot labels on graph.xlsx", 
                                      sheet = "Wildtype control vs Wildtype st")
genes_to_label <- c(genes_to_label$Up, genes_to_label$Down)

Idents(samples.integrated.filt) <- "Genotype_Treatment"

DimPlot(samples.integrated.filt, reduction="umap")
DEG_Wt_CtrlvsStroke <-FindMarkers(samples.integrated.filt, 
                                  ident.1 = "WT_Stroke", ident.2 = "WT_Ctrl", 
                                  assay = "RNA", slot = "data",
                                  test.use = "wilcox", logfc.threshold = 0)

genes_to_label[!genes_to_label %in% rownames(DEG_Wt_CtrlvsStroke)]
```


```{r, fig.width=6, fig.height=6}
p = EnhancedVolcano::EnhancedVolcano(DEG_Wt_CtrlvsStroke, 
                                     x = "avg_log2FC", 
                                     y = "p_val_adj", 
                                     pCutoff=0.05,
                                     lab=rownames(DEG_Wt_CtrlvsStroke),
                                     selectLab = genes_to_label,
                                     title = "WT_Ctrl vs WT_Stroke")



display_tab(DEG_Wt_CtrlvsStroke)
p
write.xlsx(DEG_Wt_CtrlvsStroke, file = paste0(outputdir, "Microglia_DEX_WTCtrlvsWTStroke_integrated.xlsx"),rowNames = TRUE)
ggsave(filename = paste0(outputdir, "Microglia_Volcano_WTCtrlvsWTStroke_integrated.pdf"), p)
```

## APPPS1 Stroke vs APPPS1 no Stroke

avg_logFC: log fold-chage of the average expression between the two groups. Positive values indicate that the gene is more highly expressed in the first group

```{r fig.width=6, fig.height=6}
genes_to_label <- openxlsx::read.xlsx("/files/scSeq_Hefendehl/data/Volcano pot labels on graph.xlsx", 
                                      sheet = "APPPS1 control Vs APPPS1 stroke")
genes_to_label <- c(genes_to_label$Up, genes_to_label$Down)



DEG_APPPS1_CtrlvsStroke <-FindMarkers(samples.integrated.filt, 
                                      ident.1 = "APPPS1_Stroke", ident.2 = "APPPS1_Ctrl", 
                                      assay = "RNA", slot = "data",
                                      test.use = "wilcox", logfc.threshold = 0)
display_tab(DEG_APPPS1_CtrlvsStroke)

genes_to_label[!genes_to_label %in% rownames(DEG_APPPS1_CtrlvsStroke)]

p<- EnhancedVolcano::EnhancedVolcano(DEG_APPPS1_CtrlvsStroke, 
                                     x = "avg_log2FC", 
                                     y = "p_val_adj",
                                     lab=rownames(DEG_APPPS1_CtrlvsStroke), 
                                     selectLab = genes_to_label,
                                     pCutoff = 0.05,
                                     title = "APPPS1_Ctrl vs APPPS1_Stroke")


p
write.xlsx(DEG_APPPS1_CtrlvsStroke, file = paste0(outputdir, "Microglia_DEX_APPPS1CtrlvsAPPPS1Stroke_integrated.xlsx"),rowNames = TRUE)
ggsave(filename = paste0(outputdir, "Microglia_Volcano_APPPS1CtrlvsAPPPS1Stroke_integrated.pdf"), p)

```


## WT Stroke vs APPPS1Stroke

```{r fig.width=6, fig.height=6}


# plot here only DAM markers 

genes_to_label <- unique(c(DAM_up, DAM1_down, DAM1_up, DAM1vsDAM2, DAM2_up))

DEG_Stroke_APPPS1vsWT <-FindMarkers(samples.integrated.filt, 
                                    ident.1 = "APPPS1_Stroke", ident.2 = "WT_Stroke", 
                                    assay = "RNA", slot = "data",
                                    test.use = "wilcox", logfc.threshold = 0)
display_tab(DEG_Stroke_APPPS1vsWT)

genes_to_label[!genes_to_label %in% rownames(DEG_Stroke_APPPS1vsWT)]

p<- EnhancedVolcano::EnhancedVolcano(DEG_Stroke_APPPS1vsWT, 
                                     x = "avg_log2FC", 
                                     y = "p_val_adj",
                                     lab=rownames(DEG_Stroke_APPPS1vsWT), 
                                     selectLab = genes_to_label,
                                     pCutoff = 0.05,
                                     title = "WT_stroke vs APPPS1_Stroke")

p
write.xlsx(DEG_Stroke_APPPS1vsWT, file = paste0(outputdir, "Microglia_DEX_Stroke_APPPS1vsWT_integrated.xlsx"),rowNames = TRUE)
ggsave(filename = paste0(outputdir, "Microglia_Volcano_WTstrokevsAPPPS1StrokeDAMgenes_integrated.pdf"), p)

```

## WT ctr; vs APPPS1 ctrl

avg_logFC: log fold-chage of the average expression between the two groups. Positive values indicate that the gene is more highly expressed in the first group

```{r fig.width=6, fig.height=6}

genes_to_label <- openxlsx::read.xlsx("/files/scSeq_Hefendehl/data/Volcano pot labels on graph.xlsx", 
                                      sheet = "Wild type control Vs APPPS1 con")
genes_to_label <- c(genes_to_label$Up, genes_to_label$Down)

DEG_APPPS1_CtrlvsWT_Ctrl <-FindMarkers(samples.integrated.filt, 
                                       ident.1 = "APPPS1_Ctrl", ident.2 = "WT_Ctrl", 
                                       assay = "RNA", slot = "data",
                                       test.use = "wilcox", logfc.threshold = 0)
display_tab(DEG_APPPS1_CtrlvsWT_Ctrl)
genes_to_label[!genes_to_label %in% rownames(DEG_APPPS1_CtrlvsWT_Ctrl)]

p<- EnhancedVolcano::EnhancedVolcano(DEG_APPPS1_CtrlvsWT_Ctrl, 
                                     x = "avg_log2FC", 
                                     y = "p_val_adj", 
                                     lab=rownames(DEG_APPPS1_CtrlvsWT_Ctrl), 
                                     selectLab = genes_to_label,
                                     pCutoff = 0.05,
                                     title = "WT_Ctrl vs APPPS1_Ctrl")


p
write.xlsx(DEG_APPPS1_CtrlvsWT_Ctrl, file = paste0(outputdir, "Microglia_DEX_APPPS1_CtrlvsWT_Ctrl_integrated.xlsx"),rowNames = TRUE)
ggsave(filename = paste0(outputdir, "Microglia_Volcano_WTCtrlvsAPPPS1ctrl_integrated.pdf"), p)

```



## Scatterplot of DEGs of all four conditions


```{r}
genelist_both= intersect(rownames(DEG_Wt_CtrlvsStroke), rownames(DEG_APPPS1_CtrlvsStroke))

plotdata=data.frame(log2FC_WTCtrl_vs_WTStroke=DEG_Wt_CtrlvsStroke[genelist_both,]$avg_log2FC, 
                    padj_WTCtrl_vs_WTStroke=DEG_Wt_CtrlvsStroke[genelist_both,]$p_val, 
                    log2FC_APPPS1Ctrl_vs_APPPS1Stroke = DEG_APPPS1_CtrlvsStroke[genelist_both,]$avg_log2FC, 
                    padj_APPPS1Ctrl_vs_APPPS1Stroke = DEG_APPPS1_CtrlvsStroke[genelist_both,]$p_val) %>% 
  mutate(
    collab=ifelse((padj_WTCtrl_vs_WTStroke < 0.05) & (padj_APPPS1Ctrl_vs_APPPS1Stroke < 0.05), "Both", 
                  ifelse((padj_WTCtrl_vs_WTStroke > 0.05) & (padj_APPPS1Ctrl_vs_APPPS1Stroke < 0.05), "APPPS1Stroke",
                         ifelse((padj_WTCtrl_vs_WTStroke < 0.05) & (padj_APPPS1Ctrl_vs_APPPS1Stroke > 0.05), "WTStroke", NA)))
  )

rownames(plotdata) = genelist_both

plotdata$label = rownames(plotdata)

# the next lines plot the top 20 genes of either analyses 
n=20
plotdata$label[! c(1:nrow(plotdata)) %in% c(
  order(abs(plotdata$log2FC_WTCtrl_vs_WTStroke), decreasing = T)[1:n], 
  order(abs(plotdata$log2FC_APPPS1Ctrl_vs_APPPS1Stroke), decreasing = T)[1:n])]=""


#if you execute the next two lines the labels you choose will be plotted
# to execute set FALSE to TRUE and change the list of gene names 
if(TRUE){
  genestoplot = c("S100a6", "Apoc4", "C4b", "Atp1a2", "Apoc1", "Fn1","Ntrk2")
  print(paste0(genestoplot[! genestoplot %in% rownames(plotdata)], " cannot be found"))
  plotdata$label[! rownames(plotdata) %in% genestoplot] = ""
}
```


```{r}
plotdata <- plotdata %>% 
  arrange(desc(is.na(collab)))

p <- ggplot(plotdata, aes(x=log2FC_WTCtrl_vs_WTStroke, y=log2FC_APPPS1Ctrl_vs_APPPS1Stroke, label=label))+
  geom_point(aes(colour = collab), alpha=0.7)+ 
  geom_text_repel(box.padding = 0.5, max.overlaps = Inf, color="black")+ 
  scale_color_discrete(palette=c( "#56B4E9", "#009E73","#F0E442"))+
  #geom_abline(slope=-1, intercept=0, lty=2)+geom_abline(slope=1, intercept=0, lty=2)+
  theme_classic()+ggtitle("log_FC")
p
ggsave(filename = paste0(outputdir, "Microglia_logFCCtrlvsStroke_APPPS1vsWT_all_integrated.pdf"), p)

```


```{r, fig.width=9, fig.height=7}
# sig only

genelist_both= 
  unique(c(rownames(DEG_Wt_CtrlvsStroke %>% dplyr::filter(p_val<0.05 & abs(avg_log2FC)>0.5)),
           rownames(DEG_APPPS1_CtrlvsStroke%>% dplyr::filter(p_val<0.05 & abs(avg_log2FC)>0.5))))

plotdata=data.frame(log2FC_WTCtrl_vs_WTStroke=DEG_Wt_CtrlvsStroke[genelist_both,]$avg_log2FC, 
                    padj_WTCtrl_vs_WTStroke=DEG_Wt_CtrlvsStroke[genelist_both,]$p_val, 
                    log2FC_APPPS1Ctrl_vs_APPPS1Stroke = DEG_APPPS1_CtrlvsStroke[genelist_both,]$avg_log2FC, 
                    padj_APPPS1Ctrl_vs_APPPS1Stroke = DEG_APPPS1_CtrlvsStroke[genelist_both,]$p_val) %>% 
  mutate(
    collab=ifelse((padj_WTCtrl_vs_WTStroke < 0.05) & (padj_APPPS1Ctrl_vs_APPPS1Stroke < 0.05), "Both", 
                  ifelse((padj_WTCtrl_vs_WTStroke > 0.05) & (padj_APPPS1Ctrl_vs_APPPS1Stroke < 0.05), "APPPS1Stroke",
                         ifelse((padj_WTCtrl_vs_WTStroke < 0.05) & (padj_APPPS1Ctrl_vs_APPPS1Stroke > 0.05), "WTStroke", NA)))
  )

rownames(plotdata) = genelist_both

plotdata$label = rownames(plotdata)




#if you execute the next lines the labels you choose will be plotted
# if set to TRUE, change the list of gene names and they will be plotted
# if set to FALSE the top n=15 of both will be plotted 

mode = "both" # alternative "selected" or "topN" or "both"
genestoplot = c("S100a6", "Apoc4", "C4b", "Atp1a2", "Apoc1", "Fn1","Ntrk2") # name here your gene of interest
topn = 15 # name the top number of genes to be plotted 


### don't change anything below the line here 

plotdata$label = rownames(plotdata)

if(mode=="selected"){
  
  print(paste0(genestoplot[! genestoplot %in% rownames(plotdata)], " cannot be found"))
  plotdata$label[which(! rownames(plotdata) %in% genestoplot)] = ""
} else if (mode=="topN"){
  n=topn
  genestoplotn15 <-c(1:nrow(plotdata)) %in% 
    c(
      order(abs(plotdata$log2FC_WTCtrl_vs_WTStroke), decreasing = T)[1:n], 
      order(abs(plotdata$log2FC_APPPS1Ctrl_vs_APPPS1Stroke), decreasing = T)[1:n])
  plotdata$label[!genestoplotn15]=""
} else if (mode =="both") {
  print(paste0(genestoplot[! genestoplot %in% rownames(plotdata)], " cannot be found"))
  genestoplotcmbn <-(c(1:nrow(plotdata)) %in% 
                       c(
                         order(abs(plotdata$log2FC_WTCtrl_vs_WTStroke), decreasing = T)[1:n], 
                         order(abs(plotdata$log2FC_APPPS1Ctrl_vs_APPPS1Stroke), decreasing = T)[1:n])) | (rownames(plotdata) %in% genestoplot)
  plotdata$label[which(! genestoplotcmbn)] = ""
}

```


```{r, fig.width=9, fig.height=7}

p <- ggplot(plotdata, aes(x=log2FC_WTCtrl_vs_WTStroke, y=log2FC_APPPS1Ctrl_vs_APPPS1Stroke, label=label))+
  geom_point(aes(colour = collab), alpha=0.7)+ 
  geom_text_repel(box.padding = 0.5, max.overlaps = Inf,
                  color="black")+ 
  scale_color_discrete(palette=c( "#56B4E9", "#009E73","#F0E442"))+
  #geom_abline(slope=-1, intercept=0, lty=2)+geom_abline(slope=1, intercept=0, lty=2)+
  theme_classic()+ggtitle("log_FC")

p
ggsave(filename = paste0(outputdir, "Microglia_logFCCtrlvsStroke_APPPS1vsWT_sig_only_integrated.pdf"), p)



```


```{r fig.width=8, fig.height=8}
reads_data = JoinLayers(samples.integrated.filt, assay = "RNA") %>%    
  GetAssayData(assay = "RNA", layer="data")  # loaded data as used above 
#reads_data = samples.integrated.filt@assays$RNA@counts # Raw counts/reads  
#reads_data = samples.integrated.filt@assays$RNA@scale.data  #scaled/reads

barplot_genes = function(gene, split.by =T){
  require(ggbeeswarm)
  
  plotdata = data.frame(log2Reads = log2(reads_data[gene,]+1), 
                        Genotype_Treatment = samples.integrated.filt@meta.data$Genotype_Treatment, 
                        Cluster = samples.integrated.filt@meta.data$labels)
  
  if(split.by=="none" | split.by=="cluster") {
    p=ggplot(plotdata, aes(x=Genotype_Treatment, y=log2Reads, fill=Genotype_Treatment))+
      geom_violin()+
      geom_beeswarm(alpha=0.4, method = "hex")+ggtitle(gene)+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
    if(split.by=="cluster"){p=p+facet_wrap(~Cluster)}
  } else if (split.by=="Microglia") {
    p=ggplot(plotdata, aes(x=Cluster, y=log2Reads, fill=Cluster))+
      geom_violin()+
      geom_beeswarm(alpha=0.4, method = "hex")+
      ggtitle(gene)+
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  }
  p
}

barplot_genes(gene="Ptprc", split.by ="cluster") # alternative split.by = "none" oder " Micorglia
barplot_genes(gene="Ptprc", split.by ="none") # just across the genotype_treatments 
barplot_genes(gene="Ptprc", split.by ="Microglia") # add +coord_cartesian(ylim=c(0, 7)) if you want to change the ylims
barplot_genes(gene="Ptprc", split.by ="Microglia")+coord_cartesian(ylim=c(0, 2))
barplot_genes(gene="Ptprc", split.by ="Microglia") + facet_wrap(~Genotype_Treatment)

FeaturePlot_scCustom(samples.integrated.filt, features = sort(genestoplot), reduction="umap", colors_use = viridis_dark_high)

```

## DEG Stroke Cells only 

```{r, eval=FALSE}

DEG_Strokebins_Wt_CtrlvsStroke <-FindMarkers(samples.integrated.filt, subset.ident = "Stroke_bin", subset="TRUE", 
                                             assay = "RNA", slot = "data",
                                             ident.1 = "WT_Stroke", ident.2 = "WT_Ctrl", 
                                             test.use = "wilcox", logfc.threshold = 0)

display_tab(DEG_Strokebins_Wt_CtrlvsStroke)

DEG_Strokebins_Appps1_CtrlvsStroke <-FindMarkers(samples.integrated.filt, subset.ident = "Stroke_bin", subset="TRUE", 
                                                 assay = "RNA", slot = "data",
                                                 ident.1 = "APPPS1_Stroke", ident.2 = "APPPS1_Ctrl", 
                                                 test.use = "wilcox", logfc.threshold = 0)

display_tab(DEG_Strokebins_Appps1_CtrlvsStroke)

genelist_both= intersect(rownames(DEG_Strokebins_Appps1_CtrlvsStroke), rownames(DEG_Strokebins_Wt_CtrlvsStroke))

plotdata=data.frame(WT=DEG_Strokebins_Wt_CtrlvsStroke[genelist_both,]$avg_log2FC, APPPS1 = DEG_Strokebins_Appps1_CtrlvsStroke[genelist_both,]$avg_log2FC)
rownames(plotdata) = genelist_both

plotdata$label = rownames(plotdata)
n=5
plotdata$label[! c(1:nrow(plotdata)) %in% 
                 c(
                   order(abs(plotdata$WT), decreasing = T)[1:n], 
                   order(abs(plotdata$APPPS1), decreasing = T)[1:n])]=""

plotdata$col="WT"
apps1=abs(plotdata$APPPS1)>abs(plotdata$WT)
plotdata$col[ apps1]="APPPS1"
plotdata$col[(abs(plotdata$APPPS1)<0.25) & (abs(plotdata$WT)<0.25)] <- NA
ggplot(plotdata, aes(x=WT, y=APPPS1, color=col, label=label))+geom_point(alpha=0.7)+
  geom_text(vjust=-0.7, color="black")+geom_abline(slope=-1, intercept=0, lty=2)+geom_abline(slope=1, intercept=0, lty=2)+theme_classic()+ggtitle("log_FC_Stroke_bins")

```

## DEG comparing Cluster 
All Clusters versus All Genes versus all Conditions 

### Define Main analysis pipeline 

```{r}

compareCluster <- function(
    ClusterA = "Microglia6",
    ClusterB = "Microglia0",
    seuratObject = samples.integrated.filt,
    output = "./",
    genelist = NULL,
    FCcutoff = 1,
    comorbid = FALSE,
    evgenecode=F
) {
  
  # ensure output exists
  if (!dir.exists(output)) dir.create(outputdir, recursive = TRUE)
  
  # name + logfile ALWAYS defined
  resname <- if (isTRUE(comorbid)) {
    paste0(ClusterA, "vs", ClusterB, "_comorbid")
  } else {
    paste0(ClusterA, "vs", ClusterB)
  }
  
  logfile <- file.path(output, paste0("log_", resname, ".txt"))
  if (file.exists(logfile)) file.remove(logfile)
  
  # helper to append to log
  log_line <- function(level, msg) {
    cat(
      paste0(
        format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
        " | ", level,
        " | ", ClusterA, " vs ", ClusterB,
        " | ", msg, "\n"
      ),
      file = logfile,
      append = TRUE
    )
  }
  
  # FindMarkers call (different args if comorbid)
  DEG_tmp <- tryCatch(
    withCallingHandlers(
      {
        if (isTRUE(comorbid)) {
          FindMarkers(
            seuratObject,
            assay = "RNA",
            subset.ident = "Genotype_Treatment",
            subset = "APPPS1_Stroke",
            layer = "data",
            ident.1 = ClusterA, ident.2 = ClusterB,
            test.use = "wilcox", logfc.threshold = 0
          )
        } else {
          FindMarkers(
            seuratObject,
            assay = "RNA",
            layer = "data",
            ident.1 = ClusterA, ident.2 = ClusterB,
            test.use = "wilcox", logfc.threshold = 0
          )
        }
      },
      warning = function(w) {
        log_line("WARNING", conditionMessage(w))
        invokeRestart("muffleWarning")
      }
    ),
    error = function(e) {
      log_line("ERROR", conditionMessage(e))
      return(NULL)
    }
  )
  
  # If FindMarkers failed, still return log + empty slots
  if (is.null(DEG_tmp)) {
    return(list(
      Markers = NULL,
      VulcanoPlot = NULL,
      qqplot = NULL,
      Goresults = NULL,
      GOplots = NULL,
      log = logfile
    ))
  }
  
  openxlsx::write.xlsx(
    DEG_tmp %>% dplyr::filter(p_val_adj<0.1),
    file = file.path(output, paste0("DEG_Gene_list_", resname, ".xlsx")),
    rowNames = TRUE
  )
  
  # label genes that actually exist in DEG table
  genes_to_label2 <- intersect(genelist, rownames(DEG_tmp))
  
  q <- gg_qqplot(DEG_tmp$p_val)
  
  p <- EnhancedVolcano::EnhancedVolcano(
    DEG_tmp,
    x = "avg_log2FC",
    y = "p_val_adj",
    lab = rownames(DEG_tmp),
    selectLab = genes_to_label2,
    pCutoff = 0.05,
    title = resname
  )
  
  ggsave(
    filename = file.path(output, paste0("DEG_Vulcano_", resname, "_integrated.pdf")),
    plot = (p | q),
    width = 10, height = 5
  )
  
  gene_univers <- rownames(seuratObject)
  
  gogenes <- list(
    genesigall  = rownames(DEG_tmp)[DEG_tmp$p_val_adj < 0.05 & abs(DEG_tmp$avg_log2FC) > FCcutoff],
    genesigup   = rownames(DEG_tmp)[DEG_tmp$p_val_adj < 0.05 & DEG_tmp$avg_log2FC > FCcutoff],
    genesigdown = rownames(DEG_tmp)[DEG_tmp$p_val_adj < 0.05 & DEG_tmp$avg_log2FC < -FCcutoff]
  )
  
  Gores <- getGOresults(
    geneset = gogenes,
    domain_scope = "known",
    genereference = gene_univers,
    return_genelist = evgenecode,
    organism = "mmusculus"
  )
  
  openxlsx::write.xlsx(
    Gores$result %>% dplyr::select(-parents),
    file = file.path(output, paste0("DEG_GOlist_", resname, ".xlsx")),
    rowNames = TRUE
  )
  
  plotlist <- list()
  for (m in unique(Gores$result$query)) {
    idx <- Gores$result$query == m & grepl("GO", Gores$result$source)
    
    if (sum(idx) >= 1) {
      plotlist[[m]] <- GOplot(Gores$result[idx, ], N = 10, Title = m)
    } else {
      plotlist[[m]] <- ggplot() +
        annotate(
          "text", x = 10, y = 10, size = 6,
          label = "no significant GO term\ncheck excel sheet for other enrichments"
        ) +
        theme_void() +
        ggtitle(m)
    }
  }
  
  ggsave(
    filename = file.path(output, paste0("DEG_GO_plots_", resname, ".pdf")),
    plot = ggarrange(plotlist = plotlist),
    width = 10, height = 5
  )
  
  return(list(
    Markers = DEG_tmp,
    VulcanoPlot = p,
    qqplot = q,
    Goresults = Gores$result,
    GOplots = plotlist,
    log = logfile
  ))
}

```

### DEG run for each combination 


```{r, warning=FALSE, message=FALSE}
if(T){
  Idents(samples.integrated.filt) <- "labels"
  
  labels_unique = sort(unique(samples.integrated.filt$labels))
  label_pairs <- combn(labels_unique, 2, simplify = FALSE)
  
  results_list <- list()
  
  for (pair in label_pairs) {
    print(paste(pair, collapse = "_vs_"))
    print("all")
    res <- compareCluster(
      ClusterA = pair[1],
      ClusterB = pair[2],
      FCcutoff = 0.5, 
      output = outputdir,
      genelist = genes_to_label
    )
    print("comorbid")
    res_comorbid <- compareCluster(
      ClusterA = pair[1],
      ClusterB = pair[2],
      FCcutoff = 0.5, 
      output = outputdir,
      genelist = genes_to_label,
      comorbid=T
    )
    results_list[[paste(pair, collapse = "_vs_")]] <- res
    results_list[[paste0(paste(pair, collapse = "_vs_"), 
                         "_comorbid")]] <- res_comorbid
  }
  
}
```


```{r, results='asis', message=FALSE, warning=FALSE, fig.width=10, fig.height=5}

# results_list: named list where each element is res from compareCluster()

for (nm in names(results_list)) {
  res <- results_list[[nm]]
  cat("\n\n## ", nm, "\n\n", sep = "")
  # --- 1) Volcano + QQ (patchwork-style)
  # If you're using patchwork, this should work:
  if (!is.null(res$VulcanoPlot) && !is.null(res$qqplot)) {
    p1 <- res$VulcanoPlot | res$qqplot
    print(p1)
  } else {
    cat("*No VulcanoPlot and/or qqplot found for this comparison.*\n\n")
  }
  
  # --- 2) GO plots (arranged)
  if (!is.null(res$GOplots) && length(res$GOplots) > 0) {
    p2 <- ggarrange(plotlist = res$GOplots, ncol = 2)
    print(p2)
  } else {
    cat("*No GOplots found for this comparison.*\n\n")
  }
  
  # --- 3) Markers table
  cat("\n\n### Markers\n\n")
  if (!is.null(res$Markers) && nrow(res$Markers) > 0) {
    print(kable(res$Markers %>% arrange(p_val) %>% slice_head(n = 100), format = "markdown"))
  } else {
    cat("*No markers found for this comparison.*\n\n")
  }
}
```

### within cluster Wt vs WT Stroke

```{r, results='asis', message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
Idents(samples.integrated.filt) <- "Genotype_Treatment"

for(clust in unique(samples.integrated.filt$labels)){
  print(clust)
  FCcutoff = 0.5
  resname=paste0(clust, "_WT_Ctrl_vs_WT_Stroke_")
  DEG_tmp<- FindMarkers(samples.integrated.filt[,samples.integrated.filt$labels==clust],
                        assay = "RNA",
                        layer = "data",
                        ident.1 = "WT_Ctrl", 
                        ident.2 = "WT_Stroke",
                        test.use = "wilcox", logfc.threshold = 0
  )
  
  res_sig <- DEG_tmp %>% dplyr::filter(p_val<0.05)
  openxlsx::write.xlsx(res_sig, 
                       file = file.path(outputdir, paste0("DEG_Gene_list_", resname, ".xlsx")),
                       rowNames = TRUE
  )
  q <- gg_qqplot(DEG_tmp$p_val)
  p <- EnhancedVolcano::EnhancedVolcano(
    DEG_tmp,
    x = "avg_log2FC",
    y = "p_val_adj",
    lab = rownames(DEG_tmp),
    selectLab = genes_to_label,
    pCutoff = 0.05,
    title = resname
  )
  ggsave(
    filename = file.path(outputdir, paste0("DEG_Vulcano_", resname, "integrated.pdf")),
    plot = (p | q),
    width = 10, height = 5
  )
  
  gene_univers <- rownames(samples.integrated.filt)
  
  gogenes <- list(
    genesigall  = rownames(DEG_tmp)[DEG_tmp$p_val_adj < 0.05 & abs(DEG_tmp$avg_log2FC) > FCcutoff],
    genesigup   = rownames(DEG_tmp)[DEG_tmp$p_val_adj < 0.05 & DEG_tmp$avg_log2FC > FCcutoff],
    genesigdown = rownames(DEG_tmp)[DEG_tmp$p_val_adj < 0.05 & DEG_tmp$avg_log2FC < -FCcutoff]
  )
  
  if(any(lapply(gogenes, length)>0)){
    Gores <- getGOresults(
      geneset = gogenes,
      domain_scope = "known",
      genereference = gene_univers,
      return_genelist = F,
      organism = "mmusculus"
    )
  } else {
    print("no significant genes identified")
    Gores <- NULL
  }
  
  
  if(length(Gores)!=0){
    openxlsx::write.xlsx(
      Gores$result %>% dplyr::select(-parents),
      file = file.path(outputdir, paste0("DEG_GOlist_", resname, ".xlsx")),
      rowNames = TRUE
    )
    plotlist <- list()
    for (m in unique(Gores$result$query)) {
      idx <- Gores$result$query == m & grepl("GO", Gores$result$source)
      
      if (sum(idx) >= 1) {
        plotlist[[m]] <- GOplot(Gores$result[idx, ], N = 10, Title = m)
      } else {
        plotlist[[m]] <- ggplot() +
          annotate(
            "text", x = 10, y = 10, size = 6,
            label = "no significant GO term\ncheck excel sheet for other enrichments"
          ) +
          theme_void() +
          ggtitle(m)
      }
    }
    
    ggsave(
      filename = file.path(outputdir, paste0("DEG_GO_plots_", resname, ".pdf")),
      plot = ggarrange(plotlist = plotlist),
      width = 10, height = 5
    ) } else {
      openxlsx::write.xlsx(data.frame(resname="No results identified"),
                           file = file.path(outputdir, paste0("DEG_GOlist_", resname, ".xlsx")))
      p <- ggplot() +
        annotate(
          "text", x = 10, y = 10, size = 6,
          label = "no significant enrichments"
        ) +
        theme_void()
      ggsave(
        filename = file.path(outputdir, paste0("DEG_GO_plots_", resname, ".pdf")),
        plot = p,
        width = 10, height = 5
      )
      
    }
  
  
}

Idents(samples.integrated.filt) <- "labels"
```


### within cluster APPPS1_Stroke vs WT Stroke

```{r, results='asis', message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
Idents(samples.integrated.filt) <- "Genotype_Treatment"

for(clust in unique(samples.integrated.filt$labels)){
  print(clust)
  FCcutoff = 0.5
  resname=paste0(clust, "_WT_stroke_vs_APPPS1_Stroke_")
  DEG_tmp<- FindMarkers(samples.integrated.filt[,samples.integrated.filt$labels==clust],
                        assay = "RNA",
                        layer = "data",
                        ident.1 = "APPPS1_Stroke", 
                        ident.2 = "WT_Stroke",
                        test.use = "wilcox", logfc.threshold = 0
  )
  
  res_sig <- DEG_tmp %>% dplyr::filter(p_val<0.05)
  openxlsx::write.xlsx(res_sig, 
                       file = file.path(outputdir, paste0("DEG_Gene_list_", resname, ".xlsx")),
                       rowNames = TRUE
  )
  q <- gg_qqplot(DEG_tmp$p_val)
  p <- EnhancedVolcano::EnhancedVolcano(
    DEG_tmp,
    x = "avg_log2FC",
    y = "p_val_adj",
    lab = rownames(DEG_tmp),
    selectLab = genes_to_label,
    pCutoff = 0.05,
    title = resname
  )
  ggsave(
    filename = file.path(outputdir, paste0("DEG_Vulcano_", resname, "integrated.pdf")),
    plot = (p | q),
    width = 10, height = 5
  )
  
  gene_univers <- rownames(samples.integrated.filt)
  
  gogenes <- list(
    genesigall  = rownames(DEG_tmp)[DEG_tmp$p_val_adj < 0.05 & abs(DEG_tmp$avg_log2FC) > FCcutoff],
    genesigup   = rownames(DEG_tmp)[DEG_tmp$p_val_adj < 0.05 & DEG_tmp$avg_log2FC > FCcutoff],
    genesigdown = rownames(DEG_tmp)[DEG_tmp$p_val_adj < 0.05 & DEG_tmp$avg_log2FC < -FCcutoff]
  )
  
  if(any(lapply(gogenes, length)>0)){
    Gores <- getGOresults(
      geneset = gogenes,
      domain_scope = "known",
      genereference = gene_univers,
      return_genelist = F,
      organism = "mmusculus"
    )
  } else {
    print("no significant genes identified")
    Gores <- NULL
  }
  
  
  if(length(Gores)!=0){
    openxlsx::write.xlsx(
      Gores$result %>% dplyr::select(-parents),
      file = file.path(outputdir, paste0("DEG_GOlist_", resname, ".xlsx")),
      rowNames = TRUE
    )
    plotlist <- list()
    for (m in unique(Gores$result$query)) {
      idx <- Gores$result$query == m & grepl("GO", Gores$result$source)
      
      if (sum(idx) >= 1) {
        plotlist[[m]] <- GOplot(Gores$result[idx, ], N = 10, Title = m)
      } else {
        plotlist[[m]] <- ggplot() +
          annotate(
            "text", x = 10, y = 10, size = 6,
            label = "no significant GO term\ncheck excel sheet for other enrichments"
          ) +
          theme_void() +
          ggtitle(m)
      }
    }
    
    ggsave(
      filename = file.path(outputdir, paste0("DEG_GO_plots_", resname, ".pdf")),
      plot = ggarrange(plotlist = plotlist),
      width = 10, height = 5
    ) } else {
      openxlsx::write.xlsx(data.frame(resname="No results identified"),
                           file = file.path(outputdir, paste0("DEG_GOlist_", resname, ".xlsx")))
      p <- ggplot() +
        annotate(
          "text", x = 10, y = 10, size = 6,
          label = "no significant enrichments"
        ) +
        theme_void()
      ggsave(
        filename = file.path(outputdir, paste0("DEG_GO_plots_", resname, ".pdf")),
        plot = p,
        width = 10, height = 5
      )
      
    }
  
  
}

Idents(samples.integrated.filt) <- "labels"
```

### within cluster APPPS1 vs APPPS1 Stroke


```{r, results='asis', message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
Idents(samples.integrated.filt) <- "Genotype_Treatment"

for(clust in unique(samples.integrated.filt$labels)){
  print(clust)
  FCcutoff = 0.5
  resname=paste0(clust, "_APPPS1_Stroke_vs_APPPS1_Ctrl_")
  DEG_tmp<- FindMarkers(samples.integrated.filt[,samples.integrated.filt$labels==clust],
                        assay = "RNA",
                        subset.ident = "labels",
                        subset = clust,
                        layer = "data",
                        ident.1 = "APPPS1_Stroke", ident.2 = "APPPS1_Ctrl",
                        test.use = "wilcox", logfc.threshold = 0
  )
  
  res_sig <- DEG_tmp %>% dplyr::filter(p_val<0.05)
  openxlsx::write.xlsx(res_sig, 
                       file = file.path(outputdir, paste0("DEG_Gene_list_", resname, ".xlsx")),
                       rowNames = TRUE
  )
  q <- gg_qqplot(DEG_tmp$p_val)
  p <- EnhancedVolcano::EnhancedVolcano(
    DEG_tmp,
    x = "avg_log2FC",
    y = "p_val_adj",
    lab = rownames(DEG_tmp),
    selectLab = genes_to_label,
    pCutoff = 0.05,
    title = resname
  )
  ggsave(
    filename = file.path(outputdir, paste0("DEG_Vulcano_", resname, "integrated.pdf")),
    plot = (p | q),
    width = 10, height = 5
  )
  
  gene_univers <- rownames(samples.integrated.filt)
  
  gogenes <- list(
    genesigall  = rownames(DEG_tmp)[DEG_tmp$p_val_adj < 0.05 & abs(DEG_tmp$avg_log2FC) > FCcutoff],
    genesigup   = rownames(DEG_tmp)[DEG_tmp$p_val_adj < 0.05 & DEG_tmp$avg_log2FC > FCcutoff],
    genesigdown = rownames(DEG_tmp)[DEG_tmp$p_val_adj < 0.05 & DEG_tmp$avg_log2FC < -FCcutoff]
  )
  
  if(any(lapply(gogenes, length)>0)){
    Gores <- getGOresults(
      geneset = gogenes,
      domain_scope = "known",
      genereference = gene_univers,
      return_genelist = F,
      organism = "mmusculus"
    )
  } else {
    print("no significant genes identified")
    Gores <- NULL
  }
  
  
  if(length(Gores)!=0){
    openxlsx::write.xlsx(
      Gores$result %>% dplyr::select(-parents),
      file = file.path(outputdir, paste0("DEG_GOlist_", resname, ".xlsx")),
      rowNames = TRUE
    )
    plotlist <- list()
    for (m in unique(Gores$result$query)) {
      idx <- Gores$result$query == m & grepl("GO", Gores$result$source)
      
      if (sum(idx) >= 1) {
        plotlist[[m]] <- GOplot(Gores$result[idx, ], N = 10, Title = m)
      } else {
        plotlist[[m]] <- ggplot() +
          annotate(
            "text", x = 10, y = 10, size = 6,
            label = "no significant GO term\ncheck excel sheet for other enrichments"
          ) +
          theme_void() +
          ggtitle(m)
      }
    }
    
    ggsave(
      filename = file.path(outputdir, paste0("DEG_GO_plots_", resname, ".pdf")),
      plot = ggarrange(plotlist = plotlist),
      width = 10, height = 5
    ) } else {
      openxlsx::write.xlsx(data.frame(resname="No results identified"),
                           file = file.path(outputdir, paste0("DEG_GOlist_", resname, ".xlsx")))
      p <- ggplot() +
        annotate(
          "text", x = 10, y = 10, size = 6,
          label = "no significant enrichments"
        ) +
        theme_void()
      ggsave(
        filename = file.path(outputdir, paste0("DEG_GO_plots_", resname, ".pdf")),
        plot = p,
        width = 10, height = 5
      )
      
    }
  
  
}

Idents(samples.integrated.filt) <- "labels"
```



```{r}
rm(list=c("seuratObject", "sce.filt", "reads", "reads_data", "comb", 
          "ref.sce.mm",
          "e2",
          "res",
          "res_comorbid", 
          "samples.integrated"))
gc()
```



# Pseudotime analysis

## slinghsot  starting Microglia 0

```{r, fig.width==7, fig.height=7}
samples.integrated.filt.meta = samples.integrated.filt@meta.data %>% dplyr::select(- contains("Pseudo"))
samples.integrated.filt@meta.data <- samples.integrated.filt.meta

sce.microglia <- as.SingleCellExperiment(samples.integrated.filt, assay = "RNA")
sce.microglia <- slingshot(sce.microglia, 
                           clusterLabels = 'labels', 
                           reducedDim = 'UMAP', 
                           start.clus="Microglia0", 
                           end.clus=c("Microglia3", "Microglia4", "Microglia5"),
                           omega=T, 
                           omega_scale=4, 
                           extend="n")

pt_col <- grep("slingPseudotime", names(colData(sce.microglia)), value = TRUE)

metasce <- data.frame(colData(sce.microglia)[,pt_col], 
                      row.names = rownames(colData(sce.microglia)))

colnames(metasce) <- pt_col

samples.integrated.filt.meta <- cbind(samples.integrated.filt.meta,metasce)

samples.integrated.filt.meta -> samples.integrated.filt@meta.data


dataset = data.frame(reducedDim(sce.microglia, "UMAP"), 
                     cluster = as.character(sce.microglia$labels), 
                     DAMs = as.character(sce.microglia$DAM_binary))

curves <- slingCurves(sce.microglia, as.df=T)

p = ggplot(dataset, aes(x=umap_1, y=umap_2))+
  geom_point(aes(col = cluster), size=2)+
  geom_point(data=subset(dataset, dataset$DAMs==T), aes(fill=DAMs), pch=21, size=0.7)+
  scale_fill_manual(values="black")
curves = curves %>% arrange(Lineage, Order)

p = p + geom_path(data = curves, aes(group = Lineage), lwd=1, lineend = "butt",)+theme_classic()


curves_arrows = curves %>% group_by(Lineage) %>%  mutate(xend = lead(umap_1), yend = lead(umap_2))
curves_arrows = curves_arrows[curves_arrows$Order %in% seq(10,150, by=floor((150-10)/10)),]

p = p + geom_segment(data = curves_arrows,
                     aes(x = umap_1, y = umap_2, xend = xend, yend = yend),
                     linewidth=0.5, 
                     arrow = arrow(length=unit(0.3, "cm"), type="closed"))
p

ggsave(filename = paste0(outputdir, "Pseudotime_linages_startMicroglia0_integrated.pdf"), p)

write_rds(p, paste0(outputdir, "Pseudotime_linages_startMicroglia0_integrated.rds"))
```

```{r}
library(TSCAN)

n.top.gene = 10

tscan_res <- list()
for (pt in pt_col){
  print(pt)
  tscan_res [[pt]]<- testPseudotime(sce.microglia, pseudotime = sce.microglia[[pt]])
}

```

Look for genes that are up/down-regulated along pseudotime, for the re-clustered cells (Same paper Fig 2G/H)

## Trajectory genes

By default, estimates of the spline coefficients are not returned as they are difficult to interpret. Rather, a log-fold change of expression along each path is estimated to provide some indication of the overall magnitude and direction of any change.

```{r}
reslist <- list()
reslist_feat <- list()

for (pt in names(tscan_res)){
  scePT <- tscan_res[[pt]]
  scePT<- scePT[order(scePT$FDR, decreasing = F),]
  display_tab(as.data.frame(scePT)[1:20,])
  reslist_feat[[pt]] <- FeaturePlot_scCustom(samples.integrated.filt, features = pt, reduction = "umap", colors_use = viridis_dark_high)
  
  write.xlsx(as.data.frame(scePT), rowNames = TRUE,
             file = paste0(outputdir, pt, "_genes.xslx"))
  PT1_up = scePT[scePT$logFC>0,]
  PT1_low = scePT[scePT$logFC<0,]
  p <- scater::plotExpression(sce.microglia, features=rownames(PT1_up)[1:n.top.gene], x=pt, colour_by="labels")+scale_color_hue() + ggtitle(paste(pt, "up reg"))
  p2 <- scater::plotExpression(sce.microglia, features=rownames(PT1_low)[1:n.top.gene], x=pt, colour_by="labels")+scale_color_hue()+ ggtitle(paste(pt, "down reg"))
  reslist[[pt]]=p|p2
}



```

```{r, fig.height=15, fig.width=36}

a <- ggarrange(plotlist = reslist)  
a

ggsave(filename = paste0(outputdir, "Pseudotime_Top10up_down_reg_Genes-correct_integrated.pdf"), a)

```



```{r, fig.height=8, fig.width=10}

a <- ggarrange(plotlist = reslist_feat)  
a

ggsave(filename = paste0(outputdir, "Pseudotime_Umap_Genes-correct_integrated.pdf"), a)

```
# WGCNA

In the paper <https://www.nature.com/articles/s41586-018-0023-4> this analysis was used to generate different modules of gene lists that can then be mapped onto different treatment/genotypes within one experiment (see Figure 4). This is an analysis that can be done in Seurat <https://smorabit.github.io/tutorials/9_scWGCNA_tutorial/> Which could be very cool for our data. In the paper they also have 3 different conditions in 2 genotypes.

avoid running this section as it takes very long

```{r}
nclusters <- length(unique(samples.integrated.filt$labels))
genes.use <- rownames(samples.integrated.filt)
targets <- samples.integrated.filt@meta.data
group <- as.factor(samples.integrated.filt$Genotype_Treatment)
```

```{r fig.width=8, fig.height=6}
force.recalc = T
powers = c(seq(1,10,by=1), seq(12,20, by=2));

# skips recalculation of net object as it takes long... 
reads <- JoinLayers(samples.integrated.filt, assay = "RNA") %>%    GetAssayData(assay = "RNA", layer="data") 
datExpr <- as.data.frame(reads[genes.use,])
rm("reads")
datExpr <- as.data.frame(t(datExpr))
datExpr <- datExpr[,goodGenes(datExpr)]

# expressed in more than 10% of cells 
exprgenes = colSums(datExpr>0)>nrow(datExpr)*0.1
datExpr <- datExpr[,exprgenes]

#annotated only 


mmacc=as.list(org.Mm.egALIAS2EG)
annotatedgenes = colnames(datExpr) %in% names(mmacc)
datExpr <- datExpr[,annotatedgenes]

powers = c(seq(1,10,by=1), seq(12,20, by=2));


enableWGCNAThreads(nThreads = 7)

if(force.recalc | (! file.exists(file=paste0(outputdir, "powertable.rds")))){  
  # Call the network topology analysis function for each set in turn
  powerTable = list(
    data = pickSoftThreshold(
      datExpr,
      powerVector=powers,
      verbose = 100,
      networkType="unsigned",
      corFnc="cor"
    )[[2]]
  );
  saveRDS(powerTable, file=paste0(outputdir, "powertable.rds"))
} else {
  powerTable = readRDS(file=paste0(outputdir, "powertable.rds"))
}

colors = c("blue", "red","black")
# Will plot these columns of the returned scale free analysis tables
plotCols = c(2,5,6,7)
colNames = c("Scale Free Topology Model Fit", "Mean connectivity", "mean connectivity",
             "Max connectivity");

# Get the minima and maxima of the plotted points
ylim = matrix(NA, nrow = 2, ncol = 4);
for (col in 1:length(plotCols)){
  ylim[1, col] = min(ylim[1, col], powerTable$data[, plotCols[col]], na.rm = TRUE);
  ylim[2, col] = max(ylim[2, col], powerTable$data[, plotCols[col]], na.rm = TRUE);
}

# Plot the quantities in the chosen columns vs. the soft thresholding power

par(mfcol = c(2,2));
par(mar = c(4.2, 4.2 , 2.2, 0.5))
cex1 = 0.7;

for (col in 1:length(plotCols)){
  plot(powerTable$data[,1], -sign(powerTable$data[,3])*powerTable$data[,2],
       xlab="Soft Threshold (power)",ylab=colNames[col],type="n", ylim = ylim[, col],
       main = colNames[col]);
  addGrid();
  
  if (col==1){
    text(powerTable$data[,1], -sign(powerTable$data[,3])*powerTable$data[,2],
         labels=powers,cex=cex1,col=colors[1]);
  } else
    text(powerTable$data[,1], powerTable$data[,plotCols[col]],
         labels=powers,cex=cex1,col=colors[1]);
  if (col==1){
    legend("bottomright", legend = 'Metacells', col = colors, pch = 20) ;
  } else
    legend("topright", legend = 'Metacells', col = colors, pch = 20) ;
}

softPower=3

nSets = 1
setLabels = 'OCS'
shortLabels = setLabels

multiExpr <- list()
multiExpr[['ODC']] <- list(data=datExpr)

checkSets(multiExpr) # check data size

# construct network
if(force.recalc | (! file.exists(file=paste0(outputdir, "net.rds")))){
  net=blockwiseConsensusModules(multiExpr, blocks = NULL,
                                maxBlockSize = 30000, ## This should be set to a smaller size if the user has limited RAM
                                randomSeed = 12345,
                                corType = "pearson",
                                power = softPower,
                                consensusQuantile = 0.3,
                                networkType = "unsigned",
                                TOMType = "signed",
                                TOMDenom = "min",
                                scaleTOMs = TRUE, scaleQuantile = 0.8,
                                sampleForScaling = TRUE, sampleForScalingFactor = 1000,
                                useDiskCache = TRUE, chunkSize = NULL,
                                deepSplit = 4,
                                pamStage=FALSE,
                                detectCutHeight = 0.995, minModuleSize = 50,
                                mergeCutHeight = 0.2,
                                saveConsensusTOMs = TRUE,
                                consensusTOMFilePattern = paste0(outputdir,"ConsensusTOM-block.%b.rda"))
  saveRDS(net, file=paste0(outputdir, "net.rds"))
  
} else {
  net = readRDS(file=paste0(outputdir, "net.rds"))
}



consMEs = net$multiMEs;
moduleLabels = net$colors;


# Convert the numeric labels to color labels
moduleColors = as.character(moduleLabels)
consTree = net$dendrograms[[1]];



# module eigengenes
MEs=moduleEigengenes(multiExpr[[1]]$data, colors = moduleColors, nPC=1)$eigengenes
MEs=orderMEs(MEs)
meInfo<-data.frame(rownames(datExpr), MEs)
colnames(meInfo)[1]= "SampleID"

# intramodular connectivity
KMEs<-signedKME(datExpr, MEs,outputColumnName = "kME",corFnc = "bicor")

# compile into a module metadata table
geneInfo=as.data.frame(cbind(colnames(datExpr),moduleColors, KMEs))

# how many modules did we get?
nmodules <- length(unique(moduleColors))

# merged gene symbol column
colnames(geneInfo)[1]= "GeneSymbol"
colnames(geneInfo)[2]= "Initially.Assigned.Module.Color"

# save info
write.csv(geneInfo,file=paste0(outputdir,'/geneInfoSigned.csv'))

PCvalues=MEs
```


```{r}
p = WGCNA::plotDendroAndColors(consTree, moduleColors[net$goodGenes], "Module colors", dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05,
                               main = paste0("Microglia  gene dendrogram and module colors"))


ggsave(file=paste0(outputdir, "Microglia_WGCNA_Dendrocolplot_integrated.pdf"))

```
## WGCNA by Genotype_treatment

```{r, fig.width=7, fig.height=4}

plot_df <- cbind(dplyr::select(targets, c(Genotype, Treatment, labels)), PCvalues)
plot_df <- reshape2::melt(plot_df, id.vars = c('Genotype', 'Treatment', "labels"))
plot_df$Genotype <- factor(plot_df$Genotype, levels=c('WT','APPPS1'))
plot_df$Genotype_Treatment <- paste0(plot_df$Genotype, "_", plot_df$Treatment)
plot_df$Genotype_Treatment <- factor(plot_df$Genotype_Treatment, levels=c('WT_Ctrl','WT_Stroke', "APPPS1_Ctrl", "APPPS1_Stroke"))

colors <- sub('ME', '', as.character(levels(plot_df$variable)))
p <- ggplot(plot_df, aes(x=variable, y=value, fill=Treatment)) +
  geom_boxplot(notch=FALSE) +
  RotatedAxis() + ylab('Module Eigengene') + xlab('') +
  theme(
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
  )

p <- p + facet_wrap(~variable+Genotype, scales='free_x', ncol=nmodules)+theme_classic()

p

ggsave(file=paste0(outputdir, "Microgila_WGCNA_Eigenscore_byGenotype_integrated.pdf"), p)
```

## WGCNA violin by condition

```{r, fig.height=2, fig.width=8}
p2 <- ggplot(plot_df, aes(x=variable, y=value, fill=Genotype_Treatment)) +
  geom_violin() +
  # RotatedAxis() + 
  ylab('Module Eigengene') + xlab('') +
  theme(
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
  )

p2 <- p2 + facet_wrap(~variable, scales='free', ncol=nmodules)+theme_classic()
p2
ggsave(file=paste0(outputdir, "Microgila_WGCNA_Eigenscore_byGenotypeTreatment_integrated.pdf"), p2)

```

## WGCNA violin by cluster

```{r, fig.height=2, fig.width=10}
p2 <- ggplot(plot_df, aes(x=variable, y=value, fill=labels)) +
  geom_violin() +
  # RotatedAxis() + 
  ylab('Module Eigengene') + xlab('') +
  theme(
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
  )

p2 <- p2 + facet_wrap(~variable, scales='free', ncol=nmodules)+theme_classic()
p2
ggsave(file=paste0(outputdir, "Microgila_WGCNA_Eigenscore_byCluster_integrated.pdf"), p2)

```

## WGCNA Heatmap  by condition

```{r, fig.width=6, fig.height=4}
reslist = list()
for(M in unique(plot_df$variable)){
  
  Mdata = plot_df[plot_df$variable==M,]
  fit= lm(value~Genotype*Treatment, data = Mdata)
  resfit = summary(fit)
  reslist[[M]] <- c(Genotype_Treatment = resfit$coefficients["GenotypeAPPPS1:TreatmentStroke", "Pr(>|t|)"])
}

ME_avg <- plot_df %>% dplyr::group_by(variable, Treatment, Genotype) %>% summarise(avg_Eigenmodule = mean(value),
                                                                                   sd_Eigenvalue = sd(value),
                                                                                   med_Eigenvalue=median(value)) 
ME_avg$pval=unlist(reslist[ME_avg$variable])
ME_avg$pval_ast=convertpvaltostars(ME_avg$pval)
ME_avg$Genotype_Treatment = paste0(ME_avg$Genotype, "_",ME_avg$Treatment)
ME_avg$Genotype_Treatment <- factor(ME_avg$Genotype_Treatment, levels=c('WT_Ctrl','WT_Stroke', "APPPS1_Ctrl", "APPPS1_Stroke"))
ME_avg$Module = paste0(ME_avg$pval_ast, ME_avg$variable)

ME_avg
openxlsx::write.xlsx(ME_avg, file=paste0(outputdir, "WGCNA_LinModStat_GTxTreatm.xlsx"),rowNames = TRUE)

```


```{r, fig.width=8, fig.height=5}
Meandata = MEs %>% group_by(samples.integrated.filt$Genotype_Treatment) %>% summarise_all(mean) %>% column_to_rownames("samples.integrated.filt$Genotype_Treatment") %>% t()
SDdata = MEs %>% group_by(samples.integrated.filt$Genotype_Treatment) %>% summarise_all(sd)%>% column_to_rownames("samples.integrated.filt$Genotype_Treatment") %>% t()
colLabels=colnames(Meandata)
rowLabels=rownames(Meandata)
rowLabels=paste0(convertpvaltostars(unlist(reslist[rowLabels])), " ",rowLabels=rownames(Meandata))


textlab = paste0(signif(Meandata,2), "\n", "(", signif(SDdata, 2), ")")

pdf(paste0(outputdir, "WGCNA_labeled_Heatmap_integrated.pdf"))
par(mar=c(5,8,5,5))
labeledHeatmap(Meandata, xLabels = colLabels, 
               yLabels = rowLabels,
               yColorLabels = T,
               colors = viridis(50),
               setStdMargins = FALSE, textMatrix = textlab,
               main = "Mean(SD) Eigengene-expression");
dev.off()

par(mar=c(5,8,5,5))
labeledHeatmap(Meandata, xLabels = colLabels, 
               yLabels = rowLabels,
               yColorLabels = T,
               colors = viridis(50),
               setStdMargins = FALSE, textMatrix = textlab,
               main = "Mean(SD) Eigengene-expression");

```

## WGCNA Heatmap  by cluster

```{r, fig.width=6, fig.height=4}
reslist = list()
for(M in unique(plot_df$variable)){
  
  Mdata = plot_df[plot_df$variable==M,]
  fit= aov(value~labels, data = Mdata)
  resfit = summary(fit)
  reslist[[M]] <- c(Clusters = resfit[[1]]["labels", "Pr(>F)"])
}

ME_avg <- plot_df %>% dplyr::group_by(variable, labels) %>% summarise(avg_Eigenmodule = mean(value),
                                                                      sd_Eigenvalue = sd(value),
                                                                      med_Eigenvalue=median(value)) 
ME_avg$pval=unlist(reslist[ME_avg$variable])
ME_avg$pval_ast=convertpvaltostars(ME_avg$pval)
ME_avg$Module = paste0(ME_avg$pval_ast, ME_avg$variable)

ME_avg
openxlsx::write.xlsx(ME_avg, file=paste0(outputdir, "WGCNA_LinModStat_ByCluster.xlsx"),rowNames = TRUE)

```

```{r, fig.width=10, fig.height=5}
Meandata = MEs %>% group_by(samples.integrated.filt$labels) %>% summarise_all(mean) %>% column_to_rownames("samples.integrated.filt$labels") %>% t()
SDdata = MEs %>% group_by(samples.integrated.filt$labels) %>% summarise_all(sd)%>% column_to_rownames("samples.integrated.filt$labels") %>% t()
colLabels=colnames(Meandata)
rowLabels=rownames(Meandata)
rowLabels=rownames(Meandata)
rowLabels=paste0(convertpvaltostars(unlist(reslist[rowLabels])), " ",rowLabels=rownames(Meandata))
textlab = paste0(signif(Meandata,2), "\n", "(", signif(SDdata, 2), ")")

pdf(paste0(outputdir, "WGCNA_labeled_Heatmap_byCluster_integrated.pdf"))
par(mar=c(5,8,5,5))
labeledHeatmap(Meandata, xLabels = colLabels, 
               yLabels = rowLabels,
               colorLabels = TRUE,
               colors = viridis(50),
               setStdMargins = FALSE, textMatrix = textlab,
               main = "Mean(SD) Eigengene-expression");
dev.off()

par(mar=c(5,8,5,5))
labeledHeatmap(Meandata, xLabels = colLabels, 
               yLabels = rowLabels,
               colorLabels = TRUE,
               colors = viridis(50),
               setStdMargins = FALSE, textMatrix = textlab,
               main = "Mean(SD) Eigengene-expression");


```

## Heatmaps Genewise per module 

```{r}

Modules = unique(moduleColors)
Modules = Modules[! Modules %in% "grey"] # drop grey one

for(m in Modules){
  
  p = DoHeatmap(samples.integrated.filt, features = rownames(samples.integrated.filt)[moduleColors==m], slot="data",
                group.by = "Genotype_Treatment")+scale_fill_viridis()
  
  ggsave(paste0(paste0(outputdir, "WGCNAGeneWise_ExpressionModule", m, "_integrated.pdf")), p)
}


```

## Heatmpas top 50 genes per module Conditions wise

```{r, fig.width=14, fig.height=14}
genelist = split(names(net$colors), net$colors)
gene_univers = names(net$colors)
plotlist = list()

reads = rowSums(JoinLayers(samples.integrated.filt, assay = "RNA") %>%    GetAssayData(assay = "RNA", layer="data") )


for (m in names(genelist)){
  
  targets = genelist[[m]]
  targets <- reads[targets] %>% sort(., decreasing = T)
  targets <- names(targets)[1:50]
  
  p<-DoHeatmap(samples.integrated.filt,features = targets, 
               group.by = "Genotype_Treatment", slot = "data")
  ggsave(p, file=paste0(outputdir, "WGCNA_Module_",m,"_allgenes_Heatmaps_integrated.pdf"))
  plotlist[[m]]<-p
}

p<- ggarrange(plotlist = plotlist, labels = names(genelist))
p
ggsave(file=paste0(outputdir, "WGCNA_Module_Heatmaps_integrated.pdf"), p)

names(plotlist)

```

```{r, fig.width=5, fig.height=5}
module = "green"
featureslist=c("Apoe", "Axl", "Cst7", "Tyrobp","Cd63", "Clec7a", "H2-D1", "Lyz2", "Npc2")
pdf(paste0(outputdir, "WGCNA_", module, "selectedgenes_heatmap_integrated.pdf"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "Genotype_Treatment", slot = "data")
dev.off()
```

```{r, fig.width=5, fig.height=5}
module = "yellow"
featureslist=c("Trem2", "Csf1r", "C1qa", "C1qb", "C1qc", "Ctss",
               "Ctsl", "Ctsh", "Cyba", "Lamp1")
pdf(paste0(outputdir, "WGCNA_", module, "selectedgenes_heatmap_integrated.pdf"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "Genotype_Treatment", slot = "data")
dev.off()
```

```{r, fig.width=5, fig.height=5}
module = "turqouise"
featureslist=c("Olfr1307", "Olfr635", "Olfr1336", "Olfr46", 
               "Olfr344", "Fgd4", "Pkib", "Shisa5", "Senp7", "Snhg14")
pdf(paste0(outputdir, "WGCNA_", module, "selectedgenes_heatmap_integrated.pdf"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "Genotype_Treatment", slot = "data")
dev.off()
```

```{r, fig.width=5, fig.height=5}
module = "grey"
featureslist=c("Adam10", "Abca1", "CD74","Arpc1b", "Eef2",
               "Itgb2", "Pdia3", "Aldoa", "Cfh", "Ssh2")
pdf(paste0(outputdir, "WGCNA_", module, "selectedgenes_heatmap_integrated.pdf"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "Genotype_Treatment", slot = "data")
dev.off()
```


## Heatmpas top 50 genes per module Cluster wise

```{r, fig.width=14, fig.height=14}
plotlist = list()

reads = rowSums(JoinLayers(samples.integrated.filt, assay = "RNA") %>%    GetAssayData(assay = "RNA", layer="data") )


for (m in names(genelist)){
  
  targets = genelist[[m]]
  targets <- reads[targets] %>% sort(., decreasing = T)
  targets <- names(targets)[1:50]
  
  p<-DoHeatmap(samples.integrated.filt, features = targets, 
               group.by = "labels", slot = "data")
  ggsave(p, file=paste0(outputdir, "WGCNA_Module_",m,"_Heatmaps_byCluster_integrated.pdf"))
  plotlist[[m]]<-p
}

p<- ggarrange(plotlist = plotlist, labels = names(genelist))
p
ggsave(file=paste0(outputdir,"WGCNA_Module_Heatmaps_byCluster_integrated.pdf"), p)

names(plotlist)

```


```{r, fig.width=5, fig.height=5}
module = "green"
featureslist=c("Apoe", "Axl", "Cst7", "Tyrobp","Cd63", "Clec7a", "H2-D1", "Lyz2", "Npc2")
pdf(paste0(outputdir, "WGCNA_", module, "selectedgenes_heatmap_byCluster_integrated.pdf"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "labels", slot = "data")
dev.off()
```

```{r, fig.width=5, fig.height=5}
module = "yellow"
featureslist=c("Trem2", "Csf1r", "C1qa", "C1qb", "C1qc", "Ctss",
               "Ctsl", "Ctsh", "Cyba", "Lamp1")
pdf(paste0(outputdir, "WGCNA_", module, "selectedgenes_heatmap_byCluster_integrated.pdf"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "labels", slot = "data")
dev.off()
```

```{r, fig.width=5, fig.height=5}
module = "turqouise"
featureslist=c("Olfr1307", "Olfr635", "Olfr1336", "Olfr46", 
               "Olfr344", "Fgd4", "Pkib", "Shisa5", "Senp7", "Snhg14")
pdf(paste0(outputdir, "WGCNA_", module, "selectedgenes_heatmap_byCluster_integrated.pdf"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "labels", slot = "data")
dev.off()
```

```{r, fig.width=5, fig.height=5}
module = "grey"
featureslist=c("Adam10", "Abca1", "CD74","Arpc1b", "Eef2",
               "Itgb2", "Pdia3", "Aldoa", "Cfh", "Ssh2")
pdf(paste0(outputdir, "WGCNA_", module, "selectedgenes_heatmap_byCluster_integrated.pdf"))
DoHeatmap(samples.integrated.filt,features = featureslist, 
          group.by = "labels", slot = "data")
dev.off()
```


## WGCNA GO terms

```{r, out.width='130%',out.height='100%'}


pvallist = unlist(reslist)
MOI = gsub("ME", "", names(pvallist)) %>% gsub(".Clusters", "", .)
# MOI = MOI[pvallist<0.05]

MOI <- MOI[MOI != "grey"]

# if you want to have the affected genes set return_genelist = T; this takes roughly 4-8 hours 

Gores = getGOresults(geneset = genelist[MOI], domain_scope = "known",
                     genereference = gene_univers, 
                     return_genelist = T,
                     organism = "mmusculus")

display_tab(Gores$result)
openxlsx::write.xlsx(Gores$result%>% dplyr::select(-parents), file=paste0(outputdir, "WGCNA_GO_terms.xlsx"),rowNames = TRUE)

plotlist <- list()
for(m in unique(Gores$result$query)){
  p <- GOplot(Gores$result[Gores$result$query ==m,], N = 10, Title = m)
  plotlist[[m]] <- p
}

```

```{r, fig.width=15, fig.height=6}
p<- ggarrange(plotlist = plotlist)
p 
ggsave(file=paste0(outputdir,"WGCNA_GO_terms.pdf"), p)
```



# Exdendet Plots and DAta inspection

## Dot plot showing top markers for condition clusters, same paper Fig 2f

### MG related genes 

```{r, fig.width=18, fig.height=5}

reads = JoinLayers(samples.integrated.filt, assay = "RNA") %>%    
  GetAssayData(assay = "RNA", layer="data") 

reads.filt= reads[cluster.genes,]

datatoplot_mean = aggregate(t(reads.filt), 
                            list(paste0(samples.integrated.filt$labels, "/",samples.integrated.filt$Genotype_Treatment)), 
                            mean, na.rm=T) %>% 
  as.data.frame() %>% 
  column_to_rownames("Group.1") %>% scale()



datatoplot_mean <- datatoplot_mean %>% as.data.frame() %>% 
  rownames_to_column("Var1") %>%
  pivot_longer(cols = -Var1, 
               names_to = "Genes", 
               values_to = "value")
datatoplot_mean$Microglia = factor(gsub("/.*", "", datatoplot_mean$Var1), levels=rev(levels(samples.integrated.filt$labels)))
datatoplot_mean$Genotype_Treatment = factor(gsub(".*/", "", datatoplot_mean$Var1), levels=c("WT_Ctrl", "WT_Stroke", "APPPS1_Ctrl","APPPS1_Stroke")  )                                

colnames(datatoplot_mean) = c("Cluster_Condition", "Gene", "Scaled_mean", "Microglia", "Genotype_Treatment")

datatoplot_perc = aggregate(t(reads.filt), list(paste0(samples.integrated.filt$labels, "/",samples.integrated.filt$Genotype_Treatment)), 
                            function(x){length(which(x>0))/length(x)})

datatoplot_perc <- datatoplot_perc %>% as.data.frame() %>% 
  pivot_longer(cols = -Group.1, 
               names_to = "Genes", 
               values_to = "perc")


datatoplot = cbind(datatoplot_mean, perc=datatoplot_perc$perc)

p <- ggplot(datatoplot, aes(x=Gene, y=Microglia, col=Scaled_mean, size=perc))+geom_point()+scale_size_continuous(range = c(0,3))+
  theme_classic()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+facet_grid(~Genotype_Treatment)

p

ggsave(filename = paste0(outputdir, "Microglia_Dotplot_ClsuterCondition_Genelists_integrated.pdf"), p)


```

### Spatial_mutual genes 


```{r}

Spatial_mutual <- c("S100a6", "Apoc4", "C4b", "Atp1a2", "Apoc1", "Fn1", "Ntrk2") # genes up in both spatial and APPPS1_stroke vs APPPS1

### Dot plot showing top markers for condition clusters, same paper Fig 2f


reads = JoinLayers(samples.integrated.filt, assay = "RNA") %>%    
  GetAssayData(assay = "RNA", layer="data") 

reads.filt= reads[Spatial_mutual,]

datatoplot_mean = aggregate(t(reads.filt), 
                            list(paste0(samples.integrated.filt$labels, "/",samples.integrated.filt$Genotype_Treatment)), 
                            mean, na.rm=T) %>% 
  as.data.frame() %>% 
  column_to_rownames("Group.1") %>% scale() %>% as.data.frame() %>% 
  rownames_to_column("Var1")



datatoplot_mean=datatoplot_mean %>% pivot_longer(cols = -Var1, names_to = "Gene", values_to = "value")
datatoplot_mean$Microglia = factor(gsub("/.*", "", datatoplot_mean$Var1), levels=rev(levels(samples.integrated.filt$labels)))
datatoplot_mean$Genotype_Treatment = factor(gsub(".*/", "", datatoplot_mean$Var1), levels=c("WT_Ctrl", "WT_Stroke", "APPPS1_Ctrl","APPPS1_Stroke")  )                                

colnames(datatoplot_mean) = c("Cluster_Condition", "Gene", "Scaled_mean", "Microglia", "Genotype_Treatment")

datatoplot_perc = aggregate(t(reads.filt), list(paste0(samples.integrated.filt$labels, "/",samples.integrated.filt$Genotype_Treatment)), 
                            function(x){length(which(x>0))/length(x)})

datatoplot_perc=datatoplot_perc %>% pivot_longer(cols=-Group.1, 
                                                 values_to =  "perc", 
                                                 names_to = "Gene")

datatoplot = cbind(datatoplot_mean, perc=datatoplot_perc$perc)

p <- ggplot(datatoplot, aes(x=Gene, y=Microglia, col=Scaled_mean, size=perc))+geom_point()+scale_size_continuous(range = c(0,3))+
  theme_classic()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+facet_grid(~Genotype_Treatment)

p

ggsave(filename = paste0(outputdir, "Microglia_Dotplot_Up_spatial_and_APPPS1vsAPPPS1stroke_integrated.pdf"), p)


```

### Custom Genes 

```{r}

Test_genes <- c("S100a6", "S100a9", "S100a8", "S100a10", "S100a11", "S100b", "S100a7a") # genes up in both spatial and APPPS1_stroke vs APPPS1

### Dot plot showing top markers for condition clusters, same paper Fig 2f


reads = JoinLayers(samples.integrated.filt, assay = "RNA") %>%    
  GetAssayData(assay = "RNA", layer="data") 

reads.filt= reads[Test_genes,]

datatoplot_mean = aggregate(t(reads.filt), 
                            list(paste0(samples.integrated.filt$labels, "/",samples.integrated.filt$Genotype_Treatment)), 
                            mean, na.rm=T) %>% 
  as.data.frame() %>% 
  column_to_rownames("Group.1") %>% scale() %>% as.data.frame() %>% 
  rownames_to_column("Var1")



datatoplot_mean=datatoplot_mean %>% pivot_longer(cols=-Var1, 
                                                 names_to = "Gene", 
                                                 values_to = "value")
datatoplot_mean$Microglia = factor(gsub("/.*", "", datatoplot_mean$Var1), levels=rev(levels(samples.integrated.filt$labels)))
datatoplot_mean$Genotype_Treatment = factor(gsub(".*/", "", datatoplot_mean$Var1), levels=c("WT_Ctrl", "WT_Stroke", "APPPS1_Ctrl","APPPS1_Stroke")  )                                

colnames(datatoplot_mean) = c("Cluster_Condition", "Gene", "Scaled_mean", "Microglia", "Genotype_Treatment")

datatoplot_perc = aggregate(t(reads.filt), list(paste0(samples.integrated.filt$labels, "/",samples.integrated.filt$Genotype_Treatment)), 
                            function(x){length(which(x>0))/length(x)})

datatoplot_perc=datatoplot_perc %>% pivot_longer(
  cols=-Group.1, values_to = "perc", names_to = "Gene")

datatoplot = cbind(datatoplot_mean, perc=datatoplot_perc$perc)

p <- ggplot(datatoplot, aes(x=Gene, y=Microglia, col=Scaled_mean, size=perc))+geom_point()+scale_size_continuous(range = c(0,3))+
  theme_classic()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+facet_grid(~Genotype_Treatment)

p

ggsave(filename = paste0(outputdir, "Microglia_Dotplot_test_integrated.pdf"), p)


```

## cell cylce scoring plots 

```{r}
cell_cycle_markers <- dplyr::left_join(cell_cycle_genes, annotations, by = c("geneID" = "gene_id"))

# Acquire the S phase genes
s_genes <- cell_cycle_markers %>%
  dplyr::filter(phase == "S") %>%
  pull("gene_name")

# Acquire the G2M phase genes        
g2m_genes <- cell_cycle_markers %>%
  dplyr::filter(phase == "G2/M") %>%
  pull("gene_name")


samples.integrated.filt<-
  CellCycleScoring(samples.integrated.filt, layer="scaled.data",
                   s.features = s_genes, 
                   g2m.features = g2m_genes)

a<-DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Phase", pt.size =1)
b<-DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Phase", split.by = "Plate", pt.size =1)
c<-DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Phase", split.by = "Genotype_Treatment", pt.size =1)

a
ggsave(filename = paste0(outputdir, "G2M_Opt_All_genes_all_cells_Clusters_integrated.pdf"), a)
```


```{r, fig.width=18, fig.height=4}
b
ggsave(filename = paste0(outputdir, "G2M_Opt_All_genes_all_cells_Clusters_byPlate_integrated.pdf"), b)
```


```{r fig.width=10, fig.height=4}
c
ggsave(filename = paste0(outputdir, "G2M_Opt_All_genes_all_cells_Clusters_byCondition_integrated.pdf"), c)

```

```{r, fig.width=7, fig.height=7}
p <- DimPlot(samples.integrated.filt, reduction = "umap", group.by = "Phase")
p
ggsave(file=paste0(outputdir,"CClustering_CellPhase.pdf_integrated.pdf"), p)

p <- FeaturePlot_scCustom(samples.integrated.filt,  
                          reduction = "umap",
                          features =  "S.Score",  
                          colors_use = viridis_dark_high)
p
ggsave(file=paste0(outputdir,"CClustering_CellPhaseScore.pdf"), p)

p <- FeaturePlot_scCustom(samples.integrated.filt,  reduction = "umap",features =  "G2M.Score",  colors_use = viridis_dark_high)
p
ggsave(file=paste0(outputdir,"CClustering_CellPhaseG2MScore.pdf"), p)


metadata <- samples.integrated.filt@meta.data

p<- ggplot(metadata, aes(x=labels, y= S.Score))+geom_violin(aes(fill=labels))+geom_boxplot(aes(alpha=1))+theme_classic()
p
ggsave(file=paste0(outputdir,"CClustering_CellPhaseScoreViolin.pdf"), p)
p<- ggplot(metadata, aes(x=labels, y= G2M.Score))+geom_violin(aes(fill=labels))+geom_boxplot(aes(alpha=1))+theme_classic()
p
ggsave(file=paste0(outputdir,"CClustering_CellPhaseG2MScoreViolin.pdf"), p)

pairwise.t.test(metadata$S.Score,  metadata$labels)
pairwise.t.test(metadata$G2M.Score,  metadata$labels)


```

## Additonal Microglia subtype scores

```{r, fig.width=6, fig.height=6}

DefaultAssay(samples.integrated.filt) <- "RNA"

BAM = c("Cd36", "Cd38","Lyve1","Cd206","Cd163", "Cd169")
CAM = c("Emilin2", "Pf4", "Msa4a", "Hp","F5","Mki67")
Lipid = c("Lipe", "Acnat1", "Lratd1", "Or8b44", "Vmn1r13")



samples.integrated.filt<- AddModuleScore(object = samples.integrated.filt, 
                                         assay="RNA", slot="data",
                                         features = list(INF_resp_microglia), 
                                         name = "MG_score_INF_resp_microglia")

samples.integrated.filt<- AddModuleScore(object = samples.integrated.filt, 
                                         assay="RNA", slot="data",
                                         features = list(Cycling_microglia), 
                                         name = "MG_score_Cycling_microglia")
samples.integrated.filt<- AddModuleScore(object = samples.integrated.filt, 
                                         assay="RNA", slot="data",
                                         features = list(Act_resp_microglia), 
                                         name = "MG_score_Act_resp_microglia")
samples.integrated.filt<- AddModuleScore(object = samples.integrated.filt, 
                                         assay="RNA", slot="data",
                                         features = list(Axon_tract_microglia), 
                                         name = "MG_score_Axon_tract_microglia")

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          assay = "RNA", layer="data",
                                          features = list(BAM), name = "BAM_score")

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          assay = "RNA", layer="data",
                                          features = list(CAM), name = "CAM_score")

samples.integrated.filt <- AddModuleScore(object = samples.integrated.filt, 
                                          assay = "RNA", layer="data",
                                          features = list(Lipid), name = "Lipid_score")


featlist=c("MG_score_Axon_tract_microglia1", 
           "MG_score_Act_resp_microglia1", 
           "MG_score_Cycling_microglia1",  
           "MG_score_INF_resp_microglia1",
           "BAM_score1", 
           "CAM_score1", 
           "Lipid_score1"
)

p <- FeaturePlot_scCustom(samples.integrated.filt, 
                          na_cutoff = 0, reduction="umap",
                          features = featlist,
                          colors_use = viridis_dark_high)

```
```{r, fig.width=18, fig.height=14}
p

ggsave(paste0(outputdir, "DimPlots_MGSubtypes_Genes_integrated.pdf"), p)

```


```{r, fig.width=12, fig.height=16}
p <- RidgePlot(samples.integrated.filt,
               features = c("MG_score_Axon_tract_microglia1", 
                            "MG_score_Act_resp_microglia1", 
                            "MG_score_Cycling_microglia1",  
                            "MG_score_INF_resp_microglia1",
                            "BAM_score1", 
                            "CAM_score1", 
                            "Lipid_score1") , 
               group.by = "labels", ncol = 2)

p

ggsave(paste0(outputdir, "Ridge_MGSubtypes_Genes_integrated.pdf"), p)


```


## Additional QC plots final dataset

```{r fig.width=12, fig.height=4}
DefaultAssay(samples.integrated.filt) <- "RNA"

metadata <- samples.integrated.filt@meta.data

metadata$nCount_RNA_finset <- colSums(samples.integrated.filt)
metadata$nFeature_RNA_finset <- colSums((JoinLayers(samples.integrated.filt, assay = "RNA") %>%    GetAssayData(assay = "RNA", layer="data")) >0)

metadata -> samples.integrated.filt@meta.data
p <- VlnPlot(samples.integrated.filt, features = c("nCount_RNA_finset") , group.by = "labels")  +geom_hline(yintercept = 200)
p
p2 <- VlnPlot(samples.integrated.filt, features = c("nFeature_RNA_finset")  , group.by = "labels")  +geom_hline(yintercept = 150)
p2

p3 <- VlnPlot(samples.integrated.filt, features = "percent.mt" , group.by = "labels") + ylim(c(0,10)) +geom_hline(yintercept = 5)
p3


pcomb <- ggarrange(p,p2,p3, common.legend = T, ncol = 3, legend = "right")
pcomb
ggsave(paste0(outputdir, "Qualityplots_integrated.pdf"), pcomb)


```


## Additional Plots DEX 

```{r, fig.width=4, fig.height=6}

genes_to_label_dwn <- c("Siglech", "Ccl6" , "Cst7", "Tyrobp")
genes_to_label_up <- c("H2-D1", "B2m" , "C4b", "Lipe", "Apoe")

all_genes <- c(genes_to_label_up, genes_to_label_dwn)

all_genes <- c(genes_to_label_up, genes_to_label_dwn)

# Get remaining genes not in all_genes
rest_genes <- rownames(DEG_APPPS1_CtrlvsStroke)[!rownames(DEG_APPPS1_CtrlvsStroke) %in% all_genes]

# Combine: labeled genes first, then the rest
df_res <- DEG_APPPS1_CtrlvsStroke[rev(c(all_genes, rest_genes)), ]

df_res <- df_res %>% 
  mutate(sign = ifelse(df_res$avg_log2FC < 0, "down", 
                       ifelse(df_res$avg_log2FC > 0, "up", "no.reg")), 
         sign = ifelse(df_res$p_val_adj > 0.05, "no.reg", sign), 
         sign = ifelse(rownames(df_res) %in% genes_to_label_up, "up", sign),
         sign = ifelse(rownames(df_res) %in% genes_to_label_dwn, "down", sign), 
         goi = ifelse(rownames(df_res) %in% all_genes, 1, 0))


library(ggrepel)

panelA  <- ggplot(df_res, aes(x=avg_log2FC, y=-log10(p_val_adj), 
                              col=sign, alpha=goi, size=as.factor(goi))) + 
  geom_point() + 
  theme_minimal() + 
  geom_hline(yintercept = -log10(0.05), lwd=0.3, linetype="dashed") +
  geom_vline(xintercept = 0, lwd=0.3, linetype="dashed") +
  scale_color_manual(name = "",
                     values = c("up" = "red3", "no.reg" = "gray60", "down" = "royalblue")) +
  guides(alpha = "none", size = "none") +
  geom_label_repel(data = df_res[all_genes, ],
                   aes(label = rownames(df_res[all_genes, ])),
                   size = 3,
                   max.overlaps = Inf,
                   box.padding = 0.5,
                   point.padding = 0.3,
                   segment.color = "gray40",
                   segment.size = 0.3,
                   min.segment.length = 0,
                   force = 2,
                   show.legend = FALSE) +
  theme(legend.position = "top",
        legend.justification = "left") + 
  ggtitle("Bulk differential expression \nall micorglia integrated")

panelA 

```

```{r}
reads = JoinLayers(samples.integrated.filt, assay = "RNA") %>%    
  GetAssayData(assay = "RNA", layer="data") 

subset = samples.integrated.filt$Genotype =="APPPS1"

reads.filt= reads[all_genes, subset]

datatoplot_mean = aggregate(t(reads.filt), 
                            list(paste0(samples.integrated.filt$labels[subset], "/",samples.integrated.filt$Genotype_Treatment[subset])), 
                            mean, na.rm=T) %>% 
  as.data.frame() %>% 
  column_to_rownames("Group.1")



datatoplot_mean <- datatoplot_mean %>% as.data.frame() %>% 
  rownames_to_column("Var1") %>%
  pivot_longer(cols = -Var1, 
               names_to = "Genes", 
               values_to = "value")
datatoplot_mean$Microglia = factor(gsub("/.*", "", datatoplot_mean$Var1), levels=rev(levels(samples.integrated.filt$labels)))
datatoplot_mean$Genotype_Treatment = factor(gsub(".*/", "", datatoplot_mean$Var1), levels=c("WT_Ctrl", "WT_Stroke", "APPPS1_Ctrl","APPPS1_Stroke")  )                                

colnames(datatoplot_mean) = c("Cluster_Condition", "Gene", "Scaled_mean", "Microglia", "Genotype_Treatment")

datatoplot_perc = aggregate(t(reads.filt), list(paste0(samples.integrated.filt$labels[subset], "/",samples.integrated.filt$Genotype_Treatment[subset])), 
                            function(x){length(which(x>0))/length(x)})

datatoplot_perc <- datatoplot_perc %>% as.data.frame() %>% 
  pivot_longer(cols = -Group.1, 
               names_to = "Genes", 
               values_to = "perc")


datatoplot = cbind(datatoplot_mean, perc=datatoplot_perc$perc)

datatoplot$Microglia <- as.character(datatoplot$Microglia)
datatoplot$Gene <- factor(datatoplot$Gene, levels=all_genes, labels=all_genes)





datatoplot <- datatoplot %>%
  group_by(Microglia, Gene) %>%
  mutate(Scaled_Diff = Scaled_mean - mean(Scaled_mean)) %>%
  ungroup()


library(ggh4x)
Microglia_color_palette <- scales::hue_pal()(nlevels(factor(datatoplot$Microglia)))
# Define colors for each Microglia level
strip_colors <- setNames(Microglia_color_palette, levels(factor(datatoplot$Microglia)))

strip <- strip_themed(background_x = elem_list_rect(fill = strip_colors))


panelB <- ggplot(datatoplot, aes(x=Genotype_Treatment, y=Gene, col=Scaled_Diff, size=perc)) +
  geom_point() +
  scale_size_continuous(range = c(1,7)) +
  theme_classic() +
  geom_hline(yintercept = 5.5, lwd=0.5, linetype="dashed", color="gray40")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        strip.text = element_text(face = "bold", color = "white")) +
  facet_grid2(~Microglia, strip = strip) +
  scale_color_continuous(palette = colorRampPalette(rev(RColorBrewer::brewer.pal(5,"RdBu")))(100))+
  ggtitle("Per cluster difference in expression \n normlaized to within cluster mean")

```


```{r, fig.width=15, fig.height=6}


p<- ggarrange(panelA,panelB, widths = c(1.2,4), labels = c("A", "B"))

ggsave(filename = paste0(outputdir, "APPPS1_Bulk_vs_Microglia_Dotplot_ClusterCondition_Genelists_integrated.pdf"), p, width = 15, height = 6)


```




```{r}
rm(list=c("reads", "counts", "counts_run2", "counts_run1", 
          "counts_sel", "results_list", "sce.microglia"))
gc()
```


```{r}
saveRDS(samples.integrated.filt, paste0(outputdir, "SeuratObjectafterProcessing.rds"))
# samples.integrated.filt <- readRDS( paste0(outputdir, "SeuratObjectafterProcessing.rds"))
```



